---
title: "scRNA analysis"
author: "Johan Reimegård"
date: "2023-02-22"
output: html_document

params:
  repositoryPath: ~/git/spidersilk
  geneInfo: Rscripts/loadGeneInfo.R
  extraFunctions: Rscripts/customFunctions.R
  submajormarker: results/scRNA/majorGland/major.glandSpecific.csv
  minorResultDir: results/scRNA/minorGland/integration_add_MarkerGenes
  majorResultDir: results/scRNA/majorGland
  ampullateResultDir: results/scRNA/AmpullateGland
  
  
---

# library
```{r setup2, include=FALSE}

library(Seurat)
library(tidyverse)
library(dplyr) 
library(cowplot)
library(ggplot2)
library(pheatmap)
library(rafalib)
library(ggraph)
library(clustree)
library(ComplexHeatmap)
library(circlize)
library(cluster)


# Source scripts from local project folder
source(here("MinorGland_paper/scripts", "loadGeneInfo.R"))
source(here("MinorGland_paper/scripts", "colorSchemes_minor.R"))
source(here("MinorGland_paper/scripts", "customFunctions.R"))


repositoryPath = "~/git/Rising-Lab/MajorGland_paper"
geneInfo = getGeneInfoFull(repositoryPath = repositoryPath)

#install.packages("wordcloud")
#library(wordcloud)
#install.packages("wordcloud2")
#library(wordcloud2)




```


# Annotation
## Gene annotation
```{r}

geneInfo = read_csv(here("annotation","Lsc.v1.0.3.geneInfo.csv.gz"))


termOrder = c("MiSp","AmSp-like","Spidroins","SpiCE",
             "SignalPeptide","RibosomalProteins","NoSignalPeptide")

geneInfo %>%
  mutate(Term = signalP) %>%
  mutate(Term = replace(Term,Term == "Yes","SignalPeptide")) %>%
  mutate(Term = replace(Term,Term == "No","NoSignalPeptide")) %>%
  mutate(Term = replace(Term, grepl("MiSp",geneName),"MiSp")) %>%
  mutate(Term = replace(Term, grepl("MaSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("AcSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("AgSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("FlSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("TuSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("PySp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("^(rpl|rps)",geneName),"RibosomalProteins")) %>%
  mutate(Term = replace(Term, grepl("^SpiCE",geneNameProtein),"SpiCE")) %>%
  mutate(Term = replace(Term, grepl("^AmSp-like",geneNameProtein),"AmSp-like")) %>%
  mutate(Term = factor(Term, 
                       levels = termOrder) ) ->
  geneInfo



proteinInfo =read_csv(here("annotation","MajorSilk.proteins.1.2.csv"))
  geneInfo %>%
    select(gene_id,geneName) %>%
    inner_join(proteinInfo)->
    proteinInfo
    
  
```
## Protein annotation
```{r}

# Proteomics data
MQ_data.AmpGl.QC.summary = read_csv(here("MinorGland_paper","data",
                                         "proteomics","MQ_data.AmpGl.QC.summary.csv"))


MQ_data.AmpGl.QC.summary %>%
  separate_rows(geneName, sep = ";") %>%
  inner_join(geneInfo) %>%
  select(geneName,Part,gland,relativeIntensity, geneNameProtein,scGeneName,Term) ->
MQ_data.AmpGl.QC.summary.QC  


MQ_data.AmpGl.QC.summary.QC %>% pivot_wider(names_from = c("gland","Part"), values_from = "relativeIntensity", values_fill = 0 )  ->presentInBoth


presentInBoths = MQ_data.AmpGl.QC.summary.QC %>%
  distinct(geneName,gland,Term) %>%
  group_by(geneName,Term) %>%
  summarise(n = n()) %>%
  filter(n == 2)



presentInBothsGlands = MQ_data.AmpGl.QC.summary.QC %>%
  filter(Part == "Gland") %>%
  group_by(geneName) %>%
  summarise(n = n()) %>%
  filter(n == 2)

presentOnlyInOneGland = MQ_data.AmpGl.QC.summary.QC %>%
  filter(Part == "Gland") %>%
  group_by(geneName) %>%
  summarise(n = n()) %>%
  filter(n == 1) %>%
  ungroup() %>%
  inner_join(MQ_data.AmpGl.QC.summary.QC) %>%
  filter(Part == "Gland") %>%
  arrange(geneName)

presentOnlyInMajorGland= presentOnlyInOneGland %>%
  filter(gland == "Major") %>%
  arrange(-relativeIntensity)

presentOnlyInMinorGland= presentOnlyInOneGland %>%
  filter(gland == "Minor") %>%
  arrange(-relativeIntensity)



presentInBothsDope = MQ_data.AmpGl.QC.summary.QC %>%
  filter(Part == "Dope") %>%
  group_by(geneName) %>%
  summarise(n = n()) %>%
  filter(n == 2)

presentInOneDope = MQ_data.AmpGl.QC.summary.QC %>%
  filter(Part == "Dope") %>%
  group_by(geneName) %>%
  summarise(n = n()) %>%
  filter(n == 1) %>%
  inner_join(MQ_data.AmpGl.QC.summary.QC) %>%
  filter(Part == "Dope") %>%
  arrange(geneName)


presentOnlyInMajorDope= presentInOneDope %>%
  filter(gland == "Major") %>%
  arrange(-relativeIntensity)

presentOnlyInMinorDope= presentInOneDope %>%
  filter(gland == "Minor") %>%
  arrange(-relativeIntensity)


presentOnlyMajorBoth = presentOnlyInMajorGland %>%
  filter(geneName %in% presentOnlyInMajorDope$geneName)

presentOnlyMinorBoth = presentOnlyInMinorGland %>%
  filter(geneName %in% presentOnlyInMinorDope$geneName)



```





#Before integration


```{r}



markerGenes.minor =
  read_csv(here("annotation","scRNA","Minor","Minor.cluster.markers.csv.gz"))%>%
  mutate(gland = "Minor") %>%
  group_by(cluster) %>% 
  arrange(p_val_adj , -avg_log2FC) %>%
  select(scGeneName,cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2 )



markerGenes.major =
  read_csv(here("annotation","scRNA","Major","Major.cluster.markers.csv.gz"))%>%
  mutate(gland = "Major") %>%
  group_by(cluster) %>% 
  arrange(p_val_adj , -avg_log2FC) %>%
  select(scGeneName,cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2 )



  


markerGenes = bind_rows(markerGenes.major,markerGenes.minor) 


markerGenes %>%
  group_by(scGeneName, gland) %>%
  summarise(present = n()) %>%
  group_by(scGeneName) %>%
  summarise(n = n()) %>%
  mutate(type = "Common") %>%
  mutate(type = replace(type, n == 1, "Unique"))  %>%
  select(scGeneName,type) %>% 
  inner_join(markerGenes) %>%
  mutate(zone = "Duct") %>%
  mutate(zone = replace(zone, grepl("ZoneA",cluster),"ZoneA")) %>%
  mutate(zone = replace(zone, grepl("ZoneB",cluster),"ZoneB")) %>%
  mutate(zone = replace(zone, grepl("ZoneC",cluster),"ZoneC")) %>%
  mutate(zone = replace(zone, grepl("ZoneABC",cluster),"ZoneABC")) %>%
  mutate(cluster = gsub("ZoneA_SpiCE-LMa3","ZoneA_SpiCE",cluster)) %>%
  mutate(cluster = gsub("ZoneC_SpiCE-LMa1","ZoneC_SpiCE",cluster)) %>%
  mutate(cluster = gsub("ZoneC_SpICE-LMa1","ZoneC_SpiCE",cluster))->
  markerGenes.ind

markerGenes.ind %>% distinct(cluster,gland,zone) ->
  cluster_zone.association


markerGenes %>% group_by(scGeneName) %>% arrange(p_val_adj,-avg_log2FC) %>%
  slice_head(n = 1) %>%
  select(scGeneName,cluster,gland)%>%
  mutate(clusterlong = cluster) %>%
  mutate(cluster = gsub("ZoneA_SpiCE-LMa3","ZoneA_SpiCE",cluster)) %>%
  mutate(cluster = gsub("ZoneC_SpiCE-LMa1","ZoneC_SpiCE",cluster)) %>%
  mutate(cluster = gsub("ZoneC_SpICE-LMa1","ZoneC_SpiCE",cluster)) %>%
  left_join(cluster_zone.association) %>%
  ungroup()-> 
  markerGenes.association

markerGenes %>%
  select(scGeneName,cluster,gland)%>%
  mutate(clusterlong = cluster) %>%
  mutate(cluster = gsub("ZoneA_SpiCE-LMa3","ZoneA_SpiCE",cluster)) %>%
  mutate(cluster = gsub("ZoneC_SpiCE-LMa1","ZoneC_SpiCE",cluster)) %>%
  mutate(cluster = gsub("ZoneC_SpICE-LMa1","ZoneC_SpiCE",cluster)) %>%
  left_join(cluster_zone.association) %>%
  ungroup()-> 
  markerGenes.all.association

markerGenes.all.association %>% inner_join(geneInfo) %>%
  distinct(geneNameProtein,cluster,gland,zone) %>% 
  mutate(glandzone = paste(gland,zone,sep = "_"))->
  markerGenes.all.association

markerGenes.all.association %>% 
  mutate(present = 1) %>%
  distinct(geneNameProtein, glandzone,present) %>%
  pivot_wider(names_from ="glandzone",
              values_from = "present",
              values_fill = 0 ) %>%
  column_to_rownames("geneNameProtein")->
  markerGenes.all.wide


```






### Common
```{r }

markerGenes.ind %>%
  filter(type == "Common") %>%
  inner_join(geneInfo)->
  markerGenes.ind.com

markerGenes.ind.com %>%
  ungroup() %>%
  group_by(scGeneName,gland,zone ) %>%
  arrange(p_val_adj,-avg_log2FC) %>%
  slice_head(n = 1) %>%
    ungroup() ->
  markerGenes.ind.com

markerGenes.ind.com %>%
  mutate(present = 1) %>%
  mutate(zone2 = paste(gland,zone,sep = "_")) %>%
  mutate(zone3 = paste(gland,cluster,sep = "_")) %>%
  ungroup() ->
  markerGenes.ins.HM 


markerGenes.ins.HM %>%
  select(scGeneName,zone2,present ) %>%
  pivot_wider(names_from = "zone2", values_from = "present", values_fill = 0) %>%
  column_to_rownames(var = "scGeneName")->
markerGenes.ind.com

markerGenes.ins.HM %>%
  select(scGeneName,zone2,zone3 ) %>%
  pivot_wider(names_from = "zone2", values_from = "zone3", values_fill = "No") %>%
  column_to_rownames(var = "scGeneName")->
markerGenes.ind.names
````



### Uniuqe    
```{r }

markerGenes.ind %>%
  filter(type != "Common") %>%
  inner_join(geneInfo)->
  markerGenes.ind.uni

markerGenes.ind.uni %>%
  ungroup() %>%
  group_by(scGeneName,gland,zone ) %>%
  arrange(p_val_adj,-avg_log2FC) %>%
  slice_head(n = 1) %>%
    ungroup() ->
  markerGenes.ind.uni


markerGenes.ind.uni %>%
  group_by(gland, zone) %>%
  arrange(gland,zone, p_val_adj,-avg_log2FC)  %>%
  filter(zone != "Duct") %>%
  filter(zone != "ZoneABC") 


```


## Supplementary tables 5 & 6

```{r}
markerGenes.minor %>%
  inner_join(geneInfo) %>%
  filter(signalP == "Yes") %>%
  mutate(cluster = factor(cluster, levels = names(scCellOrderColorMinor))) %>%
select(geneNameProtein, cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2) %>%
  arrange(cluster, p_val_adj , -avg_log2FC) ->
  supp.table.5



# Functional Annotation results
write_csv(supp.table.5, 
          here("MinorGland_paper", "results", "supplementaryTables", "Suppl.table.5.csv.gz"))




markerGenes.major %>%
  inner_join(geneInfo) %>%
  filter(signalP == "Yes") %>%
  mutate(cluster = factor(cluster, levels = rev(names(scCellOrderColorMajor)))) %>%
select(geneNameProtein, cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2) %>%
  arrange(cluster, p_val_adj , -avg_log2FC) ->
  supp.table.6

write_csv(supp.table.6, 
          here("MinorGland_paper", "results", "supplementaryTables", "Suppl.table.6.csv.gz"))




```

## Figure 4 

### Figure 4A

```{r}

# Minor data --- you can get this data from: 

minor.sct.integrated <- readRDS(file = here("MinorGland_paper","data",
             "scRNA","minor",  
             "scRNA.minor.2024.rds"))



DefaultAssay(minor.sct.integrated) <- "Recluster"
minor.sct.integrated$Batch <- "Minor_B1"
## add one new column named CellTypes
celltype <- rep("unsure",ncol(minor.sct.integrated))
celltype[minor.sct.integrated$Re_celltype == "Tail_MiSpabd"] <- "ZoneA_MiSpabd"
celltype[minor.sct.integrated$Re_celltype == "Sac_MiSpc"] <- "ZoneB_MiSpc"
celltype[minor.sct.integrated$Re_celltype == "Sac_1"] <- "ZoneC_SpICE-LMa1"
celltype[minor.sct.integrated$Re_celltype == "Duct_1"] <- "Duct_2"
celltype[minor.sct.integrated$Re_celltype == "Duct_2"] <- "Duct_1"
minor.sct.integrated <- AddMetaData(minor.sct.integrated, celltype, col.name = "Latest_CellTypes")
minor.sct.integrated$CellTypes <- paste0("Minor_",minor.sct.integrated$Latest_CellTypes)


minor.sct.integrated$Latest_CellTypes = factor(minor.sct.integrated$Latest_CellTypes,
                                              levels = names(scCellOrderColorMinor))
minor.sct.integrated = SetIdent(minor.sct.integrated,value =  "Latest_CellTypes")  

DimPlot(minor.sct.integrated, cols =scCellOrderColorMinor ) -> minor.dimplot
minor.dimplot

ggsave(here("MinorGland_paper","figures","figure4","Figure4A.pdf"),
       minor.dimplot,
       width = 8,
       height = 10)

```

### Figure 4B

```{r } 


gower.dist <- daisy(t(markerGenes.ind.com),metric = "manhattan")  
aggl.clust.c <- hclust(gower.dist, method = "ward")
plot(aggl.clust.c)

gower.dist.genes <- daisy((markerGenes.ind.com),metric = "manhattan")  
aggl.clust.genes <- hclust(gower.dist.genes, method = "ward")
plot(aggl.clust.genes, main = "Hierarchical Clustering Dendrogram", sub = "", xlab = "Samples", ylab = "Height")




ComplexHeatmap::Heatmap(markerGenes.ind.names,
                        cluster_columns = aggl.clust.c ,
                        col = scCellCelltypeOrderColorBoth,
                        rect_gp = gpar(col = "white", lwd = 0.1),
                        show_row_names = FALSE,
                        cluster_rows = aggl.clust.genes,
                        column_split = 5,split = 4) -> Fig4B


pdf(here("MinorGland_paper","figures","figure4","Figure4B.pdf"),height = 4, width = 4)
  Fig4B
dev.off()

Fig4B


```



### Figure 4C
```{r marker genes}
#Major data --- you can get this data from: Box/bridgespider/scRNA/LarSco.combined_filtered.integrated.MajorSpecific.SCT.2.rds


major.sct.integrated = readRDS( here("MinorGland_paper","data",
             "scRNA","major",  
             "scRNA.major.2024.rds"))



Major_data <- major.sct.integrated
DefaultAssay(Major_data) <- "integrated"
major.sct.integrated = SetIdent(major.sct.integrated,value =  "clusterZones")  

DimPlot(major.sct.integrated, reduction = "umap", cols = scCellOrderColorMajor) -> major.dimplot

ggsave(here("MinorGland_paper","figures","figure4","Figure4C.pdf"),
       major.dimplot,
       width = 8,
       height = 10)

```

## Figure 4D


```{r} 
# PCA loading for minor parts
minor.parts.PCA = read_csv(here("MinorGland_paper","results","bulkRNA","minor.parts.PCA.csv.gz"))


minor.parts.PCA%>%
  mutate(Tissue = gsub("Sac_distal_Duct","Duct",Tissue)) %>%
  ggplot(aes(PC1,PC2,fill = Tissue))+
  geom_point(size = 1, shape = 23)+
  theme_classic()+
  scale_color_manual(values = BulkPartMinor2)+
  scale_fill_manual(values = BulkPartMinor2)+
theme(
    text = element_text(size = 10)  # Increase overall text size
  ) ->
  bulk.PCA.plot.2d

ggsave(here("MinorGland_paper","figures","figure4","Figure4D.pdf"),
       bulk.PCA.plot.2d,
       width = 5,
       height = 3)







```


## Figure 4E

```{r}



markerGenes.minor =
  read_csv(here("annotation","scRNA","Minor","Minor.cluster.markers.csv.gz")) %>%
  mutate(cluster = replace(cluster,cluster =="ZoneC" ,"ZoneC_SpiCE" ))  %>% 
  mutate(gland = "Minor") %>%
  group_by(scGeneName) %>%
  arrange(p_val_adj , -avg_log2FC) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  group_by(cluster) %>%
  arrange(p_val_adj , -avg_log2FC) %>%
  select(scGeneName,cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2 )



# PCA loading for minor parts
minor.parts.PCA.loading = read_csv(here("MinorGland_paper","results","bulkRNA","minor.parts.PCA.loading.csv.gz"))
DEseq.minor.parts.location= read_csv(file = here("MinorGland_paper","results",
                                                 "bulkRNA",
                                                 "DEseq.minor.parts.location.csv.gz")
                                     )


markerGenes.minor %>% inner_join(geneInfo) %>%
  select(cluster,geneNameProtein) %>%
  inner_join(minor.parts.PCA.loading) %>%
  pivot_longer(names_to = "PCs", values_to = "loading", c(PC1,PC2)) %>%
  ggplot(aes(PCs,loading,fill = cluster))+
  geom_boxplot()+
  theme_minimal()+ 
  scale_fill_manual(values = scCellOrderColorMinor)-> 
  scRNA.PCA.loading.plot


markerGenes.minor %>% inner_join(geneInfo) %>% 
  filter(geneNameProtein %in% c("MiSpa","MiSpb","MiSpc","MiSpd",                              "SpiCE-LMa1","SpiCE-LMa4","SpiCE-LMa6"))%>%
  select(cluster,geneNameProtein) %>%
  inner_join(minor.parts.PCA.loading) %>%
  mutate(cluster = factor(cluster,levels = names(scCellOrderColorMinor)))->
  markerGenes

library(ggrepel)

markerGenes.minor %>%
  inner_join(geneInfo) %>%
  select(cluster,geneNameProtein) %>%
  inner_join(minor.parts.PCA.loading) %>%
  mutate(cluster = factor(cluster,levels = names(scCellOrderColorMinor))) %>%
  ggplot(aes(PC1,PC2,fill = cluster, label = geneNameProtein))+
  geom_point(alpha = 0.5,size = 1, shape = 22)+
  geom_point(data = markerGenes,alpha = 1,size = 2, shape = 22)+
  geom_text(data = markerGenes) +
  theme_classic()+
  theme(
    text = element_text(size = 10)  # Increase overall text size
  )  +
  scale_color_manual(values = scCellOrderColorMinor)+
  scale_fill_manual(values = scCellOrderColorMinor)-> 
  scRNA.PCA.loading.plot.2d

ggsave(here("MinorGland_paper","figures","figure4","Figure4E.pdf"),
       scRNA.PCA.loading.plot.2d,
       width = 5,
       height = 3)



bulk.PCA.plot.2d  + scRNA.PCA.loading.plot.2d + plot_layout(guides = "collect") ->
  bulk.PCA.plot

```





#After integration



## Integrate data 
Done once. Can download the resulting merged file here


```{r}


if(!file.exists(here("MinorGland_paper","data",
                                        "scRNA","integrated",  
                                        "scRNA.major.minor.integrated.2024.rds"))){
  # load major_minor_data
Major_Minor_data <- readRDS(file = paste(params$repositoryPath, params$ampullateResultDir, "Major_Minor_data.rds", sep = "/"))

ifnb.list <- SplitObject(Major_Minor_data, split.by = "Batch")
ifnb.list <- lapply(X = ifnb.list, FUN = SCTransform)
features <- SelectIntegrationFeatures(object.list = ifnb.list, nfeatures = 3000)
ifnb.list <- PrepSCTIntegration(object.list = ifnb.list, anchor.features = features)

# normalize and identify variable features for each dataset independently
immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, normalization.method = "SCT",
    anchor.features = features)


major.minor.sct.integrated <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")

major.minor.sct.integrated <- RunPCA(major.minor.sct.integrated)
major.minor.sct.integrated <- RunUMAP(major.minor.sct.integrated, dims = 1:30)
DimPlot(major.minor.sct.integrated, 
        reduction = "umap", group.by = c("clusterZones", "Latest_CellTypes"))

major.minor.sct.integrated@meta.data$cellID = rownames(major.minor.sct.integrated@meta.data) 
major.minor.sct.integrated@meta.data %>% select(cellID,clusterZones,Latest_CellTypes) %>%
  mutate(cellType = Latest_CellTypes) %>%
  mutate(cellType = replace(cellType, is.na(cellType), 
                major.minor.sct.integrated@meta.data$clusterZones[which(!is.na(clusterZones))])) %>%
  select(cellID,cellType) ->md.2
major.minor.sct.integrated = addMetaDataToSingleCellData(major.minor.sct.integrated,md.2)

major.minor.sct.integrated <- FindNeighbors(major.minor.sct.integrated, dims = 1:20)
major.minor.sct.integrated <- FindClusters(major.minor.sct.integrated, resolution = 0.1)

major.minor.sct.integrated@meta.data %>%
  ggplot(aes(x = seurat_clusters,fill= cellType )) + geom_bar()



#Clean the metadata  
MD = major.minor.sct.integrated@meta.data 



#Annotate celltypes
MD %>%
  select(c(1:9,11,24,25,29)) %>%
  mutate(GlandType = replace(GlandType, is.na(GlandType),"Major"))%>%
  mutate(rn = cellID) %>%
  mutate(CellTypes = gsub("Minor_ZoneC","Minor_ZoneC_SpiCE",CellTypes)) %>%
  mutate(CellTypes = gsub("Ma_","Major_",CellTypes)) %>%
  mutate(CellTypes = gsub("ZoneA_SpiCE-LMa3","ZoneA_SpiCE",CellTypes)) %>%
  mutate(CellTypes = gsub("ZoneC_SpiCE-LMa1","ZoneC_SpiCE",CellTypes)) %>%
  mutate(CellTypes = factor(CellTypes,
                            levels = names(scCellCelltypeOrderColorBoth)))   ->  
  MD

# Annotate zones
MD %>% 
  mutate(Zone = "ZoneA") %>%
  mutate(Zone = replace(Zone,seurat_clusters == 1,"Duct" )) %>%
  mutate(Zone = replace(Zone,seurat_clusters == 2,"ZoneA" )) %>%
  mutate(Zone = replace(Zone,seurat_clusters == 3,"ZoneC" )) %>%
  mutate(Zone = replace(Zone,seurat_clusters == 4,"ZoneABC" )) %>%
  mutate(Zone = replace(Zone,seurat_clusters == 5,"ZoneB" )) %>%
  mutate(Zone = factor(Zone,
                       levels = names(scCellZoneOrderColorBoth)))->
  MD

MD = MD[rownames(major.minor.sct.integrated@meta.data),]

major.minor.sct.integrated@meta.data = MD


saveRDS(major.minor.sct.integrated, here("MinorGland_paper","data",
                                        "scRNA","integrated",  
                                        "scRNA.major.minor.integrated.2024.rds"))

}







```



## Remove uncertain cells
```{r}
## Fix the  metaData

major.minor.sct.integrated = readRDS(here("MinorGland_paper","data",
                                        "scRNA","integrated",  
                                        "scRNA.major.minor.integrated.2024.rds"))





#Filter the data 
## Removing the small clusters 

MD.clean = major.minor.sct.integrated@meta.data %>%
  filter(seurat_clusters %in%c(0,1,2,3,4,5))
# Keep only overlapping annotation

MD.clean %>%
  filter(Zone == "ZoneA" &
           CellTypes %in% c("Minor_ZoneA_MiSpabd","Major_ZoneA_MaSp1","Major_ZoneA_MaSp2","Major_ZoneA_SpiCE"))->zoneA

MD.clean %>%
  filter(Zone == "ZoneB" &
           CellTypes %in% c("Minor_ZoneB_MiSpc","Major_ZoneB_MaSp3"))->zoneB

MD.clean %>%
  filter(Zone == "ZoneC" &
           CellTypes %in% c("Major_ZoneC_SpiCE","Minor_ZoneC_SpiCE"))->zoneC

MD.clean %>%
  filter(Zone == "ZoneABC" &
           CellTypes %in% c("Major_ZoneABC"))->zoneABC

MD.clean %>%
  filter(Zone == "Duct" &
           CellTypes %in% c("Major_Duct_1","Major_Duct_2","Minor_Duct_1","Minor_Duct_2"))->Duct

MD.clean = bind_rows(zoneA,zoneB,zoneABC,zoneC,Duct) 



major.minor.sct.integrated.clean = subset(major.minor.sct.integrated, 
                                          cellID %in% c(MD.clean$cellID))

MD.clean = MD.clean[rownames(major.minor.sct.integrated.clean@meta.data),]

major.minor.sct.integrated.clean@meta.data = MD.clean

```


## Figure 6
### Figure 6A
```{r}
colors = brewer.pal(n = 8, name = "Dark2")

Fig6A = DimPlot(major.minor.sct.integrated, cols =colors, label = TRUE )

ggsave(here("MinorGland_paper","figures","figure6","Figure6A.pdf"),
       Fig6A,
       width = 5,
       height = 4)

Fig6A

````


### Figure 6B

```{r }



major.minor.sct.integrated@meta.data %>%
  ggplot(aes(x = seurat_clusters,fill= CellTypes )) + geom_bar() +  scale_fill_manual(values = scCellCelltypeOrderColorBoth) -> fig6B

ggsave(here("MinorGland_paper","figures","figure6","Figure6B.pdf"),
       fig6B,
       width = 5,
       height = 4)

fig6B

```


### Figure 6C


```{r}

major.minor.sct.integrated@meta.data %>%
  ggplot(aes(x = seurat_clusters,fill= GlandType )) + geom_bar() +  scale_fill_manual(values = c("black","grey")) -> fig6C


ggsave(here("MinorGland_paper","figures","figure6","Figure6C.pdf"),
       fig6C,
       width = 5,
       height = 4)

fig6C

```


### Figure 6D
```{r}

DimPlot(major.minor.sct.integrated, reduction = "umap",group.by = "CellTypes",cols = scCellCelltypeOrderColorBoth) -> before.dimplot

DimPlot(major.minor.sct.integrated.clean, reduction = "umap",group.by = "CellTypes",cols = scCellCelltypeOrderColorBoth) -> after.dimplot

before.dimplot+after.dimplot+ plot_layout(guides = "collect")


DimPlot(major.minor.sct.integrated.clean, reduction = "umap",group.by = "CellTypes",cols = scCellCelltypeOrderColorBoth, split.by = "GlandType") -> Fig6D

ggsave(here("MinorGland_paper","figures","figure6","Figure6D.pdf"),
       Fig6D,
       width = 9,
       height = 4)

Fig6D
```






## Figure 7


### Independent marker genes
```{r}
markerGenes.major.all =
  read_csv(here("annotation","scRNA","Major","Major.cluster.markers.csv.gz"))%>%
  mutate(gland = "Major") %>%
  select(scGeneName,cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2 ) %>%
  filter(cluster %in% c("ZoneA_MaSp2","ZoneA_MaSp1",
                        "ZoneA_SpiCE-LMa3","ZoneB_MaSp3",
                        "ZoneC_SpiCE-LMa1")) 



markerGenes.minor.all =
  read_csv(here("annotation","scRNA","Minor","Minor.cluster.markers.csv.gz")) %>%
  mutate(cluster = replace(cluster,cluster =="ZoneC" ,"ZoneC_SpiCE" ))  %>% 
  mutate(gland = "Minor") %>%
  select(scGeneName,cluster,gland, p_val_adj, avg_log2FC,pct.1, pct.2 )%>%
  filter(cluster %in% c("ZoneA_MiSpabd","ZoneA_MaSp1",
                        "ZoneA_SpiCE-LMa3","ZoneB_MiSpc",
                        "ZoneC_SpICE-LMa1")) 




```

###Common Zone

```{r}


if(!file.exists(here("MinorGland_paper", "data", 
                     "scRNA","integrated",
                     "scRNA.major.minor.integrated.2024.common.markers.csv.gz"))){

seeds =c(58, 83, 54, 68, 53,  4, 71, 11, 75, 90)

results.common = list()
for(i in 1:10){
  set.seed(seeds[i])
  MD.clean %>% group_by(GlandType,Zone) %>%
    slice_sample(n = 300,replace = FALSE) %>%
    filter( Zone %in% c("ZoneA","ZoneB","ZoneC")) -> 
    MD.clean.sub
  
  
  major.minor.sct.integrated.clean.sub = subset(major.minor.sct.integrated.clean, cellID %in% c(MD.clean.sub$cellID))
  
  major.minor.sct.integrated.clean.sub = SetIdent(major.minor.sct.integrated.clean.sub,value =  "Zone")  
  
  
  major.minor.sct.integrated.clean.sub.sct =
    PrepSCTFindMarkers(major.minor.sct.integrated.clean.sub)
  markerGenes.integrated.both.Zones.sct =
    FindAllMarkers(major.minor.sct.integrated.clean.sub.sct,
                   logfc.threshold = 1 , only.pos = FALSE,assay = "SCT") 
  markerGenes.integrated.both.Zones.sct$run = i
  
  results.common[[i]] = markerGenes.integrated.both.Zones.sct
  
}


bind_rows(results.common) %>%
  dplyr::filter(gene %in% markerGenes.major.all$scGeneName) %>%
  dplyr::filter(gene %in% markerGenes.minor.all$scGeneName) %>%
  dplyr::filter(avg_log2FC  > 1) %>%
  dplyr::rename(scGeneName = "gene") %>%
  inner_join(geneInfo)-> common.sct.10

common.sct.10 %>% 
  distinct(scGeneName,run) %>%
  group_by(scGeneName) %>%
  summarise(n = n()) %>%
  filter(n >5) ->common.sct.10.found10
  

common.sct.10 %>% 
  distinct(scGeneName,run) %>%
  group_by(scGeneName) %>%
  summarise(n = n())->common.sct.10.found10
  


markerGenes.integrated.both.Zones.sct %>%
  filter(gene %in% markerGenes.major.all$scGeneName) %>%
  filter(gene %in% markerGenes.minor.all$scGeneName) %>%
  filter(avg_log2FC  > 1) %>%
  rename(scGeneName = "gene") %>%
  inner_join(geneInfo) -> common.sct

common.sct %>%
  inner_join(presentInBothsGlands) %>%
  filter(signalP == "Yes") ->
  common.sct.sp.prot



common.sct.sp.prot %>% select (1:11) -> common.short
write_csv(common.short, here("MinorGland_paper", "data", 
                     "scRNA","integrated",
                     "scRNA.major.minor.integrated.2024.common.markers.csv.gz"))


}

# Read in common markers for the different zones.
read_csv(here("MinorGland_paper", "data", 
                     "scRNA","integrated",
                     "scRNA.major.minor.integrated.2024.common.markers.csv.gz"))%>%
  mutate(signalP ="Yes") ->
  common.sct.sp.prot

```


### Unique Zone

#### ZoneA

```{r}
if(!file.exists(here("MinorGland_paper", "data", 
                     "scRNA","integrated","ZoneA.sct.integrated.rds"))){
  
  major.minor.sct.integrated.clean.zoneA = subset(major.minor.sct.integrated.clean, Zone == "ZoneA")
  
  DefaultAssay(major.minor.sct.integrated.clean.zoneA) <- "RNA"
  major.minor.sct.integrated.clean.zoneA@meta.data %>%
    group_by(CellTypes) %>%
    slice_sample(n = 1000)->
    MD.sampled
  
  major.minor.sct.integrated.clean.zoneA = subset(major.minor.sct.integrated.clean.zoneA,
                                                  cellID %in% MD.sampled$cellID)
  
  ifnb.list <- SplitObject(major.minor.sct.integrated.clean.zoneA, split.by = "orig.ident")
  ifnb.list <- lapply(X = ifnb.list, FUN = SCTransform)
  features <- SelectIntegrationFeatures(object.list = ifnb.list, nfeatures = 3000)
  ifnb.list <- PrepSCTIntegration(object.list = ifnb.list, anchor.features = features)
  
  # normalize and identify variable features for each dataset independently
  immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, normalization.method = "SCT",
                                           anchor.features = features)
  
  
  zoneA.sct.integrated <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")
  
  zoneA.sct.integrated <- RunPCA(zoneA.sct.integrated)
  zoneA.sct.integrated <- RunUMAP(zoneA.sct.integrated, dims = 1:30)
  
  zoneA.sct.integrated <- FindNeighbors(zoneA.sct.integrated, dims = 1:20)
  zoneA.sct.integrated <- FindClusters(zoneA.sct.integrated, resolution = 0.3)
  saveRDS(zoneA.sct.integrated, 
          file = here("MinorGland_paper", "data", 
                      "scRNA","integrated","ZoneA.sct.integrated.rds"))
  
}
```

```{r}

zoneA.sct.integrated = readRDS(here("MinorGland_paper", "data", 
                    "scRNA","integrated","ZoneA.sct.integrated.rds"))
zoneA.sct.integrated@meta.data %>%
  ggplot(aes(x = seurat_clusters,fill= CellTypes )) + geom_bar()

DimPlot(zoneA.sct.integrated, reduction = "umap", group.by = "CellTypes", split.by = "CellTypes", ncol = 2)
DimPlot(zoneA.sct.integrated, reduction = "umap")


if(!file.exists(here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneA.sct.integrated.markers.csv.gz"))){
  Seurat::SetIdent(zoneA.sct.integrated, value = "GlandType")
  zoneA.sct.integrated.sct = PrepSCTFindMarkers(zoneA.sct.integrated)

  zoneAmarkers = Seurat::FindAllMarkers(zoneA.sct.integrated.sct,
                                        only.pos = TRUE, logfc.threshold = 0.25,
                                        min.pct = 0.25, assay = "SCT",
                                        group.by = "GlandType")
  write_csv(zoneAmarkers, here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneA.sct.integrated.markers.csv.gz"))
}
zoneAmarkers = read_csv(here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneA.sct.integrated.markers.csv.gz"))

```

#### ZoneB


```{r}
if(!file.exists(here("MinorGland_paper", "data", 
                     "scRNA","integrated","ZoneB.sct.integrated.rds"))){
  
  
  
  major.minor.sct.integrated.clean.zoneB = subset(major.minor.sct.integrated.clean, Zone == "ZoneB")
  
  DefaultAssay(major.minor.sct.integrated.clean.zoneB) <- "RNA"
  
  ifnb.list <- SplitObject(major.minor.sct.integrated.clean.zoneB, split.by = "GlandType")
  ifnb.list <- lapply(X = ifnb.list, FUN = SCTransform)
  features <- SelectIntegrationFeatures(object.list = ifnb.list, nfeatures = 3000)
  ifnb.list <- PrepSCTIntegration(object.list = ifnb.list, anchor.features = features)
  
  # normalize and identify variable features for each dataset independently
  immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, normalization.method = "SCT",
                                           anchor.features = features)
  
  
  zoneB.sct.integrated <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")
  
  zoneB.sct.integrated <- RunPCA(zoneB.sct.integrated)
  zoneB.sct.integrated <- RunUMAP(zoneB.sct.integrated, dims = 1:30)
  
  zoneB.sct.integrated <- FindNeighbors(zoneB.sct.integrated, dims = 1:20)
  zoneB.sct.integrated <- FindClusters(zoneB.sct.integrated, resolution = 0.3)
  saveRDS(zoneB.sct.integrated, 
          file = here("MinorGland_paper", "data", 
                      "scRNA","integrated","ZoneB.sct.integrated.rds"))
  
}
```


```{r}
zoneB.sct.integrated = readRDS(here("MinorGland_paper", "data", 
                    "scRNA","integrated","ZoneB.sct.integrated.rds"))
zoneB.sct.integrated@meta.data %>%
  ggplot(aes(x = seurat_clusters,fill= CellTypes )) + geom_bar()

DimPlot(zoneB.sct.integrated, reduction = "umap", group.by = "CellTypes", split.by = "CellTypes", ncol = 2)
DimPlot(zoneB.sct.integrated, reduction = "umap")




if(!file.exists(here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneB.sct.integrated.markers.csv.gz"))){
Seurat::SetIdent(zoneB.sct.integrated, value = "GlandType")

zoneB.sct.integrated.sct = PrepSCTFindMarkers(zoneB.sct.integrated)
zoneBmarkers = Seurat::FindAllMarkers(zoneB.sct.integrated.sct, only.pos = TRUE, logfc.threshold = 0.25, min.pct = 0.25, assay = "SCT", group.by = "GlandType")

write_csv(zoneBmarkers, here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneB.sct.integrated.markers.csv.gz"))
}
read_csv(here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneB.sct.integrated.markers.csv.gz")) ->
  zoneBmarkers


```

#### ZoneC


```{r}

if(!file.exists(here("MinorGland_paper", "data", 
                     "scRNA","integrated","ZoneC.sct.integrated.rds"))){

major.minor.sct.integrated.clean.ZoneC = subset(major.minor.sct.integrated.clean, Zone == "ZoneC")

DefaultAssay(major.minor.sct.integrated.clean.ZoneC) <- "RNA"

ifnb.list <- SplitObject(major.minor.sct.integrated.clean.ZoneC, split.by = "GlandType")
ifnb.list <- lapply(X = ifnb.list, FUN = SCTransform)
features <- SelectIntegrationFeatures(object.list = ifnb.list, nfeatures = 3000)
ifnb.list <- PrepSCTIntegration(object.list = ifnb.list, anchor.features = features)

# normalize and identify variable features for each dataset independently
immune.anchors <- FindIntegrationAnchors(object.list = ifnb.list, normalization.method = "SCT",
    anchor.features = features)


ZoneC.sct.integrated <- IntegrateData(anchorset = immune.anchors, normalization.method = "SCT")

ZoneC.sct.integrated <- RunPCA(ZoneC.sct.integrated)
ZoneC.sct.integrated <- RunUMAP(ZoneC.sct.integrated, dims = 1:30)

ZoneC.sct.integrated <- FindNeighbors(ZoneC.sct.integrated, dims = 1:20)
ZoneC.sct.integrated <- FindClusters(ZoneC.sct.integrated, resolution = 0.3)

ZoneC.sct.integrated@meta.data %>%
  ggplot(aes(x = seurat_clusters,fill= CellTypes )) + geom_bar()

saveRDS(ZoneC.sct.integrated, file = here("MinorGland_paper", "data", 
                     "scRNA","integrated","ZoneC.sct.integrated.rds"))

}


```

```{r}
ZoneC.sct.integrated = readRDS("~/GoogleDriveSciLife/BridgeSpider/Data/scData/ZoneC.sct.integrated.rds")
DimPlot(ZoneC.sct.integrated, reduction = "umap", group.by = "CellTypes", split.by = "CellTypes", ncol = 2)
DimPlot(ZoneC.sct.integrated, reduction = "umap")

if(!file.exists(here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneC.sct.integrated.markers.csv.gz"))){
Seurat::SetIdent(ZoneC.sct.integrated, value = "GlandType")

zoneC.sct.integrated.sct = PrepSCTFindMarkers(ZoneC.sct.integrated)
zoneCmarkers = Seurat::FindAllMarkers(zoneC.sct.integrated.sct, only.pos = TRUE, logfc.threshold = 0.25, min.pct = 0.25, assay = "SCT", group.by = "GlandType")
write_csv(zoneCmarkers, here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneC.sct.integrated.markers.csv.gz"))
}

read_csv(here("MinorGland_paper", "data", 
                    "scRNA","integrated",
         "ZoneC.sct.integrated.markers.csv.gz")) ->
  zoneCmarkers
```


#### Combine the three zones
```{r }

zoneAmarkers = read_csv(here("MinorGland_paper", "data", 
                             "scRNA","integrated",
                             "ZoneA.sct.integrated.markers.csv.gz"))
zoneBmarkers = read_csv(here("MinorGland_paper", "data", 
                             "scRNA","integrated",
                             "ZoneB.sct.integrated.markers.csv.gz"))
zoneCmarkers = read_csv(here("MinorGland_paper", "data", 
                             "scRNA","integrated",
                             "ZoneC.sct.integrated.markers.csv.gz"))



bind_rows(zoneAmarkers,
          zoneBmarkers,
          zoneCmarkers) %>%
  rename(scGeneName = "gene") %>% inner_join(geneInfo) ->
  uniqueMarkers


## Filter to keep only the best
uniqueMarkers %>% 
  filter(signalP == "Yes") %>%
  filter(geneName %in% MQ_data.AmpGl.QC.summary.QC$geneName) %>% 
  filter(avg_log2FC > 5) ->
  uniqueMarkers.sp.prot.sig


# Major
uniqueMarkers.sp.prot.sig %>% filter(cluster == "Major") ->
  uniqueMarkers.sp.prot.major.sig

markerGenes.major.all %>%
  filter(scGeneName %in% uniqueMarkers.sp.prot.major.sig$scGeneName) %>%
  group_by(scGeneName) %>%
  arrange(p_val_adj,-avg_log2FC) %>%
  slice_head(n = 1) %>%
  arrange(cluster,-avg_log2FC) %>%
    inner_join(geneInfo) ->
  zones.unique.sp.prot.major

# Minor

uniqueMarkers.sp.prot.sig %>% filter(cluster == "Minor") ->
  uniqueMarkers.sp.prot.minor.sig


markerGenes.minor.all %>%
  filter(scGeneName %in% uniqueMarkers.sp.prot.minor.sig$scGeneName) %>%
  group_by(scGeneName) %>%
  arrange(p_val_adj,-avg_log2FC) %>%
  slice_head(n = 1) %>%
  arrange(cluster,-avg_log2FC) %>%
    inner_join(geneInfo) ->
  zones.unique.sp.prot.minor

# Combine
bind_rows(zones.unique.sp.prot.minor,zones.unique.sp.prot.major) %>%
  select(scGeneName,cluster,gland,geneNameProtein,signalP) %>%
  rename(CellTypes = "cluster")->
  uniqueMarkers.sp.prot.sig.genes





```

### Final list of genes for figure 7
```{r}

## Arrange metadata
MD.clean %>% distinct(GlandType,CellTypes,Zone) %>%
  mutate(CellTypes = gsub("_","-",CellTypes)) %>%
  mutate(glandZone = paste(GlandType, Zone, sep = "_")) %>%
  mutate(GlandType = factor(GlandType, levels = c("Minor","Major"))) %>%
  arrange(Zone,GlandType)->Celltypes.info

Celltypes.info %>%arrange(Zone,GlandType) %>%
  group_by(Zone, GlandType) %>% slice_head(n = 1) %>%
  ungroup() %>%
  select(glandZone) -> glandZoneOrder

Celltypes.info %>%  
  mutate(glandZone = 
           factor(glandZone, levels = glandZoneOrder$glandZone)) ->
  Celltypes.info  



# Unique genes
uniqueMarkers.sp.prot.sig.genes %>%
  mutate(CellTypes = paste(gland,CellTypes,sep = "-")) %>%
  mutate(CellTypes = gsub("-LMa1","",CellTypes)) %>%
  mutate(CellTypes = gsub("-LMa3","",CellTypes)) %>%
  mutate(CellTypes = gsub("SpICE","SpiCE",CellTypes)) %>%
  mutate(CellTypes = gsub("_","-",CellTypes)) %>%
  inner_join(Celltypes.info) ->
  uniqueMarkers.sp.prot.sig.genes


go = data.frame(scGeneName = c("MiSpa","MiSpb","MiSpd",
                               "MaSp1a","MaSp1b","MaSp1c",
                               "MaSp2a","MaSp2b","MaSp2c","MaSp2e","MaSp2f",
                               "Spice-8",
                               "MiSpc","MaSp3a","MaSp3b",
                               "sp-1339#prd34368.1-3","LASC010012",
                               "LASC004595","LASC010016","LASC013229"), 
                geneOrder = c("MiSpa","MiSpb","MiSpd",
                              "MaSp1a","MaSp1b","MaSp1c",
                              "MaSp2a","MaSp2b","MaSp2c","MaSp2e","MaSp2f",
                              "SpiCE-LMa5",
                              "MiSpc","MaSp3a","MaSp3b",
                              "sp-1339#prd34368.1_3","LASC010012",
                              "LASC004595","LASC010016","SpiCE-LMa1")) %>%
  mutate(geneOrder = factor(geneOrder,
                            levels = c("MiSpa","MiSpb","MiSpd",
                                       "MaSp1a","MaSp1b","MaSp1c",
                                       "MaSp2a","MaSp2b","MaSp2c","MaSp2e",
                                       "MaSp2f","SpiCE-LMa5",
                                       "MiSpc","MaSp3a","MaSp3b",
                                       "sp-1339#prd34368.1_3","LASC010012",
                                       "LASC004595","LASC010016","SpiCE-LMa1")))  %>%
  arrange(geneOrder)



go %>% inner_join(uniqueMarkers.sp.prot.sig.genes) ->
  uniqueMarkers.sp.prot.sig.genes



# Common genes

common.sct.sp.prot %>% 
  rename(Zone = "cluster") %>% 
  select(scGeneName,geneNameProtein,Zone) %>%
  select(geneNameProtein, Zone) %>%
  mutate(common = "Yes") ->
  finalList.common

read_csv(file = here("MinorGland_paper","results","bulkRNA",
                      "DEseq.common.27.csv.gz")) ->
  DEseq.common.27

# Bulk genes
DEseq.common.27 %>% select(geneNameProtein, Tissue) %>%
  rename(BulkRNA_Tissue = "Tissue") %>%
  mutate(bulk = "Yes")->
  finalList.bulk

# Unique genes
uniqueMarkers.sp.prot.sig.genes %>%
  select(geneNameProtein, GlandType, Zone) %>%
  rename(unique = "GlandType") ->
  finalList.unique

# 18 major silk proteins
geneInfo %>% filter(silkGroup %in% c("Spice-LMa","AmSp-like","MaSp4","MaSp3","MaSp2","MaSp1")) %>%
  select(geneNameProtein,silkGroup)->
  silkProteins

# Join all minor gland genes

full_join(finalList.common,finalList.unique) %>%
  full_join(finalList.bulk) %>%
  full_join(silkProteins) %>%
  arrange(Zone,common,unique,geneNameProtein)->
  finalList.genes



finalList.genes %>% filter(bulk == "Yes") %>%
  filter(common == "Yes" ) %>%
  arrange(common, Zone)->
  finalList.genes.common.fig7

finalList.genes %>% 
  filter(!is.na(unique) ) %>%
  arrange( Zone,unique)->
  finalList.genes.unique.fig7

```

## Supplementary table 7


```{r}
write_csv(finalList.genes,
          here("MinorGland_paper","results",
                "supplementaryTables",
                "Suppl.Table.7.finalList.of.genes.figure7.csv.gz"))



```


### Figure 7A
```{r}


DEseq.common.27 %>% select(geneNameProtein, Tissue) %>%
  rename(BulkRNA_Tissue = "Tissue") %>%
  mutate(Zone = "ZoneA") %>%
  mutate(Zone = replace(Zone, grepl("Sac_center",BulkRNA_Tissue),"ZoneB")) %>%
  mutate(Zone = replace(Zone, grepl("Sac_distal",BulkRNA_Tissue),"ZoneC")) ->
  finalList.bulk


list(Common_17 = finalList.genes.common.fig7$geneNameProtein,
     MinorBulk27 = finalList.bulk$geneNameProtein,
     MajorSilk18 = silkProteins$geneNameProtein) ->
  figure7VennInfo

three_set_venn_common <- ggVennDiagram(figure7VennInfo)


pdf(here("MinorGland_paper","figures","figure7","Figure7A.pdf"), width = 7, height = 7)
print(three_set_venn_common)
dev.off()
```

### Figure 7B

```{e}

list(Unique_20 = finalList.unique$geneNameProtein,
     MinorBulk27 = finalList.bulk$geneNameProtein,
     MajorSilk18 = c(silkProteins$geneNameProtein)) ->
  figure7VennInfo

three_set_venn_unique <- ggVennDiagram(figure7VennInfo)


pdf(here("MinorGland_paper","figures","figure7","Figure7B.pdf"), width = 7, height = 7)
print(three_set_venn_unique)
dev.off()
````




# Bulk RNA


All analysis made in "FigureBulk.Rmd"



# Visaulise expression

##

```{r}



major.minor.sct.integrated.clean.visual =
  SCTransform(major.minor.sct.integrated.clean,
              new.assay.name = "BothSCT",
              residual.features = c(uniqueMarkers$scGeneName,common.sct.sp.prot$scGeneName))



## Average expression
avg_expr <- AverageExpression(major.minor.sct.integrated.clean.visual,
                              group.by = "CellTypes",
                              assays = "BothSCT") 

avg_expr$BothSCT %>% as.data.frame() %>%
  rownames_to_column(var = "scGeneName") %>%
  pivot_longer(names_to = "CellTypes", values_to = "NormExpression", -scGeneName) ->
  avg_expr.BothSCT.long


 avg_expr.BothSCT.long %>%
  inner_join(Celltypes.info) %>%
   ungroup() %>%
  group_by(scGeneName,glandZone) %>%
  summarise(NormExpression = mean(NormExpression)) %>%
  inner_join(geneInfo) %>%
  select(geneNameProtein,geneName, glandZone,NormExpression) ->
avg_expr.BothSCT.long   
 
 


```

### Common genes

```{r}
order =c("lman2l","x975_2",
          "SpiCE-LMa3","SpiCE-LMa6","AmSp-like1",
         "AmSp-like2","LASC009784","SpiCE-LMa2","SpiCE-LMa4",
         "spa","LASC007099","LASC017532","mttp_2","SpiCE-LMa1")     

 finalList.genes.common.fig7  %>%
   filter(geneNameProtein %in% order) %>%
   mutate(geneNameProtein = factor(geneNameProtein,
                                   levels= order)) %>%
   arrange(geneNameProtein)->
   common.sct.sp.prot.sig.genes
 
 
 geneInfo %>%
   select(geneNameProtein,scGeneName,geneName,lascID) %>%
   inner_join(common.sct.sp.prot.sig.genes) ->
   common.sct.sp.prot.sig.genes

 
   
```




#### Complex Heatmap
##### scRNA
```{r} 




avg_expr.BothSCT.long %>% ungroup() %>%
  filter(scGeneName %in% common.sct.sp.prot.sig.genes$scGeneName)%>%
  filter(glandZone %in% c("Minor_ZoneA","Minor_ZoneB","Minor_ZoneC",
                          "Major_ZoneA","Major_ZoneB","Major_ZoneC")) %>%
  select(geneNameProtein,glandZone , NormExpression)->
  common.scRNA.expression
  

common.scRNA.expression %>%
  filter(NormExpression > 1e-04) %>%
  pivot_wider(names_from = "glandZone", values_from = "NormExpression",
              values_fill = 1e-04) %>%
  column_to_rownames(var = "geneNameProtein")->
  common

common[as.character(common.sct.sp.prot.sig.genes$geneNameProtein),] -> 
common

common.markerGenes = 
  markerGenes.all.wide[rownames(common),colnames(common) ]

scRNA.data.common = log2(common)
 
 

col_fun_scRNA <- colorRamp2(c(0, 3 ,8), c("blue", "white", "red")) # 


Heatmap(scRNA.data.common,
        col = col_fun_scRNA,
        na_col = NA,
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (common.markerGenes[i, j]==1) {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = "black",  lwd = 2), just = "center")
          }else{
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = NA,  lwd = 2), just = "center")
            
          }
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                    gp = gpar(fill = col_fun_scRNA(scRNA.data.common[i,j]), col = NA), just = "center") # Remove black outline
          #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "scRNA expression",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE,
        cluster_rows = FALSE) ->
  common.HM.scRNA
common.HM.scRNA


```
##### Proteomics data
```{r}


MQ_data.AmpGl.QC.summary.QC %>%
  inner_join(geneInfo) %>% 
  filter(geneNameProtein %in% common.sct.sp.prot.sig.genes$geneNameProtein) %>%
  arrange(geneName,Part, gland)  %>%
  mutate(percent = relativeIntensity*100) %>%
  select(geneNameProtein,Part,gland,percent ) %>%
  pivot_wider(names_from = c(gland, Part), 
              values_from = percent ,
              names_sep = "_") %>%
  column_to_rownames("geneNameProtein") ->
  MQ.data.common

minPercent = log2(min(MQ_data.AmpGl.QC.summary.QC$relativeIntensity)*100)
maxPercent = log2(max(MQ_data.AmpGl.QC.summary.QC$relativeIntensity)*100)


 MQ.data.common[as.character(common.sct.sp.prot.sig.genes$geneNameProtein),c("Minor_Gland","Major_Gland","Minor_Dope","Major_Dope","Major_Silk")]  ->
   MQ.data.common
 
 MQ.data.common = log2(MQ.data.common)

col_fun_Protein <- colorRamp2(c(-5,-3,5),
                      c("blue", "white", "red"))# 


MQ.data.common %>% select(Minor_Gland,Major_Gland) ->
  MQ.data.common.gland

Heatmap(MQ.data.common.gland,
        col = col_fun_Protein,
        na_col = "black",
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (!is.na(MQ.data.common[i, j])) {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                    gp = gpar(col = "black",  lwd = 2), just = "center")
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                              gp = gpar(fill = col_fun_Protein(MQ.data.common[i,j]),
                                        col = NA),
                    just = "center") # Remove black outline
 #Fill the cells with the color gradient
          }

        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "log2 percent",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE, cluster_rows = FALSE) ->
  common.HM.protein

common.HM.protein


 
```

##### BulkRNA data

```{r}
read_csv(here("MinorGland_paper","data","bulkRNA",
                      "DEseq2_tissue.csv.gz")) ->
  DEseq.gland.results



DEseq.gland.results %>%
  filter(scGeneName %in% common.sct.sp.prot.sig.genes$scGeneName ) %>%
  mutate(sigLevel = "None") %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.05, "***")) %>%
  mutate(dir = "Up") %>%
  mutate(dir = replace(dir, log2FoldChange < 0, "Down")) %>%
  select(geneNameProtein,Tissue,log2FoldChange,sigLevel,dir)->
  DEseq.common.results

DEseq.common.results


DEseq.gland.results%>%
  filter(scGeneName %in% common.sct.sp.prot.sig.genes$scGeneName ) %>%
  mutate(sigLevel = "None") %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.05, "***")) %>%
  mutate(dir = "Up") %>%
  mutate(dir = replace(dir, log2FoldChange < 0, "Down")) %>%
  select(geneNameProtein,Tissue,log2FoldChange,sigLevel,dir)->
  DEseq.common.results
```

```{r}

bulk.data.merged.QC.tpm = read_csv(here("MinorGland_paper","data","bulkRNA",
                                        "bulk.data.merged.QC.tpm.csv.gz"))

sampleInfo.QC.merged = read_csv(file = here("MinorGland_paper","data","bulkRNA","sampleInfo.QC.merged.csv.gz"))

bulk.data.merged.QC.tpm %>%
  inner_join(geneInfo) %>%
  filter(geneNameProtein %in% common.sct.sp.prot.sig.genes$geneNameProtein) %>%
  inner_join(sampleInfo.QC.merged)%>%
  filter(Tissue %in% c("Minor","Major","Head")) %>%
  group_by(geneNameProtein,Tissue) %>%
  summarise(mean = mean(tpm))  %>%
  pivot_wider(names_from = "Tissue", values_from = "mean") %>%
  column_to_rownames("geneNameProtein") %>%
  select(Head,Major,Minor) %>%
  as.matrix() -> mat_tissue.common

mat_tissue = log2(mat_tissue.common)
mat_tissue = mat_tissue[as.character(common.sct.sp.prot.sig.genes$geneNameProtein),
                        c("Minor","Major","Head")]


border_colors <- c("Up" = "blue", "Down" = "red","None"="black")



DEseq.common.results %>%
  select(geneNameProtein,Tissue,log2FoldChange) %>%
  pivot_wider(names_from = "Tissue", values_from = "log2FoldChange") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> border_data.common

border_data.common[as.character(common.sct.sp.prot.sig.genes$geneNameProtein),] ->
  border_data.common


DEseq.common.results %>%
  select(geneNameProtein,Tissue,sigLevel) %>%
  pivot_wider(names_from = "Tissue", values_from = "sigLevel") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> border.common

border.common[as.character(common.sct.sp.prot.sig.genes$geneNameProtein),] ->
  border.common


breaks <- seq(-10, 10, length.out = 10)

plasma_colors <- viridis::plasma(10)

col_fun_bulk_RNA <-  circlize::colorRamp2(breaks,plasma_colors) # Corrected color mapping

colnames(border_data.common) = c("Minor vs Major","Minor vs Head","Major vs Head")
Heatmap(border_data.common,
        col = col_fun_bulk_RNA,
        na_col = NA,
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (!(border.common[i, j] == "None")) {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = "black",  lwd = 2), just = "center")
          }else{
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = NA,  lwd = 2), just = "center")
          }
          
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                    gp = gpar(fill = col_fun_bulk_RNA(border_data.common[i,j]), col = NA), just = "center") # Remove black outline
          #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "log2fold change",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE,
        cluster_rows = FALSE) ->
  common.HM.RNAbulk

common.HM.RNAbulk
```

####  Figure 7C

```{r}
common.HM.scRNA+common.HM.RNAbulk+common.HM.protein -> HM.common.all

pdf(here("MinorGland_paper","figures","figure7","Figure7C.pdf"),
    width = 6, height= 5)
HM.common.all
dev.off()

HM.common.all

```


### Unique genes


##### scRNA data
```{r}
uniqueMarkers.sp.prot.sig.genes %>%
  select(scGeneName) %>%inner_join(geneInfo) %>%
  inner_join(avg_expr.BothSCT.long) %>%
  select(scGeneName,geneNameProtein, glandZone,NormExpression) ->
  unique.scRNA.expression

unique.scRNA.expression %>%
  filter(NormExpression > 1e-04) %>%
  pivot_wider(names_from = "glandZone", values_from = "NormExpression",
              values_fill = 1e-04) %>%
  column_to_rownames(var = "geneNameProtein")%>%
  select(-scGeneName)->
  unique

unique = unique[go$geneOrder,glandZoneOrder$glandZone[1:6]]

unique

scRNA.data = log2(unique)

unique.markerGenes = 
  markerGenes.all.wide[rownames(scRNA.data),colnames(scRNA.data) ]


min_val <- min(scRNA.data)
max_val <- max(scRNA.data)



Heatmap(scRNA.data,
        col = col_fun_scRNA,
        na_col = NA,
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (unique.markerGenes[i, j] ==1) {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = "black",  lwd = 2), just = "center")
          }else{
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = NA,  lwd = 2), just = "center")
          }
          
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                    gp = gpar(fill = col_fun_scRNA(scRNA.data[i,j]), col = NA), just = "center") # Remove black outline
          #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "scRNA expression",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE,
        cluster_rows = FALSE) ->
  unique.HM.scRNA
unique.HM.scRNA




```

##### Proteomics data
```{r}


MQ_data.AmpGl.QC.summary.QC %>%
  inner_join(geneInfo) %>% 
  filter(geneNameProtein %in% uniqueMarkers.sp.prot.sig.genes$geneOrder) %>%
  arrange(geneName,Part, gland)  %>%
  mutate(percent = relativeIntensity*100) %>%
  select(geneNameProtein,Part,gland,percent ) %>%
  pivot_wider(names_from = c(gland, Part), 
              values_from = percent ,
              names_sep = "_") %>%
  column_to_rownames("geneNameProtein") ->
  MQ.data


 MQ.data[as.character(uniqueMarkers.sp.prot.sig.genes$geneOrder),c("Minor_Gland","Major_Gland","Minor_Dope","Major_Dope","Minor_Silk","Major_Silk")]  ->
   MQ.data
 
 MQ.data = log2(MQ.data)

min_val <- min(MQ.data)
max_val <- max(MQ.data)

MQ.data %>% select(Minor_Gland,Major_Gland) ->
  MQ.data.unique.gland

Heatmap(MQ.data.unique.gland,
        col = col_fun_Protein,
        na_col = "black",
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (!is.na(MQ.data[i, j])) {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                    gp = gpar(col = "black",  lwd = 2), just = "center")
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                              gp = gpar(fill = col_fun_Protein(MQ.data[i,j]),
                                        col = NA),
                    just = "center") # Remove black outline
 #Fill the cells with the color gradient
          }

        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "Protein expression",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE, cluster_rows = FALSE) ->
  unique.HM.protein


 
```


##### BulkRNA data

```{r}

DEseq.gland.results %>%
  filter(geneNameProtein %in%
           as.character(uniqueMarkers.sp.prot.sig.genes$geneOrder) ) %>%
  mutate(sigLevel = "None") %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.05, "***")) %>%
  mutate(dir = "Up") %>%
  mutate(dir = replace(dir, log2FoldChange < 0, "Down")) %>%
  select(geneNameProtein,Tissue,log2FoldChange,sigLevel,dir)->
  DEseq.unique.results


```


```{r}


bulk.data.merged.QC.tpm %>%
  inner_join(geneInfo) %>%
  filter(geneNameProtein %in% uniqueMarkers.sp.prot.sig.genes$geneOrder) %>%
  inner_join(sampleInfo.QC.merged)%>%
  filter(Tissue %in% c("Minor","Major","Head")) %>%
  group_by(geneNameProtein,Tissue) %>%
  summarise(mean = mean(tpm))  %>%
  pivot_wider(names_from = "Tissue", values_from = "mean") %>%
  column_to_rownames("geneNameProtein") %>%
  select(Head,Major,Minor) %>%
  as.matrix() -> mat_tissue

mat_tissue = log10(mat_tissue)
mat_tissue = mat_tissue[as.character(uniqueMarkers.sp.prot.sig.genes$geneOrder),
                        c("Minor","Major","Head")]


border_colors <- c("Up" = "blue", "Down" = "red","None"="black")



DEseq.unique.results %>%
  select(geneNameProtein,Tissue,log2FoldChange) %>%
  pivot_wider(names_from = "Tissue", values_from = "log2FoldChange") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> border_data.l2c

dim(border_data.l2c)
as.character(uniqueMarkers.sp.prot.sig.genes$geneOrder)

border_data.l2c[as.character(uniqueMarkers.sp.prot.sig.genes$geneOrder),] ->
  border_data.l2c

DEseq.unique.results%>%
  select(geneNameProtein,Tissue,sigLevel) %>%
  pivot_wider(names_from = "Tissue", values_from = "sigLevel") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> border_data.sl

border_data.sl[as.character(uniqueMarkers.sp.prot.sig.genes$geneOrder),] ->
  border_data.sl

colnames(border_data.l2c) = c("Minor vs Major","Minor vs Head","Major vs Head")


Heatmap(border_data.l2c,
        col = col_fun_bulk_RNA,
        na_col = NA,
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (border_data.sl[i,j] != "None") {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = "black",  lwd = 2), just = "center")
          }else{
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                      gp = gpar(col = NA,  lwd = 2), just = "center")
            
          }
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                    gp = gpar(fill = col_fun_bulk_RNA(border_data.l2c[i,j]), col = NA), just = "center") # Remove black outline
          #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "log2Fold change",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE,
        cluster_rows = FALSE) ->
  unique.HM.RNAbulk
```
####  Figure 7D

```{r}
unique.HM.scRNA+unique.HM.RNAbulk+unique.HM.protein -> HM_all.unique

pdf(here("MinorGland_paper","figures","figure7","Figure7D.pdf"),
    width = 6, height= 5)
HM_all.unique
dev.off()


HM_all.unique

```


## Figure 8
```{r}


# Proteomics data
proteomics.wide.summary = read_csv(file = here("MinorGland_paper","data","proteomics","proteomics.wide.summary.csv.gz"))%>% 
  filter(Part == "Gland")


#
finalList.genes %>% filter(bulk == "Yes") %>%
  filter(common == "Yes" ) %>%
  arrange(common, Zone)->
  finalList.genes.common.fig7
finalList.genes.common.fig7

finalList.genes %>% 
  filter(!is.na(unique) ) %>%
  arrange( Zone,unique)->
  finalList.genes.unique.fig7

proteomics.wide.summary %>% 
  filter(geneNameProtein %in% finalList.genes.common.fig7$geneNameProtein) %>%
  mutate(specificity = "Common") ->
  proteomics.wide.summary.common

finalList.genes.unique.fig7 %>%
  select(geneNameProtein,unique) %>%
  rename(specificity = "unique") %>%
  inner_join(proteomics.wide.summary) ->
  proteomics.wide.summary.unique

bind_rows(proteomics.wide.summary.common,
              proteomics.wide.summary.unique) ->
  proteomics.wide.summary.finalList




  
  
proteomics.wide.summary.finalList %>%
  ggplot(aes(x = Mi*100, y = Ma*100, fill = specificity,
              label = geneNameProtein)) +
  # Add shaded 2-fold band
  geom_line(
    data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
      mutate(y = 0.1 * x),
    aes(x = x, y = y),
    inherit.aes = FALSE,
    linetype = "dotted",
    color = "grey40"
  ) +
  
  geom_line(
    data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
      mutate(y = 10 * x),
    aes(x = x, y = y),
    inherit.aes = FALSE,
    linetype = "dotted",
    color = "grey40"
  )  +
  geom_ribbon(data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
                     mutate(ymin = 0.5 * x, ymax = 2 * x),
                   aes(x = x, ymin = ymin, ymax = ymax),
                   inherit.aes = FALSE,
                   fill = "grey50", alpha = 0.3) +  

  geom_point(position = "jitter", shape = 22) +  
 geom_label_repel(aes(fill = specificity), color="black",
  label.r = unit(0, "lines"),           # square corners
  label.padding = unit(0.1, "lines"),   # tighter label
  box.padding = unit(0.15, "lines"),    # closer to points
  max.overlaps = 20,
  size = 3,
  label.size = 0.3)+# varwidth adjusts width by sample size
  scale_y_continuous(trans = "log10", limits = c(0.00001, 10)) +
  scale_x_continuous(trans = "log10", limits = c(0.00001, 10)) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    x = "Relative intensity (%) (log10)",
    y = "Relative intensity (%) (log10)",
    fill = "Gland"
  )+ 
  scale_color_manual(values = c(Minor = "#E2D799",Major = "#E2D799",Common ="#CDDCFB"))+
  scale_fill_manual(values = c(Minor = "#E2D799",Major = "#E2D799",Common ="#CDDCFB"))->
  Figure8.next 

Figure8.next
ggsave(here("MinorGland_paper","figures","figure8","Figure8.pdf"),
       plot = Figure8.next,
    width = 6, height= 5)


```

