---
title: "Bulk RNA analysis"
author: "Johan"
date: "05/26/2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    df_print: paged

params:
  RfunctionsDirectory: /Users/johanreimegard/git/RNAmappingPipeline/R
  
  
---






```{r setup, include=FALSE}

# if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")#

# BiocManager::install("ropls")

source("~/git/RNAmappingPipeline/R/ExpressionAnalysisFunctions.R")
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

library(knitr)
#install.packages("kableExtra")
#library("viridis")     

library(RColorBrewer)
#library(GenomicFeatures)
library(DESeq2)

library(ropls)

library(ggforce)
library(patchwork)


rename <- dplyr::rename


source("../loadGeneInfo.R")
source("../colorSchemes.R")
summarise = dplyr::summarise               
select = dplyr::select
group_by = dplyr::group_by
```


## Functions


```{r , message=FALSE, warning=FALSE}


getVSTexpression = function(count.df, 
                            geneColumn = "gene_id",
                            sampleColumn = "sample",
                            countColumn = "readCount",
                            md ,
                            gi, 
                            design = "Tissue"){
  

  count.df %>%
    dplyr::select(all_of(geneColumn), all_of(sampleColumn),all_of(countColumn)) %>%
    pivot_wider(names_from = all_of(sampleColumn),
                values_from = all_of(countColumn) ) %>%
    as.data.frame() %>%
    column_to_rownames(var = "gene_id")->
    CM

  
  CM = CM[gi[[all_of(geneColumn)]],
          md[[ all_of(sampleColumn)]]
  ]
  md$design = as.factor(md[[all_of(design)]])
  setdiff(gi[[all_of(geneColumn)]],rownames(CM))
  setdiff(md[[ all_of(sampleColumn)]],colnames(CM))

  dds.tissue <- DESeqDataSetFromMatrix(countData = CM,
                                       colData = md,
                                       design = ~design)
  
  
  vst <- vst(dds.tissue)
  
  normExpression = as.data.frame(assay(vst))
  return(normExpression)
}




getNormalisedExpression = function(count.df, 
                            geneColumn = "gene_id",
                            sampleColumn = "sample",
                            countColumn = "readCount",
                            md ,
                            gi, 
                            design = "Tissue"){
  

  count.df %>%
    dplyr::select(all_of(geneColumn), all_of(sampleColumn),all_of(countColumn)) %>%
    pivot_wider(names_from = all_of(sampleColumn),
                values_from = all_of(countColumn) ) %>%
    as.data.frame() %>%
    column_to_rownames(var = "gene_id")->
    CM

  
  CM = CM[gi[[all_of(geneColumn)]],
          md[[ all_of(sampleColumn)]]
  ]
  
  md$design = as.factor(md[[all_of(design)]])
  
  dds.tissue <- DESeqDataSetFromMatrix(countData = CM,
                                       colData = md,
                                       design = ~design)
  dds.tissue <- DESeq(dds.tissue)
  
  
  normCounts <- DESeq2::counts(dds.tissue, normalized=TRUE)
  
  normExpression = as.data.frame(assay(normCounts))
  return(normExpression)
}





getDEseq2Object = function(count.df, 
                            geneColumn = "gene_id",
                            sampleColumn = "sample",
                            countColumn = "readCount",
                            md ,
                            gi, 
                            design = "Tissue"){
  

  count.df %>%
    dplyr::select(all_of(geneColumn), all_of(sampleColumn),all_of(countColumn)) %>%
    pivot_wider(names_from = all_of(sampleColumn),
                values_from = all_of(countColumn) ) %>%
    as.data.frame() %>%
    column_to_rownames(var = "gene_id")->
    CM

  
  CM = CM[gi[[all_of(geneColumn)]],
          md[[ all_of(sampleColumn)]]
  ]
  
  md$design = as.factor(md[[all_of(design)]])
  
  dds.tissue <- DESeqDataSetFromMatrix(countData = CM,
                                       colData = md,
                                       design = ~design)
  
  
  return(dds.tissue)
}




getOPLS = function(normExpression, 
                   nrOfPredictive = 1,
                   nrOfOrtho = NA,
                   md,
                   gi, 
                   design = "Tissue"){

Parts.pls <- opls(t(normExpression),md[[design]] ,
                        predI = nrOfPredictive, orthoI = nrOfOrtho)
}

getPLS = function(normExpression, 
                   md,
                   gi, 
                   design = "Tissue"){

  Parts.pls <- opls(t(normExpression),md[[design]])
}




getMetaDataFromOPLS <- function(OPLS =  OPLS.parts,
                    md =sampleInfo.QC.merged.parts,
                    extension,sampleColumn = "sample")  {
  
  
  OPLS@scoreMN %>% as.data.frame() ->md.opls
  cn = paste(extension, colnames(md.opls), sep = "_")
  colnames(md.opls) = cn
  md.opls %>%
    rownames_to_column(var = all_of(sampleColumn)) ->
    md.opls 
  md = md %>% left_join(md.opls)
  return(md)
}  


getGeneInfoFromOPLS <- function(OPLS,md ,gi, 
                                extension ,
                                sampleColumn = "sample",
                                geneColumn = "gene_id", 
                                design = "Tissue")  {
  
  
  OPLS@scoreMN %>% as.data.frame() ->md.opls
  md.opls %>%
    rownames_to_column(var = all_of(sampleColumn)) ->
    md.opls 
  md %>% left_join(md.opls)%>%
    rename(design = all_of(design)) %>%
    group_by(design) %>%
    summarise(mean = mean(p1)) %>%
    mutate(dir = "neg" ) %>%
    mutate(dir = replace(x = dir, mean >=0 , values = "pos") ) %>%
    select(design,dir)  ->
    md.summary 
  
  OPLS@loadingMN %>% as.data.frame() -> loading
  
  
  loading %>% rownames_to_column(var = all_of(geneColumn)) %>%
    mutate(dir = "neg" ) %>%
    mutate(dir = replace(x = dir, p1 >=0 , values = "pos") ) %>%
    inner_join(md.summary) %>%
    select(all_of(geneColumn),design,p1) ->
    loading
  
  colnames(loading)[3] = paste(extension,colnames(loading)[3], sep = "_" )
  colnames(loading)[2] = all_of(extension)
  
  gi %>% left_join(loading) -> gi
  
  
  
  return(gi)
}

getGeneInfoFrom2DPLS <- function(OPLS,md ,gi, 
                                extension ,
                                sampleColumn = "sample",
                                geneColumn = "gene_id", 
                                design = "Tissue")  {
  
  
  OPLS@scoreMN %>% as.data.frame() ->md.opls
  md.opls %>%
    rownames_to_column(var = all_of(sampleColumn)) ->
    md.opls 
  md %>% left_join(md.opls)%>%
    rename(design = all_of(design)) %>%
    group_by(design) %>%
    summarise(mean_p1 = mean(p1),mean_p2 = mean(p2)) %>%
    mutate(dir_p1 = "neg" ) %>%
    mutate(dir_p1 = replace(x = dir_p1, mean_p1 >=0 , values = "pos") ) %>%
    mutate(dir_p2 = "neg" ) %>%
    mutate(dir_p2 = replace(x = dir_p2, mean_p2 >=0 , values = "pos") ) %>%
    select(design,dir_p1,dir_p2)  ->
    md.summary 
  
  OPLS@loadingMN %>% as.data.frame() -> loading
  
  
  loading %>% rownames_to_column(var = all_of(geneColumn)) %>%
    mutate(dir_p1 = "neg" ) %>%
    mutate(dir_p1 = replace(x = dir_p1, p1 >=0 , values = "pos") ) %>%
    mutate(dir_p2 = "neg" ) %>%
    mutate(dir_p2 = replace(x = dir_p2, p2 >=0 , values = "pos") ) %>%
    left_join(md.summary) %>%
    select(all_of(geneColumn),dir_p1,dir_p2,design,p1,p2) ->
    loading
  

  
  colnames(loading)[5] = paste(extension,colnames(loading)[5], sep = "_" )
  colnames(loading)[6] = paste(extension,colnames(loading)[6], sep = "_" )
  colnames(loading)[4] = all_of(extension)
  
  return(loading)
}


  



```



## Parameters
```{r, message=FALSE, warning=FALSE}

minExpressionCutoff =20
DataGenerated = TRUE

```
## Reference
```{r}
geneInfo = getGeneInfoFull() 
majorProteins = read_csv2("~/Box/bridgeSpider/annotation/silk.proteins.1.1.csv") %>%
  select(geneName,geneNameProtein,mean) %>% rename(masp_concentration = "mean")%>% filter(masp_concentration>0.015)

geneInfo = geneInfo %>% left_join(majorProteins)
transcriptInfo = addTranscriptInfo(geneInfo) %>%
  select(gene_id,geneName, tx_name) %>% left_join(majorProteins) 


geneInfo %>% filter(!is.na(geneNameProtein))


```



## Proteomics

#### Minor

```{r}

lines =read_lines("../../data/proteomics/minorSilk/F058708_MI-silk_1.csv")[c(68,seq(from = 69, to = 113, by = 2))] 
write_lines(lines,"../../data/proteomics/minorSilk/F058708_MI-silk_1.1.csv")

lines =read_lines("../../data/proteomics/minorSilk/F058710_MI-silk_2.csv")[c(68,seq(from = 69, to = 117, by = 2))] 
write_lines(lines,"../../data/proteomics/minorSilk/F058710_MI-silk_2.1.csv")

proteinData1 = read_csv("../../data/proteomics/minorSilk/F058710_MI-silk_2.1.csv")  %>%
  rename(tx_name = prot_acc) %>% inner_join(transcriptInfo) %>% arrange(-prot_score)  %>%
  select(geneName,geneNameProtein,prot_score,prot_matches_sig )
  
proteinData2 = read_csv("../../data/proteomics/minorSilk/F058708_MI-silk_1.1.csv") %>%
  rename(tx_name = prot_acc) %>% inner_join(transcriptInfo) %>% arrange(-prot_score)  %>%
  select(geneName,geneNameProtein,prot_score,prot_matches_sig )
  

proteinData1
proteinData2

bind_rows(proteinData1,proteinData2) %>% group_by(geneName,geneNameProtein) %>%
  summarise(prot_matches_sig = sum(prot_matches_sig), n= n()) %>% arrange(-prot_matches_sig) ->
proteinData.minor


proteinData.minor %>% inner_join(getGeneInfoFull()) %>%
  select(c(1:5,11)) -> 
  proteinData.minor

write_csv(proteinData.minor,"~/Box/bridgeSpider/proteomics/minor/minor.MS.summary.csv")

```


#### Major
```{r Loading the data}
# Retrieving the data
proteomicsData = read_tsv("../../data/proteomics/majorSilk/proteomics.tsv")
proteomicsData%>% dplyr::select(2,3,5,6,9,14,15) %>% unique() -> proteomicsData.unique 
colnames(proteomicsData.unique) = c("Sample","Replicate","gene_id","transcript_id","MolecularWeight","Percentage_Total_Spectra","Percentage_sequence_coverage")


proteomicsData.unique %>%
  mutate(Percentage_Total_Spectra = gsub("%","",Percentage_Total_Spectra)) %>%
  mutate(Percentage_Total_Spectra = gsub(",",".",Percentage_Total_Spectra)) %>%
  mutate(Percentage_Total_Spectra = as.numeric(Percentage_Total_Spectra)) ->
  proteomicsData.unique
  




```


###### Found in the gland 
```{r}
GI = getGeneInfoFull()
geneInfo.transcript = addTranscriptInfo(geneInfo = GI)

proteomicsData.unique %>% 
  filter(Sample == "Gland") %>%
  group_by(transcript_id ) %>%
  summarise( n = n(), mean = mean(Percentage_Total_Spectra))  %>%
  rename(tx_name = "transcript_id") %>%
  inner_join(geneInfo.transcript) %>%
  select(geneName, mean,n  )->
  proteomicsData.unique.gland


proteomicsData.unique.gland %>%
  rename(glandMean = "mean") %>%
  arrange(-glandMean) ->proteomicsData.unique.gland.order

proteomicsData.unique.gland.order$rank = 1:2145





proteomicsData.unique.gland.order$rank = 1:2145
proteomicsData.unique.gland.order %>%
 mutate(total =  sum(glandMean)) %>% 
  mutate(percentGland = glandMean /total * 100) %>%
  select(geneName, n , percentGland, rank) %>%
  rename(samplesGland = "n") %>%
  rename(rankGland = rank) ->
proteomicsData.unique.gland.order




```

###### Found in the Silk 
```{r}

proteomicsData.unique %>% 
  filter(Sample != "Gland") %>%
  filter(Sample != "Dope") %>%
  group_by(transcript_id, Sample ) %>%
  summarise( n = n(), mean = mean(Percentage_Total_Spectra))  %>%
  rename(tx_name = "transcript_id") %>%
  inner_join(geneInfo.transcript) %>%
  as.data.frame() %>%
  select(geneName, mean ,n) ->
  proteomicsData.unique.silk


nrow(proteomicsData.unique.silk)

proteomicsData.unique.silk %>%
  rename(silkMean = "mean") %>%
  arrange(-silkMean) ->proteomicsData.unique.silk.order

proteomicsData.unique.silk.order$rank = 1:nrow(proteomicsData.unique.silk.order)
proteomicsData.unique.silk.order %>%
 mutate(total =  sum(silkMean)) %>% 
  mutate(percentSilk = silkMean /total * 100) %>%
  select(geneName, n , percentSilk, rank) %>%
  rename(samplesSilk = "n") %>%
  rename(rankSilk = rank) ->
proteomicsData.unique.silk.order

```




###### Final set of proteins in the silk 
```{r}

 signalnPeptideInfo =  getSignalPeptideInfo(gitDir = "~/git/spidersilk", geneInfo.transcript = geneInfo.transcript)

proteomicsData.unique.silk %>%
  filter(geneName %in% proteomicsData.unique.gland$geneName) %>%
  filter(n > 4)  %>%
  left_join(signalnPeptideInfo)->
  proteomicsData.unique.silk.gland


proteomicsData.unique.silk.gland

proteomicsData.unique.silk.gland %>%
  arrange(-mean) %>%
  filter(Term == "signalP") ->
  proteomicsData.unique.silk.gland.signalP


proteomicsData.unique.silk.gland.signalP %>% filter(Term == "signalP")


```

## BulkRNA


The information about mapping can be found in the Map2Reference.Rmd

#### Samples 
##### Minor
```{r}
 order =c("ZoneA_Proximal","ZoneA_Distal","ZoneB","ZoneC","ZoneC_Duct") 
 
library(tidyverse)
read_csv("~/git/spidersilk/data/bulkRNA/minorParts/samples.txt") %>%
  mutate(Tissue= "Minor") %>%
  mutate(part = gsub("DMI","ZoneA_Distal",part)) %>%
  mutate(part = gsub("PMI","ZoneA_Proximal",part)) %>%
  mutate(part = gsub("BMI","ZoneB",part)) %>%
  mutate(part = gsub("CMI","ZoneC",part)) %>%
  mutate(part = gsub("CDMi","ZoneC_Duct",part)) %>%
  mutate(part = factor(part, levels =  order))%>%
  rename(Part = "part") %>%
  mutate(sampleID2 = paste(Part,rep, sep = "_")) -> 
  samplesInfo.minor



fc = read_tsv("~/git/spidersilk/data/bulkRNA/minorParts/minorParts.featureCount.reverse.txt", skip = 1) %>%
  select(1,7:26)  %>%
  rename(gene_id = Geneid) %>%
  pivot_longer(cols = -gene_id, names_to = "sample", values_to = "readCount") %>%
  mutate(sampleID = gsub("/Aligned.sortedByCoord.out.bam", "",sample))  %>%
  inner_join(samplesInfo.minor)

fc %>% select(gene_id,sampleID2,Tissue,Part,readCount) %>% rename(sample = "sampleID2") ->
  fc.minor





```

##### Rest
```{r}

sampleInfo.QC.merged.rest = read_csv(file = "~/Box/bridgeSpider/bulkRNA/sampleInfo.csv")

# Fix parts
sampleInfo.QC.merged.rest %>%
  mutate(Part = "Whole") %>%
  mutate(Part = replace(x = Part,list = Tissue == "major-duct", values = "Duct"))%>%
  mutate(Part = replace(x = Part,list = Tissue == "major-sac", values = "Sac"))%>%
  mutate(Part = replace(x = Part,list = Tissue == "major-tail", values = "Tail"))->
  sampleInfo.QC.merged.rest

# Fix samples names so that they correspond to the featurecount file 

sampleInfo.QC.merged.rest %>%
  mutate(sample = gsub(pattern = "wholespider",replacement = "wholebody", x = sample)) %>%
  mutate(sample = gsub(pattern = "AciniPyri",replacement = "acinipyri", x = sample)) ->
  sampleInfo.QC.merged.rest


# Fix Tissues with capital letters and remove parts information
 sampleInfo.QC.merged.rest %>%
   mutate(Tissue = gsub(pattern = "aggregate",replacement = "Aggregate",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "flagelliform",replacement = "Flagelliform",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "major-duct",replacement = "Major",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "major-sac",replacement = "Major",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "major-tail",replacement = "Major",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "majorampullate",replacement = "Major",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "minor",replacement = "Minor",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "tubuli",replacement = "Tubuli",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "head",replacement = "Head",x = Tissue)) %>%
  mutate(Tissue = gsub(pattern = "wholespider",replacement = "WholeBody",x = Tissue)) ->
  sampleInfo.QC.merged.rest

bulk.data.merged = read_csv(file = "../../data/bulkRNA/All.samples.count.csv") %>%
  rename(gene_id = "Geneid") 

fc.rest = bulk.data.merged %>%
  pivot_longer( cols = -gene_id, names_to = "sample", values_to = "readCount") %>%
  inner_join(sampleInfo.QC.merged.rest)%>% select(gene_id,sample,Tissue,Part,readCount)


```
#### Join all
##### Feature count table
```{r}

fc.all = bind_rows(fc.rest,fc.minor)  %>% filter(Tissue != "Tubuli")
  

fc.all %>% distinct(sample)
fc.all %>% distinct(sample,Tissue,Part) %>% group_by(Tissue,Part) %>%
  summarise(n = n()) 
fc.all %>% distinct(sample,Tissue,Part) %>% group_by(Tissue) %>%
  summarise(n = n()) 


```
##### Sample info
```{r}
fc.all %>%
  distinct(sample,Tissue,Part) ->
  sampleInfo.all



sampleInfo.all  %>%
  mutate(sample2 = sample) %>%
  mutate(sample2 = gsub(pattern = "acinipyri",replacement = "AciniPyri", x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "aggregate",replacement = "Aggregate",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "flagelliform",replacement = "Flagelliform",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "major-duct",replacement = "Duct",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "major-sac",replacement = "Sac",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "major-tail",replacement = "Tail",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "majorampullate",replacement = "Major",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "minor",replacement = "Minor",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "tubuli",replacement = "Tubuli",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "wholebody",replacement = "WholeBody",x = sample2)) %>%
  mutate(sample2 = gsub(pattern = "head",replacement = "Head",x = sample2)) ->
  sampleInfo.all
  
  sampleInfo.all %>%
    mutate(Gland = "YES") %>%
    mutate(Gland = replace(x = Gland,list = Tissue == "Head", values = "NO")) %>%
    mutate(Gland = replace(x = Gland,list = Tissue == "WholeBody", values = "NO")) ->
    sampleInfo.all 
  

 sampleInfo.all 
 
 sampleInfo.all %>% inner_join(fc.all) %>%
   select(gene_id,sample2,readCount, Tissue, Part,Gland)
 
```


```{r }
  fc.all %>% distinct(gene_id) -> geneInfo.all

  
```



# RNA analysis

# Gene QC filter


```{r}

# Get counts per gene as more stable than read count
fc.all %>% 
  group_by(sample) %>%
  dplyr::summarise(sampleReads = sum(readCount)) %>%
  inner_join(fc.all) %>%
  mutate(rpm = readCount / sampleReads *1000000) %>%
  group_by(Tissue,Part,gene_id) %>%
  summarise(max = max(rpm), n = n(), mean = mean(rpm)) %>%
  mutate(log = log2(mean+1)) ->
  fc.all.rpm.summary


#Visualise
  fc.all.rpm.summary %>%
  ggplot(aes(x =log, color = Tissue)) + geom_density()+ geom_vline(xintercept = 2)


#Remove lowly expressed genes
  fc.all.rpm.summary %>%
    mutate(log = log2(mean+1))%>% 
    filter(log > 1)  %>%
    as.data.frame()%>%
    distinct(gene_id)%>% 
    inner_join(geneInfo.all) ->
    geneInfo.all.qc
  
# Remove from reads 
fc.all %>% inner_join(geneInfo.all.qc) ->
  fc.all.qc


fc.all.qc %>%
  summarise(genes = length(unique(gene_id)), samples = length(unique(sample)))


fc.all.rpm.summary %>%
  select(Tissue,Part, gene_id, mean) %>%
  rename(rpm = "mean") ->
  geneinfo.rpm


geneinfo.rpm %>%
  group_by( gene_id) %>%
  summarise(mean = mean(rpm), sd = sd(rpm)) %>%
  inner_join(geneinfo.rpm) %>%
  mutate(zscore = (rpm-mean)/sd) %>%
  filter(mean>0) ->
  geneinfo.rpm

geneinfo.rpm %>% inner_join(geneInfo) %>% filter(geneName == "MiSpa") %>% arrange(-rpm)
geneinfo.rpm %>% inner_join(geneInfo) %>% filter(geneName == "MiSpc") %>% arrange(-rpm)
geneinfo.rpm %>% inner_join(geneInfo) %>% filter(geneName == "MaSp1a") %>% arrange(-rpm)
geneinfo.rpm %>% inner_join(geneInfo) %>% filter(geneName == "MaSp2c") %>% arrange(-rpm)



```


# OPLS part analysis on glands
can be found in the OPLS analysis in bulkRNA

### Head vs Body


```{r whole spider fragments5, message=FALSE, warning=FALSE}

# SampleInfo
sampleInfo.all %>%
  filter(Tissue %in%  c("Head","WholeBody")) ->
  sampleInfo.all.tissues

# geneInfo
geneInfo.all.qc ->
  geneInfo.tissues


  

normExpression.parts = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.all.tissues,
                 gi =geneInfo.tissues, 
                 design =  "Tissue" )

OPLS.parts = getOPLS(normExpression = normExpression.parts,
                 md =sampleInfo.all.tissues,
                 gi =geneInfo.tissues, 
                 design =  "Tissue"
          )


sampleInfo.tissues = getMetaDataFromOPLS(OPLS =  OPLS.parts, 
                    md = sampleInfo.all.tissues,
                    extension =  "Tissue" ) 

geneInfo.tissues = getGeneInfoFromOPLS(OPLS =  OPLS.parts, 
                    md = sampleInfo.all.tissues,
                    extension =  "Tissue",gi = geneInfo.tissues )


geneInfo.all.qc %>% left_join(geneInfo.tissues) -> geneInfo.tissues
sampleInfo.all %>% left_join(sampleInfo.tissues) -> sampleInfo.tissues



```


### Body vs glands

```{r whole spider fragments6, message=FALSE, warning=FALSE}

sampleInfo.tissues %>%
  filter(Tissue %in% c("Major","Minor","Aggregate",
                       "AciniPyri","Flagelliform"
                       ,"WholeBody")) %>% 
  filter(Part =="Whole")->
  sampleInfo.gland



geneInfo.tissues  ->
  geneInfo.gland
  

normExpression.glands = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.gland,
                 gi =geneInfo.gland, 
                 design="Tissue")


OPLS.glands = getOPLS(normExpression = normExpression.glands,
                     md =sampleInfo.gland,
                     gi =geneInfo.gland, 
                     design =  "Gland")


sampleInfo.glands = getMetaDataFromOPLS(OPLS =  OPLS.glands, 
                    md = sampleInfo.gland,
                    extension =  "Gland" )  %>%
  select(sample,Gland,Gland_p1 )

sampleInfo.tissues %>% left_join(sampleInfo.glands) -> sampleInfo.tissues.glands



geneInfo.glands = getGeneInfoFromOPLS(OPLS =  OPLS.glands,
                                      design = "Gland", 
                                      md = sampleInfo.gland,
                                      extension =  "Gland",
                                      gi = geneInfo.gland ) %>%
  select(gene_id,Gland,Gland_p1)

geneInfo.tissues %>% left_join(geneInfo.glands) ->geneInfo.tissues.glands
```

```{r}


geneInfo.tissues.glands %>%
  filter(!is.na(Gland_p1)) %>%
  ggplot(aes(x = Gland_p1, color = Gland)) + 
  geom_density()-> geneInfo.tissues.glands.plot

sampleInfo.tissues.glands %>%
  filter(!is.na(Gland_p1)) %>% 
  ggplot(aes(x = Gland_p1, color = Gland)) + 
  geom_density()-> 
  sampleInfo.tissues.glands.plot



patchwork <- sampleInfo.tissues.glands.plot / geneInfo.tissues.glands.plot
patchwork + plot_annotation(
  title = 'Figure body parts',
  subtitle = 'Comparing glands and whole body '
)

ggsave("~/Box/bridgeSpider/bulkRNA/figures/OPLS.body_vs_glands.pdf")
```




### Ampullate glands vs rest of the glands

```{r whole spider fragments7, message=FALSE, warning=FALSE}

sampleInfo.tissues.glands %>%
  filter(Tissue %in% c("Major","Minor","Aggregate",
                       "AciniPyri","Flagelliform")) %>% 
  filter(Part == "Whole") %>% 
  mutate(Ampullate = "NO") %>%
  mutate(Ampullate = replace(x = Ampullate,list = Tissue %in% c("Major","Minor"), values = "YES"))->
  sampleInfo.tissues.glands.ampullate


# geneInfo
geneInfo.tissues ->
  geneInfo.tissues.glands.ampullates


normExpression.major = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.tissues.glands.ampullate,
                 gi =geneInfo.tissues.glands.ampullates, 
                 design =  "Ampullate" )

OPLS.major = getOPLS(normExpression = normExpression.major,
                     nrOfOrtho = 1,
                     nrOfPredictive =1,
                     md =sampleInfo.tissues.glands.ampullate,
                     gi =geneInfo.tissues.glands.ampullates, 
                     design =  "Ampullate")


sampleInfo.ampullate = getMetaDataFromOPLS(OPLS =  OPLS.major, 
                                                          md = sampleInfo.tissues.glands.ampullate,
                                                          extension =  "Ampullate" ) %>%
  select(sample,Ampullate,Ampullate_p1)
  

geneInfo.ampullate = getGeneInfoFromOPLS(OPLS =  OPLS.major, design = "Ampullate",
                    md = sampleInfo.tissues.glands.ampullate,
                    gi = geneInfo.tissues.glands.ampullates,
                    extension =  "Ampullate" ) %>%
  select(gene_id, Ampullate, Ampullate_p1)


geneInfo.tissues.glands %>% left_join(geneInfo.ampullate) -> geneInfo.tissues.glands.ampullate
sampleInfo.tissues.glands %>% left_join(sampleInfo.ampullate) -> sampleInfo.tissues.glands.ampullate

```

### Major vs Minor
```{r }

sampleInfo.tissues.glands.ampullate %>%
  filter(Tissue %in% c("Major","Minor")) %>% 
  filter(Part == "Whole") ->
  sampleInfo.tissues.glands.ampullate.separated


# geneInfo
geneInfo.tissues ->
  geneInfo.tissues.glands.ampullate.separated
 

normExpression.internal = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.tissues.glands.ampullate.separated,
                 gi =geneInfo.tissues.glands.ampullate.separated, 
                 design =  "Tissue" )

OPLS.internal = getOPLS(normExpression = normExpression.internal,
                     nrOfOrtho = 1,
                     nrOfPredictive =1,
                     md =sampleInfo.tissues.glands.ampullate.separated,
                     gi =geneInfo.tissues.glands.ampullate.separated, 
                     design =  "Tissue")


sampleInfo.separated = getMetaDataFromOPLS(OPLS =  OPLS.internal, 
                    md = sampleInfo.tissues.glands.ampullate.separated,
                    extension =  "Separated" ) %>%
  select(sample,Tissue,Separated_p1)
  

geneInfo.separated = getGeneInfoFromOPLS(OPLS =  OPLS.internal, design = "Tissue",
                    md = sampleInfo.tissues.glands.ampullate.separated,
                    gi = geneInfo.tissues.glands.ampullate.separated,
                    extension =  "Separated" ) %>%
  select(gene_id,Separated , Separated_p1)


geneInfo.tissues.glands.ampullate %>% left_join(geneInfo.separated) -> geneInfo.tissues.glands.ampullate.separated


```
### Sample and Gene info 

```{r}
sampleInfo.tissues.glands.ampullate.separated -> sampleInfo.OPLS
geneInfo.tissues.glands.ampullate.separated -> geneInfo.OPLS

geneInfo.tissues.glands.ampullate.separated %>%
  inner_join(getGeneInfoFull()) %>%
  filter(Group == "Spidroin") -> test

ggplot(test, aes(Ampullate_p1,Separated_p1, label = geneName)) + geom_point()+ geom_text()


```




## DEseq analysis

### Head vs Body
```{r }


sampleInfo.OPLS %>%
  filter(Tissue %in% c("Head","WholeBody")) ->
  sampleInfo.OPLS.tissues

# geneInfo
geneInfo.OPLS -> 
  geneInfo.OPLS.tissues


Tissue.deseq = getDEseq2Object(count.df = fc.all.qc,
                 md =sampleInfo.OPLS.tissues,
                 gi =geneInfo.OPLS.tissues, 
                 design =  "Tissue" )


Tissue.deseq.DE = DESeq(Tissue.deseq)

as.data.frame(results(Tissue.deseq.DE)) %>%
  rownames_to_column(var = "gene_id") %>%
  rename(Body_FC = "log2FoldChange") %>%
  rename(Body_padj =  "padj") %>%
  select(gene_id,Body_FC,Body_padj) -> result



geneInfo.OPLS %>% left_join(result) -> geneInfo.OPLS.tissue


```

### Body vs glands

```{r whole spider fragments6, message=FALSE, warning=FALSE}

sampleInfo.tissues %>%
  filter(Tissue %in% c("Major","Minor","Aggregate",
                       "AciniPyri","Flagelliform"
                       ,"WholeBody")) %>% 
  filter(Part =="Whole")->
  sampleInfo.gland

geneInfo.OPLS ->  geneInfo.gland
  
Gland.deseq = getDEseq2Object(count.df = fc.all.qc,
                 md =sampleInfo.gland,
                 gi =geneInfo.gland, 
                 design =  "Gland" )


Gland.deseq.DE = DESeq(Gland.deseq)

as.data.frame(results(Gland.deseq.DE)) %>%
  rownames_to_column(var = "gene_id") %>%
  rename(Gland_FC = "log2FoldChange") %>%
  rename(Gland_padj =  "padj") %>%
  select(gene_id,Gland_FC,Gland_padj) -> result



geneInfo.OPLS.tissue %>% left_join(result) -> geneInfo.OPLS.tissue.gland
```

### Ampullate glands vs rest

```{r whole spider fragments7, message=FALSE, warning=FALSE}

sampleInfo.tissues.glands %>%
  filter(Tissue %in% c("Major","Minor","Aggregate",
                       "AciniPyri","Flagelliform")) %>% 
  filter(Part == "Whole") %>% 
  mutate(Ampullate = "NO") %>%
  mutate(Ampullate = replace(x = Ampullate,list = Tissue %in% c("Major","Minor"), values = "YES"))->
  sampleInfo.tissues.glands.ampullates


# geneInfo
geneInfo.OPLS ->  geneInfo.tissues.glands.ampullates


ampullate.deseq = getDEseq2Object(count.df = fc.all.qc,
                 md =sampleInfo.tissues.glands.ampullates,
                 gi =geneInfo.tissues.glands.ampullates, 
                 design =  "Ampullate" )


ampullate.deseq.DE = DESeq(ampullate.deseq)

as.data.frame(results(ampullate.deseq.DE)) %>%
  rownames_to_column(var = "gene_id") %>%
  rename(ampullate_FC = "log2FoldChange") %>%
  rename(ampullate_padj =  "padj") %>%
  select(gene_id,ampullate_FC,ampullate_padj) -> result



geneInfo.OPLS.tissue.gland %>% left_join(result) -> geneInfo.OPLS.tissue.gland.ampullate




```

### Major vs Minor
```{r }

sampleInfo.tissues.glands.ampullate %>%
  filter(Tissue %in% c("Major","Minor")) %>% 
  filter(Part == "Whole") ->
  sampleInfo.tissues.glands.ampullate.separated


# geneInfo
geneInfo.OPLS ->
  geneInfo.tissues.glands.ampullate.separated
 

ampullate.deseq.diff = getDEseq2Object(count.df = fc.all.qc,
                 md =sampleInfo.tissues.glands.ampullate.separated,
                 gi =geneInfo.tissues.glands.ampullate.separated, 
                 design =  "Tissue" )


ampullate.deseq.diff.DE = DESeq(ampullate.deseq.diff)

as.data.frame(results(ampullate.deseq.diff.DE)) %>%
  rownames_to_column(var = "gene_id") %>%
  rename(Major_FC = "log2FoldChange") %>%
  rename(Major_padj =  "padj") %>%
  select(gene_id,Major_FC,Major_padj) -> result



geneInfo.OPLS.tissue.gland.ampullate %>% left_join(result) -> geneInfo.OPLS.tissue.gland.ampullate.major



geneInfo.OPLS.tissue.gland.ampullate.major %>% filter(Major_padj < 0.05) %>% ggplot(aes(Major_FC,Separated_p1, color=log10(Major_padj) )) + geom_point()



```




### Sample and Gene info 

```{r}
sampleInfo.OPLS -> sampleInfo.OPLS.DEseq2
geneInfo.OPLS.tissue.gland.ampullate.major -> geneInfo.OPLS.DEseq2



geneInfo.OPLS.DEseq2 %>% filter(Tissue != "Head") %>% filter(Ampullate == "YES") %>%
  filter(Separated_p1 <0.1 )

```



# Part comparison

## Major 

```{r whole spider fragments4, message=FALSE, warning=FALSE}

sampleInfo.all %>%
  filter(Tissue %in% c("Major")) %>%
  filter(Part != "Whole")->
  sampleInfo.major.parts

# geneInfo
geneInfo.OPLS.DEseq2 %>%
  filter(Tissue != "Head") %>%
  filter(Ampullate == "YES") %>%
  filter(Separated_p1 <0.05 ) ->
  geneInfo.major.parts


normExpression.major.parts = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.major.parts,
                 gi =geneInfo.major.parts, 
                 design =  "Part" )

PLS.major.parts = getPLS(normExpression = normExpression.major.parts,
                     md =sampleInfo.major.parts,
                     gi =geneInfo.major.parts, 
                     design =  "Part")


sampleInfo.major.parts.PLS = getMetaDataFromOPLS(OPLS =  PLS.major.parts, 
                    md = sampleInfo.major.parts,
                    extension =  "MAG_part" ) %>%
  select(sample,Part,MAG_part_p1,MAG_part_p2)


  

geneInfo.major.parts.PLS = getGeneInfoFrom2DPLS(OPLS =  PLS.major.parts, 
                    md = sampleInfo.major.parts,
                    gi = geneInfo.major.parts,
                    extension =  "MAG_part" ) %>% 
  mutate(MAG_part_Distance = (MAG_part_p1^2+MAG_part_p2^2)^0.5)




sampleInfo.major.parts.PLS %>% 
  ggplot(aes(x = MAG_part_p1, y = MAG_part_p2, color = Part)) +
  geom_point(shape = 18, size = 6)+ theme_no_axes() + 
  scale_color_manual(values = MajorColors)
#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure2/PLS.sampleInfo.majorParts.2D.pdf", width = 6, height = 5)


geneInfo.major.parts.PLS  %>%
  mutate(MajorParts = "Undefined") %>% 
  mutate(MajorParts = replace(MajorParts,
                              MAG_part_Distance > 0.01 & MAG_part_p1 < 0 ,
                              "Tail")) %>%
  mutate(MajorParts = replace(MajorParts,
                              MAG_part_Distance > 0.01 &
                                MAG_part_p1 > 0 &
                                MAG_part_p2 <= -0.005 ,"Sac")) %>%
  mutate(MajorParts = replace(MajorParts,MAG_part_Distance > 0.01 &
                                MAG_part_p1 > 0 &
                                MAG_part_p2 >= 0.005 ,"Duct")) ->
  geneInfo.major.parts.PLS
# Keeping only major expression

geneInfo.major.parts.PLS %>% 
  inner_join(getGeneInfoFull()) %>%
  inner_join(proteomicsData.unique.silk.gland.signalP)->
  geneInfo.major.parts.PLS.protein


library(ggrepel)



geneInfo.major.parts.PLS.protein %>% 
  filter(MajorParts != "Undefined") ->
  geneInfo.QC.major.PLS.proteins.all

 geneInfo.QC.major.PLS.proteins.all %>%
  ggplot(aes(x = MAG_part_p1, y = MAG_part_p2))+ 
  geom_density2d(data =geneInfo.major.parts.PLS, aes(x = MAG_part_p1, y = MAG_part_p2,color = MajorParts,), size = 0.8 , adjust = 0.8) +   geom_point( alpha = 0.6)+ geom_text_repel(aes(label = geneName), size = 3, color = "black",force = 3,max.overlaps = 20) +
theme_light() +
  xlim(c(-0.04,0.04)) + 
  ylim(c(-0.04,0.04)) + 
  scale_color_manual(values = MajorColors) 
  

genes/samples

ggsave("~/Box/bridgeSpider/paper1/main/figures/figure3/PLS_map.pdf", width = 6,height = 7)
  
```


```{r whole spider fragments4, message=FALSE, warning=FALSE}

sampleInfo.all %>%
  filter(Tissue %in% c("Minor")) %>%
  filter(Part != "Whole")->
  sampleInfo.minor.parts

# geneInfo
geneInfo.OPLS.DEseq2 %>%
  filter(Tissue != "Head") %>%
  filter(Ampullate == "YES")  ->
  geneInfo.minor.parts


normExpression.minor.parts = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.minor.parts,
                 gi =geneInfo.minor.parts, 
                 design =  "Part" )

PLS.minor.parts = getPLS(normExpression = normExpression.minor.parts,
                     md =sampleInfo.minor.parts,
                     gi =geneInfo.minor.parts, 
                     design =  "Part")

getP

sampleInfo.minor.parts.PLS = getMetaDataFromOPLS(OPLS =  PLS.minor.parts, 
                    md = sampleInfo.minor.parts,
                    extension =  "MIG_part" ) %>%
  select(sample,Part,MIG_part,MIG_part)


  

geneInfo.minor.parts.PLS = getGeneInfoFrom2DPLS(OPLS =  PLS.minor.parts, 
                    md = sampleInfo.minor.parts,
                    gi = geneInfo.minor.parts,
                    extension =  "MAG_part" ) %>% 
  mutate(MAG_part_Distance = (MAG_part_p1^2+MAG_part_p2^2)^0.5)




sampleInfo.major.parts.PLS %>% 
  ggplot(aes(x = MAG_part_p1, y = MAG_part_p2, color = Part)) +
  geom_point(shape = 18, size = 6)+ theme_no_axes() + 
  scale_color_manual(values = MajorColors)
#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure2/PLS.sampleInfo.majorParts.2D.pdf", width = 6, height = 5)


geneInfo.major.parts.PLS  %>%
  mutate(MajorParts = "Undefined") %>% 
  mutate(MajorParts = replace(MajorParts,
                              MAG_part_Distance > 0.01 & MAG_part_p1 < 0 ,
                              "Tail")) %>%
  mutate(MajorParts = replace(MajorParts,
                              MAG_part_Distance > 0.01 &
                                MAG_part_p1 > 0 &
                                MAG_part_p2 <= -0.005 ,"Sac")) %>%
  mutate(MajorParts = replace(MajorParts,MAG_part_Distance > 0.01 &
                                MAG_part_p1 > 0 &
                                MAG_part_p2 >= 0.005 ,"Duct")) ->
  geneInfo.major.parts.PLS
# Keeping only major expression

geneInfo.major.parts.PLS %>% 
  inner_join(getGeneInfoFull()) %>%
  inner_join(proteomicsData.unique.silk.gland.signalP)->
  geneInfo.major.parts.PLS.protein


library(ggrepel)



geneInfo.major.parts.PLS.protein %>% 
  filter(MajorParts != "Undefined") ->
  geneInfo.QC.major.PLS.proteins.all

 geneInfo.QC.major.PLS.proteins.all %>%
  ggplot(aes(x = MAG_part_p1, y = MAG_part_p2))+ 
  geom_density2d(data =geneInfo.major.parts.PLS, aes(x = MAG_part_p1, y = MAG_part_p2,color = MajorParts,), size = 0.8 , adjust = 0.8) +   geom_point( alpha = 0.6)+ geom_text_repel(aes(label = geneName), size = 3, color = "black",force = 3,max.overlaps = 20) +
theme_light() +
  xlim(c(-0.04,0.04)) + 
  ylim(c(-0.04,0.04)) + 
  scale_color_manual(values = MajorColors) 
  

genes/samples

ggsave("~/Box/bridgeSpider/paper1/main/figures/figure3/PLS_map.pdf", width = 6,height = 7)
  
```




## Expression heatmap
```{r}
 
bulk.data.merged.QC.rpm.2 %>% 
  filter( Tissue %in%  c("Tail","Sac","Duct")) ->
  geneinfo.rpm.major.parts
geneinfo.rpm.major.parts %>%
  inner_join(geneInfo.extra) %>%
  dplyr::filter(geneName %in% proteomicsData.unique.silk.gland.signalP$geneName) ->
  geneinfo.rpm.major.parts.silk

  ggplot(geneinfo.rpm.major.parts.silk, aes( x = sample2, y = geneName, fill = zscore)) + 
  geom_tile()

#ggsave("~/Box/bridgeSpider/bulkRNA/figures/PLS.geneInfo.majorParts.2D.pdf", width = 6, height = 5)


geneinfo.rpm.major.parts.silk %>%
  select(zscore, geneName, sample2) %>%
  pivot_wider(names_from = "sample2", values_from = "zscore") %>%
  column_to_rownames("geneName") ->
  geneinfo.rpm.major.parts.silk.wide
  
geneinfo.rpm.major.parts.silk.wide


library(cluster)
library(factoextra)


m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

df = geneinfo.rpm.major.parts.silk.wide
ac <- function(x) {
  agnes(df, method = x)$ac
}


sapply(m, ac)
clust <- agnes(df, method = "ward")

pltree(clust, cex = 0.6, hang = -1, main = "Dendrogram") 

gap_stat <- clusGap(df, FUN = hcut, nstart = 25, K.max = 10, B = 50 )

#produce plot of clusters vs. gap statistic
fviz_gap_stat(gap_stat)

d <- dist(df, method = "euclidean")

#perform hierarchical clustering using Ward's method
final_clust <- hclust(d, method = "ward.D2" )
#cut the dendrogram into 4 clusters

genes = data.frame(geneName = names(cutree(final_clust, 7)) ,
                   hierarchicalCluster = paste("c",cutree(final_clust, 7),sep ="_")) %>%   mutate(hierarchicalCluster2 = hierarchicalCluster) %>%
  mutate(hierarchicalCluster2 = gsub("c_2","tail_bulk_1",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_3","tail_bulk_2",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_4","tail_bulk_3",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_1","tail_bulk_4",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_5","sac_bulk_2",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_7","sac_bulk_3",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_6","sac_bulk_1",hierarchicalCluster2)) %>%
  arrange(hierarchicalCluster2)






d2 <- dist(t(df), method = "euclidean")

#perform hierarchical clustering using Ward's method
row_clust <- hclust(d, method = "ward.D2" )
col_clust <- hclust(d2, method = "ward.D2" )

row_clust$order
row_clust$merge
  
clusters = genes$hierarchicalCluster2
names(clusters) = genes$geneName

clusters= clusters[rownames(geneinfo.rpm.major.parts.silk.wide)]

row_ha_left = rowAnnotation(Markers = clusters,col = list(Markers = bulkClustersH))


colnames(geneinfo.rpm.major.parts.silk.wide)  = c("Duct_1","Duct_2","Duct_3","Duct_4","Duct_5",
                                                  "Sac_1","Sac_2","Sac_3","Sac_4","Sac_5","Sac_6",
                                                  "Tail_1","Tail_2","Tail_3","Tail_4","Tail_5","Tail_6")

pdf("~/Box/bridgeSpider/paper1/main/figures/figure3/heatmap.silkproteins.expression.mojor.parts.pdf", width = 6, height = 5)

ComplexHeatmap::Heatmap(geneinfo.rpm.major.parts.silk.wide, row_split = clusters, cluster_columns = FALSE,column_order = c(17:1),
                         left_annotation = row_ha_left)
dev.off()


write_csv(x = genes, file = "../../results/bulkRNA/silkProteins.RNA.hclusters.csv")
final_clust$d



```





```{r whole spider fragments5, message=FALSE, warning=FALSE}

# SampleInfo
sampleInfo.OPLS.DEseq2 %>%
  filter(Tissue %in%  c("Head","WholeBody")) ->
  sampleInfo.all.tissues

# geneInfo
geneInfo.all.qc ->
  geneInfo.tissues
  

normExpression.parts = getVSTexpression(count.df = fc.all.qc,
                 md =sampleInfo.all.tissues,
                 gi =geneInfo.tissues, 
                 design =  "Tissue" )

OPLS.parts = getOPLS(normExpression = normExpression.parts,
                 md =sampleInfo.all.tissues,
                 gi =geneInfo.tissues, 
                 design =  "Tissue"
          )


sampleInfo.tissues = getMetaDataFromOPLS(OPLS =  OPLS.parts, 
                    md = sampleInfo.all.tissues,
                    extension =  "Tissue" ) 

geneInfo.tissues = getGeneInfoFromOPLS(OPLS =  OPLS.parts, 
                    md = sampleInfo.all.tissues,
                    extension =  "Tissue",gi = geneInfo.tissues )


geneInfo.all.qc %>% left_join(geneInfo.tissues) -> geneInfo.tissues
sampleInfo.all %>% left_join(sampleInfo.tissues) -> sampleInfo.tissues



```















#### Comparing against protein expression
```{r}

normExpression.major = getNormalisedExpression(count.df = bulk.data.merged.QC.long,
                 md =sampleInfo.QC.merged.major,
                 gi =geneInfo.QC, 
                 design =  "Tissue" )


normExpression.major %>% rownames_to_column(var = "gene_id") %>%
  pivot_longer(-gene_id,names_to = "sample", values_to = "norm_exp") %>%
  inner_join(sampleInfo.QC.merged.major) %>%
  filter(Tissue == "Major") %>%
  group_by(gene_id, Tissue) %>%
  summarise(norm_exp = mean(norm_exp)) %>%
  inner_join(getGeneInfoFull()) ->
  bulk.major

bulk.major %>% select(geneName, norm_exp ) %>% inner_join(proteinsInSilk) ->
  bulk.major.expression

bulk.major %>% select(geneName, norm_exp ) %>% inner_join(proteomicsData.unique.gland) ->
  bulk.major.gland.expression


pearson.corr = cor.test(x =bulk.major.gland.expression$norm_exp, y = bulk.major.gland.expression$mean )

write_csv(bulk.major.gland.expression, "../../data/bulkRNA/normExpression.protein.csv") 
write_csv(bulk.major.expression, "../../data/bulkRNA/normExpression.protein.silk.csv") 
  


bulk.major.expression = read_csv("../../data/bulkRNA/normExpression.protein.silk.csv")
bulk.major.gland.expression = read_csv("../../data/bulkRNA/normExpression.protein.csv")
bulk.major.expression %>% filter(geneName %in% protein.info$geneName)->
  bulk.major.expression


ggplot(data =bulk.major.gland.expression %>% filter(norm_exp > 6),  
       aes(x = mean, y = norm_exp))+
  geom_point(color = "blue", alpha = 0.3, size = 0.2)+
  geom_density2d(data =bulk.major.gland.expression,  aes(x = mean, y = norm_exp))+ 
  geom_smooth(method = "lm")+
  geom_point(data =bulk.major.expression,  aes(x = mean, y = norm_exp), color = "black") + 
  geom_text_repel(data =bulk.major.expression,
                mapping = aes(x = mean,
                              y = norm_exp,label = geneNameProtein))+ 
  stat_cor(p.accuracy = 0.0000000000000001, r.accuracy = 0.01,label.x = -2.5,label.y = 12, color = "blue")+
  theme_light()+
  scale_x_continuous(trans='log2') +
  scale_y_continuous(trans='log2') +
  ylab("Normalised RNA expression") + 
  xlab("Normalised Protein expression") 
  


ggsave("~/Box/bridgeSpider/paper1/supplementary/SupplFig/RNA.protein.expression.pdf")


bulk.major.expression
bulk.major.gland.expression
 bulk.major.gland.expression$norm_exp = 2^bulk.major.gland.expression$norm_exp

 
bulk.major %>%  arrange(-norm_exp) ->
  bulk.major.ordered
bulk.major.ordered$rank = 1:10650



bulk.major.ordered %>%
  select(geneName, rank) %>% inner_join(proteinInfo.filtered)->
  bulk.major.ordered


 bulk.major.ordered%>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder2)) %>% 
  arrange(geneNameProtein) 
```
#### Visualise pattern
```{r}



geneInfo.QC.part.major %>%
  filter(!is.na(Major_p1)) %>%
  ggplot(aes(x = Major_p1, color = Major)) + 
  geom_density()-> major.sampleInfo.plot

sampleInfo.QC.merged.part.major %>%
  filter(!is.na(Major_p1)) %>% 
  ggplot(aes(x = Major_p1, color = Tissue)) + 
  geom_density()-> 
  major.geneInfo.plot





patchwork <- major.sampleInfo.plot / major.geneInfo.plot
patchwork + plot_annotation(
  title = 'Figure Spider Major',
  subtitle = 'Comparing whole body and Major gland'
)

```
### Gland analysis

```{r whole spider fragments3, message=FALSE, warning=FALSE}

sampleInfo.QC.merged.part %>%
  filter(Tissue  %in% c("AciniPyri","Aggregate","Flagelliform","Major","Minor"))  ->
  sampleInfo.QC.merged.glands

# geneInfo
geneInfo.glands = geneInfo.QC.part %>% filter(part == "WholeBody")


normExpression.gland = getVSTexpression(count.df = bulk.data.merged.QC.long,
                 md =sampleInfo.QC.merged.glands,
                 gi =geneInfo.glands, 
                 design =  "Tissue" )

PLS.gland = getPLS(normExpression = normExpression.gland,
                     md =sampleInfo.QC.merged.glands,
                     gi =geneInfo.glands, 
                     design =  "Tissue")


sampleInfo.QC.merged.glands = getMetaDataFromOPLS(OPLS =  PLS.gland, 
                    md = sampleInfo.QC.merged.glands,
                    extension =  "glands" ) 
  

geneInfo.glands = getGeneInfoFrom2DPLS(OPLS =  PLS.gland, 
                    md = sampleInfo.QC.merged.glands,
                    gi = geneInfo.glands,
                    extension =  "glands" ) 


geneInfo.QC.part %>% left_join(geneInfo.glands) -> geneInfo.QC.part.glands
sampleInfo.QC.merged.part %>% left_join(sampleInfo.QC.merged.glands) -> sampleInfo.QC.merged.part.glands





```



#### Visualise pattern
```{r}



geneInfo.QC.part.glands %>%
  filter(!is.na(glands_p1)) %>%
  ggplot(aes(x = glands_p1, glands_p2 )) + 
  geom_point()-> Tubuli.sampleInfo.plot

sampleInfo.QC.merged.part.glands %>%
  filter(!is.na(glands_p1)) %>% 
  ggplot(aes(x = glands_p1, glands_p2, color = Tissue)) + 
  geom_point()-> 
  Tubuli.geneInfo.plot





patchwork <- Tubuli.sampleInfo.plot / Tubuli.geneInfo.plot
patchwork + plot_annotation(
  title = 'Figure Spider glands',
  subtitle = 'Seperating the glands'
)

```




# Expression glands analysis

```{r}

sampleInfo.QC.merged %>%
  filter(Tissue %in%  c("Head",
                        "WholeBody",
                        "AciniPyri",
                        "Flagelliform",
                        "Major",
                        "Minor",
                        "Tubuli",
                        "Aggregate"
                        )) ->
  sampleInfo.QC.merged.wholeTissues

				
	
# geneInfo
geneInfo.parts = geneInfo.QC



normExpression.parts = getNormalisedExpression(count.df = bulk.data.merged.QC.long,
                 md =sampleInfo.QC.merged.wholeTissues,
                 gi =geneInfo.parts, 
                 design =  "Tissue" )


normExpression.parts %>% rownames_to_column(var = "gene_id") %>%
  pivot_longer(cols = -gene_id, names_to = "sample",values_to = "normExpression") ->normExpression.parts.long 
  

normExpression.parts.long %>% inner_join(sampleInfo.QC.merged.wholeTissues)->   normExpression.parts.long.silk


normExpression.parts.long.silk %>% group_by(Tissue,gene_id) %>%
  summarise(mean = mean(normExpression)) %>%
  inner_join(getGeneInfoFull()) %>%
  filter(Group == "Spidroin") %>%
  filter(Tissue != "Tubuli") %>%
   select(geneName, mean, Tissue)%>%
  pivot_wider(values_from = "mean", names_from = "Tissue") %>%
  arrange(geneName) %>%
  select(geneName,Aggregate,AciniPyri,Flagelliform,Major,Minor, WholeBody,Head) %>%
  column_to_rownames(var = "geneName")  ->
  normExpression.parts.long.silk.summary.real

normExpression.parts.long.silk.summary.real

pdf("~/Box/bridgeSpider/paper1/main/figures/figure1/expression.heatmap.real.pdf", width = 3, height=6)
ComplexHeatmap::Heatmap(
  normExpression.parts.long.silk.summary.real, cluster_columns = FALSE, cluster_rows = FALSE)
dev.off()

#normExpression.parts.long.silk.summary %>%
#  group_by(geneName) %>%
#  summarise(max = max(mean)) %>%
#  inner_join(normExpression.parts.long.silk.summary) %>%
#  mutate(relative = mean/max)%>%
#  select(geneName, relative, Tissue) %>%
#  pivot_wider(values_from = "relative", names_from = "Tissue") %>%
#  column_to_rownames(var = "geneName")->
#  normExpression.parts.long.silk.summary.relative
# ComplexHeatmap::Heatmap(normExpression.parts.long.silk.summary.relative)


normExpression.parts.long.silk %>% group_by(Tissue,gene_id) %>%
  summarise(mean = mean(normExpression)) %>%
  inner_join(getGeneInfoFull()) %>%
  filter(Group == "Spidroin") %>%
  mutate(log = log2(mean+1)) %>%
   select(geneName, log, Tissue)  %>%
  pivot_wider(values_from = "log", names_from = "Tissue") %>%
  column_to_rownames(var = "geneName") %>%
  select(Aggregate,AciniPyri,Flagelliform,Major,Minor, WholeBody,Head)->
  normExpression.parts.long.silk.summary.log
ComplexHeatmap::Heatmap(normExpression.parts.long.silk.summary.log)


newOrder = c("MaSp1a","MaSp1c","MaSp1b",
  "MaSp2b","MaSp2c","MaSp2e","MaSp2f","MaSp3a","MaSp3b",
  "MaSp2a","MaSp4","MaSp2d","Spidroin1","Spidroin3",
  "MiSpa","MiSpb","MiSpc","MiSpd",
  "AgSp1","AgSp2b","AgSp2a","Spidroin2",
  "FlSpa","FlSpb","FlSpc","FlSp-like",
  "TuSpa","TuSpb","TuSpc","TuSp-like",
  "PySpa","PySpb",
  "AcSpa","AcSp1b","AcSp1c"
  )
setdiff(newOrder, rownames(normExpression.parts.long.silk.summary.log))
normExpression.parts.long.silk.summary.log = normExpression.parts.long.silk.summary.log[newOrder,]
ComplexHeatmap::Heatmap(normExpression.parts.long.silk.summary.log)

normExpression.parts.long.silk.summary.real = normExpression.parts.long.silk.summary.real[newOrder,]
pdf("~/Box/bridgeSpider/paper1/main/figures/figure1/expression.heatmap.real.pdf", width = 3, height=6)
ComplexHeatmap::Heatmap(
  normExpression.parts.long.silk.summary.real, cluster_columns = FALSE, cluster_rows = FALSE)
dev.off()




```





```{r}



geneInfo.QC %>% select(gene_id, AciniPyri,Major,Minor,Aggregate,Flagelliform,Tubuli  ) %>%
  pivot_longer(-gene_id, names_to = "Gland", values_to = "Selection") %>%
  filter(!is.na(Selection)) %>% filter(Selection != "WholeBody")  %>%
  select(gene_id, Gland) %>% mutate(present = as.integer(1))  %>%  pivot_wider(names_from = Gland, values_from = present, values_fill = as.integer(0)) %>% as.data.frame()->
  geneInfo.QC.Glands.present





geneInfo.QC.all %>% select(gene_id, AciniPyri,Major,Minor,Aggregate,Flagelliform,Tubuli  ) %>%
  pivot_longer(-gene_id, names_to = "Gland", values_to = "Selection") %>%
  filter(!is.na(Selection)) %>% filter(Selection != "WholeBody")  %>%
  select(gene_id, Gland) %>% mutate(present = as.integer(1)) ->geneInfo.QC.Glands.present.df

geneInfo.QC.Glands.present.df%>%
  group_by(gene_id) %>% summarise(nrOfGlands = sum(present)) ->
  geneInfo.QC.Glands.present.summary
  
geneInfo.QC.Glands.present.summary  %>% filter(nrOfGlands > 1 & nrOfGlands <6)  ->
  glandUnspecific

pdf("~/Box/bridgeSpider/bulkRNA/figures/gland.genes.common.pdf")
UpSetR::upset(geneInfo.QC.Glands.present,
      nsets = 6,  order.by = "freq", cutoff = 15,
    mainbar.y.label = "Common genes", sets.x.label = "Gland specific genes", 
    text.scale = c(1.3, 1.3, 1, 1, 2, 0.75), show.numbers = FALSE)
dev.off()


geneInfo.QC.Glands.present %>% filter(gene_id %in% glandUnspecific$gene_id) ->
  geneInfo.QC.Glands.present.2

pdf("~/Box/bridgeSpider/bulkRNA/figures/gland.genes.common.heatmap.pdf")
ComplexHeatmap::Heatmap(geneInfo.QC.Glands.present.2[,-1]) 
dev.off()


```

## Protein distributions
```{r}

genes %>%
  mutate(hierarchicalCluster2 = 
           factor(hierarchicalCluster2,
                  levels = c("tail_bulk_1","tail_bulk_2",
                           "tail_bulk_3","tail_bulk_4",
                           "sac_bulk_1","sac_bulk_2",
                           "sac_bulk_3","NotInSilk"))) ->
  genes

proteomicsData.unique.gland %>% left_join(genes) %>%
  mutate(hierarchicalCluster2 = replace(hierarchicalCluster2,
                                        is.na(hierarchicalCluster2),
                                        "NotInSilk")) %>%
  mutate(tissue = "Gland") ->
  proteomicsData.unique.gland.bulk

proteomicsData.unique.silk %>%
  inner_join(genes) %>%
  mutate(tissue = "Silk") ->
  proteomicsData.unique.silk.bulk
  
proteomics = bind_rows(proteomicsData.unique.gland.bulk,proteomicsData.unique.silk.bulk)

proteomics %>% group_by(hierarchicalCluster2,tissue ) %>%
  summarise(sum = sum(mean)) %>%
  arrange(tissue,hierarchicalCluster2 )->
  proteomics.summary


ggplot(proteomics.summary, aes(x = tissue, y = sum, fill = hierarchicalCluster2)) +
  geom_bar(, stat = "identity", position="fill" ) + 
  scale_fill_manual(values = bulkClustersH)+
  theme_light()+ theme(legend.position = "none")+
  coord_polar("y", start=0) 
  

glandCircle = proteomics.summary %>% filter(tissue == "Gland") %>%
  ggplot( aes(x = tissue, y = sum, fill = hierarchicalCluster2)) +
  geom_bar(, stat = "identity", position="fill" ) + 
  scale_fill_manual(values = bulkClustersH)+
  theme_light()+ 
  coord_polar("y", start=0) 
  

silkCircle = proteomics.summary %>% filter(tissue != "Gland") %>%
  ggplot( aes(x = tissue, y = sum, fill = hierarchicalCluster2)) +
  geom_bar(, stat = "identity", position="fill" ) + 
  scale_fill_manual(values = bulkClustersH)+
  theme_light()+ 
  coord_polar("y", start=0) 


glandCircle+silkCircle + plot_layout(guides = "collect")
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure3/protein.expression.circleDiagram.pdf", width = 2,height = 4)

  



```



### Marker genes
```{r}


sampleInfo.QC.merged %>%
  filter(Tissue %in% c("Tail","Sac","Duct")) ->
  sampleInfo.QC.merged.major.parts

# geneInfo
geneInfo.major.parts = geneInfo.QC.all %>% filter(Major == "Major")


normExpression.major.parts.DEseq2 = getDEseq2Object(count.df = bulk.data.merged.QC.long,
                 md =sampleInfo.QC.merged.major.parts,
                 gi =geneInfo.major.parts, 
                 design =  "Tissue" )


normExpression.major.parts.DEseq2 = DESeq(normExpression.major.parts.DEseq2)

results(normExpression.major.parts.DEseq2)
normExpression.major.parts.DEseq2$

geneInfo.MAGpart %>%
  mutate(majorDistance = (MAG_part_p1^2 +MAG_part_p2 ^2)^0.5) %>%
  as.data.frame() %>%
  group_by(MajorParts ) %>%
  arrange(-majorDistance) %>%
  top_n(n = 100) %>%
  inner_join(geneInfo.extra) %>%
  select(MajorParts,geneName,
         MAG_part_p1,MAG_part_p2,
         majorDistance) %>% 
  arrange(MajorParts, -majorDistance )  %>%
  filter(grep(geneName, pattern = "MaSp" ) )




```

