---
title: "Figures for BulkRNA"
author: "Johan"
date: "01/04/2024"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    df_print: paged


---






```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(RColorBrewer)
library(DESeq2)
library(ropls)
library(ggforce)
library(patchwork)
library(ggpubr)
library(ggrepel)
library(gridExtra)


# Avoid namespace conflicts
rename <- dplyr::rename
summarise <- dplyr::summarise                
select <- dplyr::select
group_by <- dplyr::group_by

# Load General Annotation Path
# Note: ".." steps out of MinorGland_paper into the shared Rising-Lab folder
signalPFile <- here("..", "annotation", "signalP", "Lsc.v1.1.SignalP.v6.tsv")

# Source scripts from local project folder
source(here("MinorGland_paper/scripts", "loadGeneInfo.R"))
source(here("MinorGland_paper/scripts", "colorSchemes_minor.R"))
source(here("MinorGland_paper/scripts", "customFunctions.R"))


directories <- c(
  here("MinorGland_paper", "results", "supplementaryTables"),
  here("MinorGland_paper", "figures", "figure2"),
  here("MinorGland_paper", "figures", "figure3"),
  here("MinorGland_paper", "figures", "supplementaryFigures"),
  here("MinorGland_paper", "figures", "supplementaryFigures","bulkRNA"),
  here("MinorGland_paper", "figures", "supplementaryFigures","proteomics"),
  here("MinorGland_paper", "figures", "supplementaryFigures","scRNA"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig1"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig2"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig3"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig4"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig5"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig6"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig7"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig8"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig9"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig10"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig11"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig12"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig13"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig14"),
  here("MinorGland_paper", "figures", "supplementaryFigures","SupplFig15")

  
)

# Loop through and create them if they don't exist
walk(directories, ~if(!dir.exists(.x)) dir.create(.x, recursive = TRUE))


```




# Functions


## Gene information
```{r Gene information, message=FALSE, warning=FALSE}


geneInfo <- read_csv(here( "annotation", "Lsc.v1.0.3.geneInfo.csv.gz"))
termOrder = c("MiSp","AmSp-like","Spidroins","SpiCE",
             "SignalPeptide","RibosomalProteins","NoSignalPeptide")


geneInfo %>%
  mutate(Term = signalP) %>%
  mutate(Term = replace(Term,Term == "Yes","SignalPeptide")) %>%
  mutate(Term = replace(Term,Term == "No","NoSignalPeptide")) %>%
  mutate(Term = replace(Term, grepl("MiSp",geneName),"MiSp")) %>%
  mutate(Term = replace(Term, grepl("MaSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("AcSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("AgSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("FlSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("TuSp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("PySp",geneName),"Spidroins")) %>%
  mutate(Term = replace(Term, grepl("^(rpl|rps)",geneName),"RibosomalProteins")) %>%
  mutate(Term = replace(Term, grepl("^SpiCE",geneNameProtein),"SpiCE")) %>%
  mutate(Term = replace(Term, grepl("^AmSp-like",geneNameProtein),"AmSp-like")) %>%
  mutate(Term = factor(Term, 
                       levels = termOrder) ) ->
  geneInfo






```
## Functional annotation 
### Supplementary table 1: Spacer domain counts
```{r}

col_names <- c(
  "protein_acc", "sequence_md5", "sequence_length", "analysis",
  "signature_accession", "signature_description", "start", "stop",
  "score", "status", "date", "interpro_accession",
  "interpro_description", "go_terms", "pathways"
)
interpro5 <- read_tsv(
  here("annotation", "interpro_5", "Lsc.v1.1.genes.AA.longest.fasta.tsv"), 
  col_names = col_names, 
  skip = 0, 
  na = "-"
) %>%
  rename(geneName = "protein_acc")

interpro5 %>% filter(grepl( "1.10.274.60",signature_accession)) %>%
  inner_join(geneInfo) %>%
  group_by(geneNameProtein,lascID,signature_accession) %>%
  summarise(nrOfDomains = n()) %>%
  mutate(class = "Other") %>%
  mutate(class = replace(class, grepl("AmSp-like",geneNameProtein), "AmSp-like")) %>%
  mutate(class = replace(class, grepl("MiSp",geneNameProtein), "MiSp")) %>%
  mutate(class = replace(class, grepl("PySp",geneNameProtein), "PySp")) %>%
  mutate(class = replace(class, grepl("TuSp",geneNameProtein), "TuSp")) %>%
  mutate(class = replace(class, grepl("AcSp",geneNameProtein), "AcSp")) %>%
  mutate(class = factor(class, levels = c("PySp","TuSp","AcSp","MiSp","AmSp-like","Other"))) %>%
  arrange(class,-nrOfDomains)  ->
  interpro5.spacer


# Updated directory creation and saving for this section
dir.create(here("MinorGland_paper", "results", "supplementaryTables"), recursive = TRUE, showWarnings = FALSE)

write_csv(
  interpro5.spacer, 
  here("MinorGland_paper", "results", "supplementaryTables", "supp.tab.1.spacer_domain_counts.csv")
)
```


## Parameters
```{r Parameters, message=FALSE, warning=FALSE}

minExpressionCutoff =20
DataGenerated = TRUE


```



# Proteomics data

## Present proteins in proteomics data
```{r}
# Load from project subfolder
minorProteomics <- read_csv(here("MinorGland_paper", "data", "proteomics", "minor", "minorAmpullateEpxression.csv.gz"))

minorProteomics %>%
  filter(Sample != "Mi_Dope_1") %>%
  group_by(Part, geneName) %>%
  summarise(numberOfSamples = n(), 
            averageIntensity = sum(relativeIntensity)/numberOfSamples,
            .groups = "drop") %>% # Added .groups = "drop" to prevent grouped-df warnings
  group_by(geneName) %>%
  summarise(max = max(averageIntensity)) %>%
  filter(max > 0.0001) %>% 
  arrange(-max) ->
  minorProteomics.expressed

minorProteomics %>% 
  inner_join(minorProteomics.expressed, by = "geneName") ->
  minorProteomics.QC

minorProteomics.QC %>% 
  distinct(spider, Part) %>%
  group_by(Part) %>%
  summarise(numberOfSpiders = n()) %>%
  inner_join(minorProteomics.QC, by = "Part") %>%
  group_by(Part, geneName) %>%
  summarise(numberOfSpiders = n(), 
            averageIntensity = sum(relativeIntensity)/numberOfSpiders,
            .groups = "drop") %>%
  select(geneName, Part, averageIntensity) ->
  minorProteomics.QC


```



## Amino acid composition 

```{r Amino acid composition , message=FALSE, warning=FALSE}
proteinSequenceData.df = read_csv(here("annotation","proteomics","proteinComposition.2024.csv.gz"))
```
### Plots

# RNA analysis
## QC
```{r, message=FALSE, warning=FALSE}


sampleInfo.QC.merged = read_csv(file = here("MinorGland_paper","data","bulkRNA","sampleInfo.QC.merged"))

bulk.data.merged = read_csv(file = here("MinorGland_paper","data","bulkRNA","All.samples.count.csv")) %>%
  rename(gene_id = "Geneid")

bulk.data.merged.QC = bulk.data.merged[,c("gene_id",sampleInfo.QC.merged$sample)]
      

sampleInfo.All = read_csv(file  = here("MinorGland_paper","data","bulkRNA","All.samples.sampleInfo.csv")) 





```


## Minor part data
```{r, message=FALSE, warning=FALSE}

MinorCounts = read_tsv(here("MinorGland_paper","data","bulkRNA","minorParts","minorParts.featureCount.reverse.txt"),
                  skip = 1) %>%
  select(-6:-2) %>%
  pivot_longer(names_to = "sample", values_to = "readCount",-Geneid) %>%
  mutate(sample = gsub("/Aligned.sortedByCoord.out.bam","",sample)) %>%
  rename(gene_id = "Geneid") %>%
  mutate(sample = gsub("PMI","TaPr",sample)) %>%
  mutate(sample = gsub("DMI","SaPr",sample)) %>%
  mutate(sample = gsub("BMI","SaCe",sample)) %>%
  mutate(sample = gsub("CMI","SaDi",sample)) %>%
  mutate(sample = gsub("CDMi","SaDi_Du",sample)) 
  
   

TissueOrder = c("Gland","Tail_proximal","Sac_proximal","Sac_center","Sac_distal","Sac_distal_Duct")

MinorCounts %>%
  group_by(sample) %>%
  summarise(totalCount = sum(readCount)) %>%
  mutate(part = "Gland") %>%
  mutate(MAG = "Part") %>%
  mutate(Tissue =  "Sac_center") %>%
  mutate(Tissue =  replace(Tissue, grepl("SaDi",sample),"Sac_distal")) %>%
  mutate(Tissue =  replace(Tissue, grepl("SaDi_Du",sample),"Sac_distal_Duct")) %>%
  mutate(Tissue =  replace(Tissue, grepl("SaPr",sample),"Sac_proximal")) %>%
  mutate(Tissue =  replace(Tissue, grepl("TaPr",sample),"Tail_proximal")) %>% 
  mutate(Tissue = factor(Tissue,levels = TissueOrder)) %>%
  arrange(Tissue)->
  MinorSampleInfo

sampleOrder = MinorSampleInfo$sample


MinorSampleInfo %>% 
  mutate(sample = factor(sample, levels = sampleOrder)) ->
        MinorSampleInfo   


```

### Gene QC filter

```{r for major, message=FALSE, warning=FALSE}

bulk.data.merged.QC %>% 
  pivot_longer(cols = -gene_id, names_to = "sample",values_to = "count") %>%
  group_by(sample) %>%
  dplyr::summarise(sampleReads = sum(count)) ->
  sample.count.info



bulk.data.merged.QC %>% 
  pivot_longer(cols = -gene_id, names_to = "sample",values_to = "count") %>%
  inner_join(geneInfo ) %>%
  mutate(normLengthCount = count/ tx_len)->
  bulk.data.merged.QC.tpm


bulk.data.merged.QC.tpm %>%  
  group_by(sample) %>%
  dplyr::summarise(sampleReads = sum(normLengthCount)) %>%
  inner_join(bulk.data.merged.QC.tpm) %>%
  mutate(tpm= (normLengthCount/sampleReads*1000000)) ->
  bulk.data.merged.QC.tpm




bulk.data.merged.QC.tpm %>% select(gene_id,sample,tx_len,count,tpm) -> bulk.data.merged.QC.tpm


bulk.data.merged.QC.tpm %>%
  group_by(gene_id) %>% 
  summarise(max_tpm = max(tpm)) %>%
  inner_join(geneInfo) %>%
  mutate(Group = replace(silkGroup, is.na(silkGroup), "Other")) ->  bulk.data.merged.QC.tpm.summary


bulk.data.merged.QC.tpm.summary %>%ggplot(aes(log10(max_tpm ))) + geom_density()


bulk.data.merged.QC.tpm.summary %>%
  select(gene_id,max_tpm) -> 
  geneInfo.tpm


geneInfo.tpm %>% filter(max_tpm > 4)  ->
geneInfo.QC

geneInfo.QC = inner_join(geneInfo.QC, geneInfo)


bulk.data.merged.QC %>% filter(gene_id %in% geneInfo.QC$gene_id) %>% pivot_longer(-gene_id, values_to = "readCount", names_to = "sample") -> bulk.data.merged.QC.long


  
```


```{r bulkRNA data major parts}


bulk.data.merged.QC.tpm %>%
  filter(gene_id %in% geneInfo.QC$gene_id) %>%
  inner_join(sampleInfo.QC.merged) %>%
  filter(Tissue == c("Major") | MAG == c("Part")) %>%
  group_by(gene_id,Tissue, MAG) %>%
  summarise(tpm = mean(tpm)) %>% 
  select(-MAG) %>%
  pivot_wider(names_from = "Tissue" , values_from = "tpm") -> 
  bulk.data.merged.QC.tpm.major.wide






```

```{r QC for minor }
MinorCounts %>% 
  inner_join(geneInfo.QC) %>%
  mutate(normLengthCount = readCount/ tx_len)->
  MinorCounts.QC.tpm



MinorCounts.QC.tpm %>%  
  group_by(sample) %>%
  dplyr::summarise(sampleReads = sum(normLengthCount)) %>%
  inner_join(MinorCounts.QC.tpm) %>%
  mutate(tpm= (normLengthCount/sampleReads*1000000)) ->
  MinorCounts.QC.tpm


MinorCounts.QC.tpm %>% inner_join(MinorSampleInfo) %>%
  group_by(gene_id,Tissue) %>% 
  summarise(max_tpm = max(tpm)) %>%
  inner_join(geneInfo.QC) ->  MinorCounts.QC.tpm.summary






MinorCounts.QC.tpm %>% inner_join(MinorSampleInfo)  %>%
  select(gene_id,Tissue,sample,tpm) %>%
  group_by(gene_id,Tissue) %>%
  summarise(tpm = mean(tpm)) %>%
  pivot_wider(names_from = "Tissue" , values_from = "tpm") %>%
    select(-Sac_distal_Duct)-> 
  MinorCounts.QC.tpm.wide



```








```{r}  

sampleInfo.QC.merged.minor.parts = MinorSampleInfo

normExpression.minor.parts = getVSTexpression(count.df = MinorCounts,
                 md =MinorSampleInfo,
                 gi =geneInfo.QC, 
                 design =  "Tissue" )


normExpression.minor.parts%>% rownames_to_column(var = "gene_id") %>%
  pivot_longer(-gene_id, values_to = "vst", names_to = "sample") %>%
  inner_join(MinorSampleInfo) %>%
  ggplot(aes(y = vst ,x = sample, fill = Tissue )) + geom_boxplot() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> bp.before


badSamples = c("SaCe_4","SaCe_1","SaPr_5","TaPr_1","SaPr_1")


MinorSampleInfo %>% filter(!sample %in% badSamples) -> 
  MinorSampleInfo.QC

normExpression.minor.parts = getVSTexpression(count.df = MinorCounts,
                 md =MinorSampleInfo,
                 gi =geneInfo.QC, 
                 design =  "Tissue" )


normExpression.minor.parts %>% distinct()

MinorCounts.QC.tpm %>% 
  inner_join(MinorSampleInfo) %>%
  mutate(sample = factor(sample, levels = sampleOrder)) %>%
  ggplot(aes(y = log10(tpm) ,x = sample, fill = Tissue )) +
  geom_boxplot()+
  theme_light()+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values = BulkPartMinor) -> tpm.before



normExpression.minor.parts%>% rownames_to_column(var = "gene_id") %>%
  pivot_longer(-gene_id, values_to = "vst", names_to = "sample") %>%
  inner_join(MinorSampleInfo.QC) %>%
  mutate(sample = factor(sample, levels = sampleOrder)) %>%
  ggplot(aes(y = vst ,x = sample, fill = Tissue )) + geom_boxplot() +
  theme_light()+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values = BulkPartMinor)-> bp.after


tpm.before/bp.after+ plot_layout(guides = "collect") -> RNA1.supp.plot
#RNA1.supp.plot
#ggsave("figures/supplementaryFigures/supplementaryFigureRNA1.pdf",width = 160,height = 160,units = "mm",plot = suppFigProt3)

```
#### Expression of silkProteins
```{r}

MinorCounts.QC.tpm  %>%
  inner_join(sampleInfo.QC.merged.minor.parts) %>%
  inner_join(geneInfo.QC) %>%
  filter(grepl("MiSp",geneName))  %>%
  select(geneNameProtein,Tissue,sample,tpm) ->
  MiSpExpression

MinorCounts.QC.tpm  %>%
  inner_join(sampleInfo.QC.merged.minor.parts) %>%
  inner_join(geneInfo.QC) %>%
  filter(!is.na(silkGroup)) %>%
  select(geneNameProtein,Tissue,sample,tpm)->
  silkProteins.expression
  
bind_rows(MiSpExpression,silkProteins.expression) %>%
  mutate(log2TPM = log2(tpm+1))->
  silkProteins.expression

silkProteins.expression %>%
  select(geneNameProtein,sample,log2TPM) %>%
  pivot_wider(names_from = "sample", values_from ="log2TPM") %>%
  column_to_rownames("geneNameProtein")->silkProteins.expression.wide



ComplexHeatmap::Heatmap(silkProteins.expression.wide, column_km = 4, row_km = 4)
pdf(here("MinorGland_paper","figures","supplementaryFigures","bulkRNA","silkProteins.heatMap.pdf"), width = 6, height = 4)
ComplexHeatmap::Heatmap(silkProteins.expression.wide, column_split =  4, row_split = 4) 
dev.off()


```


#### PCA


```{r, message=FALSE, warning=FALSE }
normExpression.minor.parts[,as.character(MinorSampleInfo.QC$sample)] ->normExpression.minor.parts


row_vars <- apply(normExpression.minor.parts, 1, var)

top_n <- 1000
top_var_rows <- normExpression.minor.parts[order(row_vars, decreasing = TRUE)[1:top_n], ]

mir.pca <- prcomp(t(top_var_rows), center = TRUE, scale = TRUE) 
e.var = (mir.pca[['sdev']]^2 / sum(mir.pca[['sdev']]^2))
e.var = as.data.frame( e.var )
e.var$PC = as.factor(1:nrow(e.var)) 
e.var$Variance = e.var$e.var*100
e.var$CumulativeVariance = cumsum(e.var$Variance)
qplot(PC, Variance, data=e.var, geom = c("point")
      , ylim = c(0,max(e.var$Variance)+5) 
      ,ylab = "Variance (%)")+ theme_light() -> PCAvariance

ggsave(here("MinorGland_paper","figures","supplementaryFigures","bulkRNA","bulkRNA.PCA.variance.pdf"),plot = PCAvariance, 
       width = 2000,height = 500,units = "px")



mir.pca$x %>% as.data.frame()%>% 
  rownames_to_column(var = "sample") %>%
  select(sample,PC1,PC2,PC3,PC4,PC5,PC6,PC7) ->
  sampleInfo.minor.parts.PCA

sampleInfo.minor.parts.PCA %>% inner_join(sampleInfo.QC.merged.minor.parts) %>%
  select(PC1,PC2,,PC3,PC4,PC5,Tissue) %>%
  ggpairs(aes(color = Tissue, alpha = 0.8) ,
          upper = "blank", 
          , lower = list(continuous = "points", combo = "box_no_facet"))+ 
  scale_color_manual(values = BulkPartMinor) +
  scale_fill_manual(values = BulkPartMinor) + theme_light()   ->
  RNAqual2.check     


ggsave(here("MinorGland_paper","figures","supplementaryFigures","bulkRNA","bulkRNA.PCA.plot.pdf")
       ,plot = RNAqual2.check, 
       width = 2000,height = 2000,units = "px")



sampleInfo.minor.parts.PCA %>% inner_join(sampleInfo.QC.merged.minor.parts) %>%
  select(PC1,PC2,Tissue) %>%
  ggpairs(aes(color = Tissue, alpha = 0.8) ,
          upper = "blank", 
          , lower = list(continuous = "points", combo = "box_no_facet"))+ 
  scale_color_manual(values = BulkPartMinor) +
  scale_fill_manual(values = BulkPartMinor) + theme_light()   ->
  RNAqual3.check     



mir.pca$rotation %>%
  as.data.frame() %>%
  rownames_to_column(var = "gene_id") %>%
  select(gene_id,PC1,PC2) %>%
  inner_join(geneInfo.QC) %>%
  select(geneNameProtein,PC1,PC2)->
  minor.parts.PCA.loading


mir.pca$x %>%
  as.data.frame() %>%
  rownames_to_column(var = "sample") %>%
  select(sample,PC1,PC2) %>%
  inner_join(sampleInfo.QC.merged.minor.parts) ->
  minor.parts.PCA


```


```{r, message=FALSE, warning=FALSE }
normExpression.minor.parts[,as.character(MinorSampleInfo.QC$sample)] ->normExpression.minor.parts
normExpression.minor.parts[,1:13] -> n2

row_vars <- apply(n2, 1, var)

top_n <- 2000
top_var_rows <- n2[order(row_vars, decreasing = TRUE)[1:top_n], ]


mir.pca <- prcomp(t(top_var_rows), center = TRUE, scale = TRUE) 
e.var = (mir.pca[['sdev']]^2 / sum(mir.pca[['sdev']]^2))
e.var = as.data.frame( e.var )
e.var$PC = as.factor(1:nrow(e.var)) 
e.var$Variance = e.var$e.var*100
e.var$CumulativeVariance = cumsum(e.var$Variance)
qplot(PC, Variance, data=e.var, geom = c("point")
      , ylim = c(0,max(e.var$Variance)+5) 
      ,ylab = "Variance (%)")+ theme_light() -> variance

#PCAvariance[["minor.part"]] = e.var


mir.pca$x %>% as.data.frame()%>% 
  rownames_to_column(var = "sample") %>%
  select(sample,PC1,PC2,PC3,PC4,PC5,PC6,PC7) ->
  sampleInfo.minor.parts.PCA

sampleInfo.minor.parts.PCA %>% inner_join(sampleInfo.QC.merged.minor.parts) %>%
  select(PC1,PC2,Tissue) %>%
  ggpairs(aes(color = Tissue, alpha = 0.8) ,
          upper = "blank", 
          , lower = list(continuous = "points", combo = "box_no_facet"))+ 
  scale_color_manual(values = BulkPartMinor) +
  scale_fill_manual(values = BulkPartMinor) + theme_light()   ->
  RNAqual2.check   


sampleInfo.minor.parts.PCA %>% inner_join(sampleInfo.QC.merged.minor.parts) %>%
  select(PC1,PC2,Tissue) %>%
  ggplot(aes(x =Tissue ,y = PC1,  fill = Tissue))+
  geom_boxplot(alpha =0.8) +
  scale_color_manual(values = BulkPartMinor) +
  scale_fill_manual(values = BulkPartMinor)  + theme_light()  + theme(legend.position = "none")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))  ->
  FigurePCA.minor.parts



variance+FigurePCA.minor.parts ->Figure2.E.PCA.minor.parts

ggsave(here("MinorGland_paper","figures","figure2","Figure2.E.PCA.minor.parts.pdf")
       ,plot = Figure2.E.PCA.minor.parts, 
       width = 160,height = 80,units = "mm")


```



#### DEseq2

##### Compare Minor parts

###### Duct 
```{r , message=FALSE, warning=FALSE}



MinorSampleInfo.QC


MinorSampleInfo.QC  ->
  MinorSampleInfo.QC.MiAG

geneInfo.QC %>% inner_join(minorProteomics.expressed)-> geneInfo.QC.minor.prot



MinorSampleInfo.QC.MiAG %>%
  mutate(Tissue = as.character(Tissue)) %>%
  mutate(sample = as.character(sample)) -> 
  MinorSampleInfo.QC.MiAG


dds.MiAG = getDEseq2Object(count.df = MinorCounts,
                 md =MinorSampleInfo.QC.MiAG,
                 gi =geneInfo.QC.minor.prot, 
                 design =  "Tissue" )
dds.MiAG = DESeq(dds.MiAG)


resDAp <-results(dds.MiAG, 
               contrast =
                 list(c("designSac_distal_Duct"),c("designSac_distal")), 
               listValues = c(1,-1))
as.data.frame(resDAp) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Duct") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.Dap.DEseq2



geneInfo.Dap.DEseq2 %>%
  filter(padj < 0.05 &log2FoldChange >log2(5) ) ->
  DEseq.minor.parts.duct.results



```

###### Exluding the duct samples

```{r , message=FALSE, warning=FALSE}



MinorSampleInfo.QC


MinorSampleInfo.QC %>% 
  filter(Tissue != "Sac_distal_Duct")   ->
  MinorSampleInfo.QC.MiAG

geneInfo.QC %>% inner_join(minorProteomics.expressed)-> geneInfo.QC.minor.prot



MinorSampleInfo.QC.MiAG %>%
  mutate(Tissue = as.character(Tissue)) %>%
  mutate(sample = as.character(sample)) -> 
  MinorSampleInfo.QC.MiAG


dds.MiAG = getDEseq2Object(count.df = MinorCounts,
                 md =MinorSampleInfo.QC.MiAG,
                 gi =geneInfo.QC.minor.prot, 
                 design =  "Tissue" )
dds.MiAG = DESeq(dds.MiAG)

#Sac_center_vs_Tail_proximal
resScTp <-results(dds.MiAG, 
               contrast =
                 list(c("designSac_center"),c("designTail_proximal")), 
               listValues = c(1,-1))

as.data.frame(resScTp) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Sac_center") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.ScTp.DEseq2


#ZoneC
resSdTp <-results(dds.MiAG, 
               contrast =
                 list(c("designSac_distal"),c("designTail_proximal")), 
               listValues = c(1,-1))
as.data.frame(resSdTp) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Sac_distal") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.SdTp.DEseq2



resSpAp <-results(dds.MiAG, 
               contrast =
                 list(c("designSac_proximal"),c("designTail_proximal")), 
               listValues = c(1,-1))
as.data.frame(resSpAp) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Sac_proximal") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.SpAp.DEseq2

bind_rows(geneInfo.SpAp.DEseq2,geneInfo.ScTp.DEseq2,geneInfo.SdTp.DEseq2)->
  DEseq.minor.parts.results



DEseq.minor.parts.results %>% select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  pivot_wider(names_from = "Tissue", values_from = c("baseMean","log2FoldChange","padj")) ->
  DEseq.minor.parts.results.wide 




```

```{r , message=FALSE, warning=FALSE}


DEseq.minor.parts.results %>%
  group_by(gene_id) %>%
  summarise(minPadj = min(padj), maxFoldChange = max(abs(log2FoldChange))) %>%
  arrange(minPadj) %>%
  inner_join(DEseq.minor.parts.results) %>%
  filter(minPadj < 0.05 &maxFoldChange >log2(5) ) ->
  DEseq.minor.parts.A.results.pruned

DEseq.minor.parts.A.results.pruned %>%
  select(geneName,signalP,log2FoldChange,baseMean,Tissue) %>%
  pivot_wider(values_from = "log2FoldChange", names_from = "Tissue") ->
  DEseq.parts.A.results.pruned.wide


DEseq.parts.A.results.pruned.wide%>% filter(signalP == "Yes") %>%
  inner_join(geneInfo)  %>%
  select(geneNameProtein,Sac_proximal,Sac_center,Sac_distal) %>%
  column_to_rownames("geneNameProtein") ->
  DEseq.parts.A.results.pruned.signalP.wide


DEseq.minor.parts.A.results.pruned %>% inner_join(geneInfo) %>%
  distinct(geneNameProtein, Term) %>%
  group_by(Term) %>% summarise(n = n())
  


DEseq.minor.parts.A.results.pruned %>%
  filter(log2FoldChange > 0) %>%
  group_by (geneName) %>%
  arrange(-log2FoldChange) %>%
  slice_head(n = 1)  ->
  geneLocations

DEseq.minor.parts.A.results.pruned %>%
  filter(abs(log2FoldChange) == maxFoldChange) %>%
  mutate(Tissue = replace(Tissue, log2FoldChange < 0,"Tail_proximal"))  %>%
  arrange(Tissue,log2FoldChange)-> DEseq.minor.parts.A.results.pruned.location


DEseq.minor.parts.A.results.pruned.location %>%
  filter(Tissue == "Tail_proximal") %>%
  select(geneNameProtein) %>% 
  inner_join(geneLocations) %>%
  select(geneNameProtein,Tissue)-> DEseq.minor.parts.A.results.pruned.location.changed

DEseq.minor.parts.A.results.pruned.location  %>%
  select(geneNameProtein,Tissue) %>%
  filter(!geneNameProtein %in% DEseq.minor.parts.A.results.pruned.location.changed$geneNameProtein) %>%
  bind_rows(DEseq.minor.parts.A.results.pruned.location.changed) %>%
  arrange(Tissue) %>%
  ungroup() ->
  DEseq.minor.parts.A.results.pruned.location



DEseq.minor.parts.duct.results %>%
  select(geneNameProtein,Tissue) %>%
  bind_rows(DEseq.minor.parts.A.results.pruned.location) ->
  DEseq.minor.parts.location


data.frame(mir.pca$rotation) %>%
  rownames_to_column(var = "gene_id") %>%
  inner_join(geneInfo) %>%
  inner_join(DEseq.minor.parts.A.results.pruned.location) %>%
  ggplot(aes(x = PC1, y = PC2, color = Tissue, label = geneNameProtein)) +
  geom_point() + 
  theme_light() +
  theme(legend.position = "bottom") -> DE.distribution


data.frame(mir.pca$rotation) %>%
  rownames_to_column(var = "gene_id") %>%
  inner_join(geneInfo) %>%
  select(geneNameProtein, PC1,PC2) ->
  PCA.rotation.data
  









```




##### Compare glands

```{r , message=FALSE, warning=FALSE}

geneInfo.QC %>% inner_join(minorProteomics.expressed)-> geneInfo.QC.minor.prot


sampleInfo.QC.merged %>%
  mutate(Tissue = as.character(Tissue)) %>%
  mutate(sample = as.character(sample)) %>%
  filter(Tissue %in% c("Head","Major","Minor"))-> 
  sampleInfo.QC.merged.DE



dds.Gland = getDEseq2Object(count.df = bulk.data.merged.QC.long,
                 md =sampleInfo.QC.merged.DE,
                 gi =geneInfo.QC.minor.prot, 
                 design =  "Tissue" )
dds.Gland = DESeq(dds.Gland)


resultsNames(dds.Gland)
#Sac_center_vs_Tail_proximal

resMiHe <-results(dds.Gland, 
               contrast =
                 list(c("designMinor"),c("designHead")), 
               listValues = c(1,-1))

as.data.frame(resMiHe) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Head") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.MiHe.DEseq2


resMiMa <-results(dds.Gland, 
               contrast =
                 list(c("designMinor"),c("designMajor")), 
               listValues = c(1,-1))

as.data.frame(resMiMa) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Major") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.MiMa.DEseq2



resMaHe <-results(dds.Gland, 
               contrast =
                 list(c("designMajor"),c("designHead")), 
               listValues = c(1,-1))

as.data.frame(resMaHe) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Major_Head") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.MaHe.DEseq2


bind_rows(geneInfo.MiMa.DEseq2,geneInfo.MiHe.DEseq2,geneInfo.MaHe.DEseq2)->
  DEseq.minor.gland.results

DEseq.minor.gland.results %>%
  group_by(gene_id) %>%
  summarise(minPadj = min(padj), maxFoldChange = max(log2FoldChange)) %>%
  arrange(minPadj) %>%
  inner_join(DEseq.minor.gland.results) %>%
  filter(minPadj < 0.05,maxFoldChange >log2(5) ) %>%
  filter(signalP == "Yes")->
  DEseq.minor.gland.pruned.results


DEseq.minor.gland.results %>%
  filter(geneNameProtein %in%  rownames(DEseq.parts.A.results.pruned.signalP.wide)) %>%
  mutate(sigLevel = "None") %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.1, "**")) %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.05, "***")) %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.0001, "****")) %>%
  mutate(dir = "Up") %>%
  mutate(dir = replace(dir, log2FoldChange < 0, "Down")) %>%
  select(geneNameProtein,Tissue,log2FoldChange,sigLevel,dir)->
  DEseq.minor.gland.HM.results



#DEseq.parts.A.results.pruned.signalP.wide = DEseq.parts.A.results.pruned.signalP.wide[DEseq.minor.gland.results.2$geneNameProtein,] 

#DEseq.parts.A.results.pruned.signalP.wide


geneInfo.MiHe.DEseq2%>%
  filter(padj < 0.05,log2FoldChange >log2(5) ) %>%
  inner_join(geneInfo) %>%
  select(geneNameProtein,Term) -> miHe.final


DEseq.minor.parts.A.results.pruned %>% inner_join(geneInfo) %>%
  distinct(geneNameProtein, Term) %>% inner_join(miHe.final) %>%
  group_by(Term) %>% summarise(n = n())



DEseq.minor.gland.results %>%
  filter(Tissue == "Head")  %>%
  filter(padj < 0.05) %>%
  filter(log2FoldChange >  log2(5)) ->
  DEseq.minor.head.results



DEseq.minor.head.results %>% inner_join(interpro5) %>%
  group_by(interpro_accession,interpro_description) %>%
  summarise(n = n()) %>%
  arrange(-n)
  

GOI_common = c("LASC007034","LASC0011995","lman2l",
        "x975_2","SpiCE-LMa5","AmSp-like1","AmSp-like2",
        "SpiCE-LMa3","SpiCE-LMa1",
        "SpiCE-LMa2","SpiCE-LMa4",
        "LASC009784","spa",
        "LASC007099","LASC017532")
geneInfo %>% 
  left_join(interpro5) %>% 
  filter(geneNameProtein %in%  GOI_common)


geneInfo %>% filter(grepl("SpiCE",geneNameProtein2 )) %>% select(geneNameProtein, geneNameProtein2) %>%
  arrange(geneNameProtein2)

interpro5  %>%
   filter(interpro_accession == "IPR043070") %>%
   group_by(geneName ) %>%
   summarise(n = n())  %>%
  arrange(-n)  -> IPR043070


DEseq.minor.gland.results%>%
  select(gene_id,Tissue,log2FoldChange,padj) %>%
  
  pivot_wider(names_from = "Tissue", values_from = c("log2FoldChange","padj")) ->
  DEseq.minor.glands.results.wide 
  
  
DEseq.minor.glands.results.wide %>%
  rename(log2FoldChange_Major_vs_Head = "log2FoldChange_Major_Head") %>%
  rename(padj_Major_vs_Head = "padj_Major_Head") %>%
  rename(log2FoldChange_Minor_vs_Head = "log2FoldChange_Head") %>%
  rename(padj_Minor_vs_Head = "padj_Head") %>%
  rename(log2FoldChange_Minor_vs_Major = "log2FoldChange_Major") %>%
  rename(padj_Minor_vs_Major = "padj_Major") ->
  DEseq.minor.glands.results.wide
  


```
#####  Common genes
```{r}


geneInfo %>% 
  filter(signalP == "Yes") ->
  signalP

DEseq.minor.gland.results %>%
  filter(Tissue == "Head")  %>%
  filter(padj < 0.05) %>%
  filter(log2FoldChange >  log2(5)) %>%
  select(geneNameProtein,signalP,Term)->
  DEseq.gland.sig


DEseq.minor.parts.results %>% 
  select(-Tissue) %>%
  right_join(DEseq.minor.parts.A.results.pruned.location)%>%
  select(geneNameProtein,signalP,Term,Tissue) %>%
  distinct()->
  DEseq.part.sig

DEseq.part.sig %>%
  arrange(Tissue)

DEseq.part.sig %>%
  filter(geneNameProtein %in% DEseq.gland.sig$geneNameProtein) %>%
  select(geneNameProtein,Tissue,Term)->
  DEseq.common

genesOfInterest = list(signalP$geneNameProtein, DEseq.part.sig$geneNameProtein, DEseq.gland.sig$geneNameProtein) 

#install.packages("ggVennDiagram")  
library(ggVennDiagram)

three_set_venn <- ggVennDiagram(genesOfInterest)
pdf("~/Google Drive/My Drive/Jobb/Minor gland/paper2/figures_new/Figure3/Venn.pdf", width = 2, height = 4)
print(three_set_venn)
dev.off()

DEseq.common %>%
  filter(Term != "RibosomalProteins") %>%
  filter(Term != "NoSignalPeptide") ->
DEseq.common.27

DEseq.common.27 %>%
  arrange(Term,Tissue)


DEseq.minor.gland.results %>%
  filter(Tissue == "Head")  %>%
  arrange(Term,geneNameProtein)


intersect(DEseq.minor.head.results$geneNameProtein,
          DEseq.minor.parts.A.results.pruned.location$geneNameProtein) -> common

intersect(common, signalP$geneNameProtein) -> common.signalP

```


## Supplementary table 
```{r}
# Proteomics
minorProteomics.QC %>% inner_join(geneInfo) %>%
  select(geneNameProtein,geneName,gene_id,  Part,  averageIntensity) %>%
  pivot_wider(names_from = "Part", values_from = "averageIntensity") ->
  proteomics.wide

# RNA parts

left_join(proteomics.wide,MinorCounts.QC.tpm.wide)
  

sampleInfo.All %>% filter(Tissue %in% c("Minor","Major","Head")) -> sampleInfo.paper
sampleInfo.paper %>% inner_join(bulk.data.merged.QC.tpm) %>%
  group_by(gene_id,Tissue) %>%
  summarise(tpm = mean(tpm)) %>%
  pivot_wider(names_from = "Tissue", values_from = "tpm") ->
  TissuesCounts.QC.tpm.wide
 
left_join(proteomics.wide,MinorCounts.QC.tpm.wide) %>%
  left_join(TissuesCounts.QC.tpm.wide) ->
  bulkRNA_expression.proteomics.table


bulkRNA_expression.proteomics.table



bulkRNA_expression.proteomics.table %>% arrange(-Sac_proximal) ->
  bulkRNA_expression.proteomics.table


# DE results 
left_join(bulkRNA_expression.proteomics.table,
          DEseq.minor.parts.results.wide)->
  bulkRNA_expression.proteomics.DE.table


left_join( bulkRNA_expression.proteomics.DE.table,
           DEseq.minor.glands.results.wide) %>%
  select(-baseMean_Sac_proximal,
         -baseMean_Sac_center,
         -baseMean_Sac_distal)%>%
  rename(log2FC_Major_vs_Head = log2FoldChange_Major_vs_Head,
         padj_Major_vs_Head = padj_Major_vs_Head,
         log2FC_Minor_vs_Head = log2FoldChange_Minor_vs_Head,
         padj_Minor_vs_Head = padj_Minor_vs_Head,
         log2FC_Minor_vs_Major = log2FoldChange_Minor_vs_Major,
         padj_Minor_vs_Major = padj_Minor_vs_Major) %>%
  rename(log2FoldChange_Sac_proximal_vs_Tail_proximal = 
           log2FoldChange_Sac_proximal ) %>%
  rename(padj_Sac_proximal_vs_Tail_proximal = padj_Sac_proximal ) %>%
  rename(log2FoldChange_Sac_center_vs_Tail_proximal = 
           log2FoldChange_Sac_center ) %>%
  rename(padj_Sac_center_vs_Tail_proximal = padj_Sac_center ) %>%
  rename(log2FoldChange_Sac_distal_vs_Tail_proximal = 
           log2FoldChange_Sac_distal ) %>%
  rename(padj_Sac_distal_vs_Tail_proximal = padj_Sac_distal ) ->
  supplementaryTable.bulkRNA.proteomics.DE
  

minorProteomics.QC %>% 
  pivot_wider(names_from = Part, values_from = averageIntensity, values_fill = 0) 


supplementaryTable.bulkRNA.proteomics.DE %>% select(geneNameProtein,gene_id,Gland,Dope) ->
  supplemntaryTable.proteomics.GlandDope

supplementaryTable.bulkRNA.proteomics.DE %>% select(geneNameProtein,gene_id,Tail_proximal,Sac_proximal,Sac_center,Sac_distal,log2FoldChange_Sac_proximal_vs_Tail_proximal,padj_Sac_center_vs_Tail_proximal,log2FoldChange_Sac_center_vs_Tail_proximal,padj_Sac_center_vs_Tail_proximal,log2FoldChange_Sac_distal_vs_Tail_proximal,padj_Sac_distal_vs_Tail_proximal)->
  supplemntaryTable.RNAexpression.parts

supplementaryTable.bulkRNA.proteomics.DE %>% select(geneNameProtein,gene_id,Minor,Major,Head,
                                                    log2FC_Minor_vs_Head,padj_Minor_vs_Head,
                                                    log2FC_Minor_vs_Major,padj_Minor_vs_Major,
                                                    log2FC_Major_vs_Head,padj_Major_vs_Head)->
  supplemntaryTable.RNAexpression.glands


# Functional Annotation results
write_csv(interpro5.spacer, 
          here("MinorGland_paper", "results", "supplementaryTables", "suppl.tab.1.spacer_domain_counts.csv"))

# Supplementary Data Tables
write_csv(supplemntaryTable.proteomics.GlandDope, 
          here("MinorGland_paper", "results", "supplementaryTables", "Suppl.Table.2.proteomics.gland.dope.csv"))

write_csv(supplemntaryTable.RNAexpression.glands, 
          here("MinorGland_paper", "results", "supplementaryTables", "Suppl.Table.3.bulkRNA.expression.glands.csv"))

write_csv(supplemntaryTable.RNAexpression.parts, 
          here("MinorGland_paper", "results", "supplementaryTables", "Suppl.Table.4.bulkRNA.expression.parts.csv"))



```


# Figure 2 
## Figure2A 
```{r statisical analysis, message=FALSE, warning=FALSE}



# Proteomics
minorProteomics.QC %>% inner_join(geneInfo) %>%
  group_by(Term, Part) %>%
  summarise(n = n()) %>%
  pivot_wider(values_from = "n", names_from = Part, values_fill = 0) ->
  counts.proteomics

# RNA parts
DEseq.minor.parts.A.results.pruned.location %>%
  inner_join(geneInfo) %>%
  group_by(Term) %>%
  summarise(RNAparts = n())  ->
  counts.RNAparts


# RNA gland vs head

DEseq.minor.head.results %>% inner_join(geneInfo) %>%
  group_by(Term) %>%
  summarise(RNAgland = n())  ->
  counts.gland

DEseq.common %>%
  inner_join(geneInfo) %>%
  group_by(Term) %>%
  summarise(RNAcommon = n())  ->
  counts.common


# Join all counts
counts.proteomics %>% left_join(counts.gland) %>%
  left_join(counts.RNAparts) %>%
  left_join(counts.common)%>%  
  mutate(RNAparts = replace(RNAparts, is.na(RNAparts),0)) %>%
  mutate(RNAcommon =replace(RNAcommon, is.na(RNAcommon),0) )->
  counts.all



geneInfo %>% mutate(Part = "All") %>% group_by(Term, Part) %>%
  summarise(n = n())%>%
  pivot_wider(values_from = "n", names_from = Part, values_fill = 0)  %>%
  left_join(counts.all) ->
  counts

long_data <- counts %>%
  pivot_longer(cols = -c(Term, All), names_to = "Group", values_to = "Count")

total_genes = nrow(geneInfo)
long_data %>%
  group_by(Group) %>%
  summarise(AllGroup = sum(Count)) %>%
  ungroup() %>%
  inner_join(long_data) %>%
  mutate(total_genes = total_genes) %>%
  group_by(Term, Group) %>%
  mutate(
    Not_in_Group = All - Count,
    Total_in_Group = mean(AllGroup), # Sum across all terms in each Group
    Not_in_Term_Group = Total_in_Group - Count,
    Not_in_Term_All = total_genes - All ) ->
  long_data

long_data %>%
  rowwise() %>%
  mutate(test = 
           list(fisher.test(
             matrix(c(Count,
                      Not_in_Group,
                      Not_in_Term_Group,
                      Not_in_Term_All),
                    nrow = 2)
           )
           ),
         Odds_Ratio = test$estimate,
         P_Value = test$p.value
  ) %>%
  select(-test) %>%
  unnest(c(Odds_Ratio, P_Value)) ->
  geneEnrichment.results

geneEnrichment.results %>%
  select(Term,All, Group, Count) %>%
  pivot_wider(names_from = "Group", values_from = "Count") ->
  counts

geneEnrichment.results %>%
  select(Term, Group, P_Value) %>%
  pivot_wider(names_from = "Group", values_from = "P_Value") ->
  pvalues

geneEnrichment.results %>%
  select(Term, Group, Count,P_Value,Odds_Ratio) 
  
geneEnrichment.results %>%
  select(Term,Group,Count) %>%
  pivot_wider(names_from = "Group", values_from = "Count") ->
counts  


counts %>% select(Term,Gland,Dope) -> counts.protein
class(counts)
table <- tableGrob(counts.protein, 
                   rows = NULL,
                   theme = ttheme_default(core = list(fg_params = list(cex = 0.7))
                                          )
)
Figure2A <- ggplot() +
  annotation_custom(table) +
  theme_void()
  
Figure2A 

ggsave(filename = here("MinorGland_paper", "figures", "figure2", "figure2A.pdf"),
       plot = Figure2A,
       create.dir = TRUE, 
       device = "pdf",
       width = 183 ,  # Convert mm to inches
       height = 120 ,  # Adjust height as needed
       units = "mm",
       dpi = 300  # High resolution for publication
       )

Figure2A
```



###  Figure2B

```{r statisical analysis, message=FALSE, warning=FALSE}



minorProteomics.QC %>%
  inner_join(geneInfo)%>%
  group_by(Part,Term) %>%
  summarise(averageIntensity = sum(averageIntensity)) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope"))) %>%
  ggplot(aes(x = Part,y = averageIntensity,  fill = Term)) +
  geom_bar( stat = "identity", position = "fill" ) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Tissue",
    y = "Normalised proteinLevels (fraction)"
  ) +
  scale_fill_brewer(palette = "Set2")  -> 
  fig2B


ggsave(filename = here("MinorGland_paper", "figures", "figure2", "figure2B.pdf"),
       plot = fig2B,
       create.dir = TRUE, 
       device = "pdf",
       width = 183 ,  # Convert mm to inches
       height = 120 ,  # Adjust height as needed
       units = "mm",
       dpi = 300  # High resolution for publication
       )

fig2B

```

### Figure 2C

```{r , message=FALSE, warning=FALSE}


# Showing the entire gland  

#sampleInfo.QC.merged %>% filter(Tissue == "Minor") %>%
#  inner_join(bulk.data.merged.QC.tpm) %>%
#  group_by(gene_id) %>%
#  summarise(tpm = mean(tpm))%>%
#  mutate(Tissue = "Gland") %>%
#  inner_join(geneInfo) %>% group_by(Tissue,Term) %>%
#  summarise(tpmTot = sum(tpm))->
#  MinorGland.QC.tpm.long.summary  

#MinorGland.QC.tpm.long.summary %>%
#  mutate(Tissue = factor(Tissue, levels = TissueOrder)) %>%
#  ggplot(aes(x = Tissue, y = tpmTot, fill = Term)) + geom_bar(stat = "identity") +
#  theme_light() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 




MinorCounts.QC.tpm.wide %>%
  pivot_longer(cols = -gene_id, names_to = "Tissue", values_to = "tpm")%>%
  inner_join(geneInfo) %>% group_by(Tissue,Term) %>%
  summarise(tpmTot = sum(tpm))->
  MinorParts.QC.tpm.long.summary 

MinorParts.QC.tpm.long.summary %>%
  mutate(Tissue = factor(Tissue, levels = TissueOrder)) %>%
  ggplot(aes(x = Tissue, y = tpmTot, fill = Term)) +
  geom_bar(stat = "identity") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Tissue",
    y = "Normalised expression (tpm)"
  ) +
  scale_fill_brewer(palette = "Set2") -> 
  fig2C

ggsave(filename = here("MinorGland_paper", "figures", "figure2", "figure2C.pdf"),
       plot = fig2C,
       create.dir = TRUE, 
       device = "pdf",
       width = 183 ,  # Convert mm to inches
       height = 120 ,  # Adjust height as needed
       units = "mm",
       dpi = 300  # High resolution for publication
       )


fig2C
```


## Figure2E  PCA

PCA related data

```{r}

# Save previous plot
ggsave(filename = here("MinorGland_paper", "figures", "figure2", "figure2E.pdf"),
       plot = FigurePCA.minor.parts, 
       width = 400,height = 800,units = "px")
  


```

## Figure 2F
```{r}
counts %>% select(Term,RNAgland,RNAparts,RNAcommon) -> counts.protein

table <- tableGrob(counts.protein, 
                   rows = NULL,
                   theme = ttheme_default(core = list(fg_params = list(cex = 0.7))
                                          )
)

Figure2F <- ggplot() +
  annotation_custom(table) +
  theme_void()
  
Figure2F 


```

## Figure2H

```{r , message=FALSE, warning=FALSE}


MinorCounts.QC.tpm.wide %>%
  pivot_longer(cols = -gene_id, names_to = "Tissue", values_to = "tpm")%>%
  inner_join(geneInfo) %>% 
  filter(geneNameProtein %in% DEseq.common$geneNameProtein  )%>%
  group_by(Tissue,Term) %>%
  summarise(tpm = sum(tpm))->
  MinorParts.QC.tpm.long.summary 

MinorParts.QC.tpm.long.summary %>%
  mutate(Tissue = factor(Tissue, levels = TissueOrder)) %>%
  ggplot(aes(x = Tissue, y = tpm, fill = Term)) +
  geom_bar(stat = "identity") +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Tissue",
    y = "Normalised expression (tpm)"
  ) +
  scale_fill_manual(values = termColors) -> 
  fig2H.together


fig2H.together

ggsave(filename = here("MinorGland_paper", "figures", "figure2", "figure2H.pdf"),
       plot = fig2H.together,
       create.dir = TRUE, 
       device = "pdf",
       width = 183 ,  # Convert mm to inches
       height = 120 ,  # Adjust height as needed
       units = "mm",
       dpi = 300  # High resolution for publication
)


```

##  Figure2G

```{r statisical analysis, message=FALSE, warning=FALSE}

DEseq.gland.sig
DEseq.part.sig
DEseq.common

minorProteomics.QC %>%
  inner_join(geneInfo)%>%
  inner_join(DEseq.gland.sig) %>%
  group_by(Part,Term) %>%
  summarise(averageIntensity = sum(averageIntensity)) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope"))) %>%
  ggplot(aes(x = Part,y = averageIntensity,  fill = Term)) +
  geom_bar( stat = "identity", position = "fill" ) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Tissue",
    y = "Normalised expression (tpm)"
  ) +
  scale_fill_manual(values = termColors) -> 
  fig2G.gland

minorProteomics.QC %>%
  inner_join(geneInfo)%>%
  inner_join(DEseq.part.sig) %>%
  group_by(Part,Term) %>%
  summarise(averageIntensity = sum(averageIntensity)) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope"))) %>%
  ggplot(aes(x = Part,y = averageIntensity,  fill = Term)) +
  geom_bar( stat = "identity", position = "fill" ) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Tissue",
    y = "Normalised expression (tpm)"
  ) +
  scale_fill_manual(values = termColors) -> 
  fig2G.part

minorProteomics.QC %>%
  inner_join(geneInfo)%>%
  inner_join(DEseq.common) %>%
  group_by(Part,Term) %>%
  summarise(averageIntensity = sum(averageIntensity)) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope"))) %>%
  ggplot(aes(x = Part,y = averageIntensity,  fill = Term)) +
  geom_bar( stat = "identity", position = "fill" ) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Tissue",
    y = "Normalised expression (tpm)"
  ) +
  scale_fill_manual(values = termColors) -> 
  fig2G.common




#ggsave(filename = "~/Google Drive/My Drive/Minor gland/paper2/figures_new/figure2/figure2ABC.pdf",
#       plot = fig2BC,
#       create.dir = TRUE, 
#       device = "pdf",
#       width = 183 ,  # Convert mm to inches
#       height = 120 ,  # Adjust height as needed
#       units = "mm",
#       dpi = 300  # High resolution for publication
#       )


(fig2G.gland+ fig2G.part +fig2G.common+fig2G.together )+
  plot_layout(guides = "collect", nrow = 1,widths = c(1,1,1,2)) &
  theme(legend.position = "none")  ->
  fig2G.GH

ggsave(filename = "~/Google Drive/My Drive/Minor gland/paper2/figures_new/figure2/figure2GH.pdf",
       plot = fig2G.GH,
       create.dir = TRUE, 
       device = "pdf",
       width = 160 ,  # Convert mm to inches
       height = 80 ,  # Adjust height as needed
       units = "mm",
       dpi = 300  # High resolution for publication
       )
```



### Distribution of hits 
```{r}


geneInfo %>%
  distinct(geneNameProtein, Term)%>%
  mutate(type = "Annotation") -> 
  geneInfo.geneInfo

minorProteomics.expressed %>% inner_join(geneInfo) %>%
  distinct(geneNameProtein, Term)%>%
  mutate(type = "Protein") -> 
  proteinPresent.geneInfo
  

DEseq.minor.head.results  %>%
  distinct(geneNameProtein, Term) %>%
  mutate(type = "DE_Tissue") -> 
  DEseq.minor.head.pruned.geneInfo


DEseq.minor.parts.A.results.pruned %>%
  distinct(geneNameProtein, Term) %>%
  mutate(type = "DE_Part")  -> 
  DEseq.minor.parts.A.results.pruned.geneInfo





inner_join(DEseq.minor.head.pruned.geneInfo[,c("geneNameProtein","Term")],
           DEseq.minor.parts.A.results.pruned.geneInfo[,c("geneNameProtein","Term")])  %>%
  mutate(type = "Both")  -> DEseq.minor.both.geneInfo


  


DEseq.minor.both.geneInfo %>%
  filter(Term != "NoSignalPeptide") %>%
  filter(Term != "RibosomalProteins") %>%
  mutate(type = "Final") ->
  DEseq.minor.final.geneInfo

bind_rows(geneInfo.geneInfo,proteinPresent.geneInfo,
          DEseq.minor.head.pruned.geneInfo,
          DEseq.minor.parts.A.results.pruned.geneInfo,
          DEseq.minor.both.geneInfo,
          DEseq.minor.final.geneInfo)  %>%
 mutate(type = factor(type, levels =c("Annotation","Protein","DE_Tissue","DE_Part", "Both","Final") )) ->
  all.geneInfo


all.geneInfo %>% group_by(Term, type) %>%
  summarise(n = n()) -> all.geneInfo.summary


all.geneInfo.summary %>% group_by(type) %>%
  summarise(total = sum(n))  %>% inner_join(all.geneInfo.summary ) %>%
  mutate(fraction = n /total) ->
  all.geneInfo.summary


all.geneInfo.summary %>% 
  ggplot(aes( x = type, y = fraction, fill = Term))+ 
  geom_bar(stat = "identity")+
  theme_light()+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none" ) + 
  scale_fill_brewer(palette = "Set2")+
  labs(
    title = "Distribution of annotation",
    x = "Selection",
    y = "Fraction"
  ) ->
  distribution

all.geneInfo.summary %>%
  distinct(type,total) %>%
  ggplot(aes(x = type, y = total )) + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = total), hjust = 0.5, vjust = 2, size = 3, color = "white")+
  theme_light()+
  scale_y_log10() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "none" 
  ) +
  labs(
    title = "Nr of genes",
    y = "log10(nrOfGenes)"
  )  -> count 

  
count/distribution  -> fig2A


ggsave("figures/figure2/geneDistribution.pdf",height = 4,width = 4,plot = fig2A)


DEseq.minor.both.geneInfo %>% group_by(Term)%>%
  summarise(n = n())



```





```{r}



DEseq.minor.parts.A.results.pruned %>% 
  distinct(geneNameProtein, baseMean)%>% 
  filter(geneNameProtein %in% rownames(DEseq.parts.A.results.pruned.signalP.wide)) %>%
  select(geneNameProtein,baseMean) %>%
  column_to_rownames("geneNameProtein") ->
  baseMean.HM

DEseq.minor.parts.A.results.pruned %>% 
  filter(geneNameProtein %in% rownames(DEseq.parts.A.results.pruned.signalP.wide)) %>%
  select(geneNameProtein,Tissue, log2FoldChange, padj,baseMean) %>%
  mutate(expLevel = baseMean*(2^log2FoldChange))  %>%     select(geneNameProtein,Tissue,expLevel,log2FoldChange,padj) %>%
  mutate(sigLevel = "None") %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.1, "**")) %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.05, "***")) %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.0001, "****")) ->
    DEseq.parts.A.results.pruned.sig
    
 DEseq.minor.parts.A.results.pruned %>% 
  distinct(geneNameProtein, baseMean)%>% 
  filter(geneNameProtein %in% rownames(DEseq.parts.A.results.pruned.signalP.wide)) %>%
  select(geneNameProtein,baseMean) %>%
   rename(expLevel = "baseMean" ) %>%
   mutate(Tissue = "Tail_proximal") %>%
   mutate(log2FoldChange = 0) %>%
   mutate(padj = 1) %>%
   mutate(sigLevel = "None") ->
     DEseq.parts.Tail_proximal
     

 bind_rows(DEseq.parts.A.results.pruned.sig,DEseq.parts.Tail_proximal) %>%
   arrange(geneNameProtein) ->DEseq.parts.A.results.pruned.sig

  
 DEseq.parts.A.results.pruned.sig %>%
  mutate(log10expLevel = log10(expLevel)) ->
  DEseq.parts.A.results.pruned.sig.2


 DEseq.parts.A.results.pruned.sig.2 %>% distinct(geneNameProtein)

 
```

```{r}



library(cluster)
library(factoextra)

DEseq.parts.A.results.pruned.sig.2 %>%
  select(geneNameProtein,Tissue,log2FoldChange) %>%
  pivot_wider(names_from = "Tissue", values_from = "log2FoldChange") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> mat

m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

df = mat
ac <- function(x) {
  agnes(df, method = x)$ac
}
sapply(m, ac)
clust <- agnes(df, method = "ward")
pltree(clust, cex = 0.6, hang = -1, main = "Dendrogram") 
gap_stat <- clusGap(df, FUN = hcut, nstart = 25, K.max = 6, B = 50 )
#produce plot of clusters vs. gap statistic

gap_stat_plot <- fviz_gap_stat(gap_stat) +
  ggtitle("Gap Statistic Plot") +
  xlab("Number of Clusters") +
  ylab("Gap Statistic") +
  theme_minimal()

d <- dist(df, method = "euclidean")

#perform hierarchical clustering using Ward's method
final_clust <- hclust(d, method = "ward.D2" )
#cut the dendrogram into 3 clusters
plot(final_clust, main = "Hierarchical Clustering Dendrogram", xlab = "Samples", ylab = "Height")

genes = data.frame(geneName = names(cutree(final_clust, 3)) ,
                   hierarchicalCluster = paste("c",cutree(final_clust, 3),sep ="_")) %>% 
  mutate(hierarchicalCluster2 = hierarchicalCluster)%>%
#  mutate(hierarchicalCluster2 = gsub("c_4","tail_bulk_2",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_3","tail_bulk_1",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_2","sac_bulk_2",hierarchicalCluster2)) %>%
  mutate(hierarchicalCluster2 = gsub("c_1","sac_bulk_1",hierarchicalCluster2)) %>%
#  mutate(hierarchicalCluster2 = gsub("c_5","tail_bulk_1",hierarchicalCluster2)) %>%
  arrange(hierarchicalCluster2)



```



# Figure 3

Keeping only the genes that are in the DEseq parts A results and have a signal peptide

```{r}
DEseq.parts.A.results.pruned.sig.2 %>%
  filter(geneNameProtein %in% DEseq.common.27$geneNameProtein ) ->
  
  
  DEseq.parts.A.results.pruned.sig.2



```

### RNA expression 

#### Fold change 
```{r}
library(circlize)
library(ComplexHeatmap)


# Sample Data
DEseq.parts.A.results.pruned.sig.2 %>%
  filter(Tissue != "Tail_proximal") %>%
  select(geneNameProtein,Tissue,log2FoldChange) %>%
  pivot_wider(names_from = "Tissue", values_from = "log2FoldChange") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> mat_DEseq

MinorCounts.QC.tpm %>% inner_join(MinorSampleInfo.QC.MiAG)  %>%
  select(geneNameProtein,Tissue,sample,tpm) %>%
  filter(geneNameProtein %in% rownames(mat_DEseq)) %>%
  group_by(geneNameProtein,Tissue) %>%
  summarise(tpm = mean(tpm)) %>%
  mutate(log10tpm = log10(tpm)) %>%
  select(-tpm) %>%
  pivot_wider(names_from = "Tissue" , values_from = "log10tpm") %>%
  column_to_rownames("geneNameProtein") %>%
   as.matrix()-> 
  mat_DEseq_exp




DEseq.parts.A.results.pruned.sig.2 %>%
  select(geneNameProtein,Tissue,sigLevel) %>%
  pivot_wider(names_from = "Tissue", values_from = "sigLevel") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> border_data

DEseq.parts.A.results.pruned.sig.2 %>%
  mutate(diff = "Up") %>%
  mutate(diff  = replace(diff, log2FoldChange <0,"Down")) %>%
  select(geneNameProtein,Tissue,diff) %>%
  pivot_wider(names_from = "Tissue", values_from = "diff") %>%
  column_to_rownames("geneNameProtein") %>%
  as.matrix() -> border_color



# Color palettes
min_val <- min(mat_DEseq_exp)
max_val <- max(mat_DEseq_exp)
col_fun <- colorRamp2(c(min_val, 1 ,max_val), c("blue", "white", "red")) # Corrected color mapping
border_colors <- c("Up" = "red", "Down" = "blue","None"="black")
border_styles <- c("None" = "blank","**" = "dotted", "***" = "dashed", "****" = "solid")

# Create the heatmap with cell gap


distance_matrix <- dist(mat_DEseq)

# Perform hierarchical clustering
hc <- hclust(distance_matrix)

# Convert hclust to a dendrogram
row_dend <- as.dendrogram(hc)


clusters <- cutree(hc, k = 3)



# Create the heatmap
mat_DEseq_exp = mat_DEseq_exp[rownames(mat_DEseq),
                              c("Tail_proximal","Sac_proximal","Sac_center","Sac_distal")]
border_data = border_data[rownames(mat_DEseq),
                          c("Tail_proximal","Sac_proximal","Sac_center","Sac_distal")]
border_color = border_color[rownames(mat_DEseq),
                          c("Tail_proximal","Sac_proximal","Sac_center","Sac_distal")]
Heatmap(mat_DEseq_exp,
        col = col_fun,
        cluster_rows = row_dend, # Add the row dendrogram
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          border_lty <- border_styles[border_data[i, j]]
          border_col <- border_colors[border_color[i, j]]
          grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                    gp = gpar(col = border_col, lty = border_lty, lwd = 2), just = "center")
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                              gp = gpar(fill = col_fun(mat_DEseq_exp[i,j]), col = NA), just = "center") # Remove black outline
 #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "DE",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
      cluster_columns = FALSE,split = 3) ->
  HM1

HM1

ordered_genes <- rownames(mat_DEseq)[hc$order]
ordered_genes

data.frame(clusters) %>%
  rownames_to_column(var = "geneNameProtein") %>%
  inner_join(DEseq.parts.A.results.pruned.sig.2) %>%
  mutate(Tissue = factor(Tissue, level = TissueOrder)) %>%
  mutate(cluster2 = "Sac strong") %>%
  mutate(cluster2 = replace(cluster2,clusters == 2,"Sac weak")) %>%
  mutate(cluster2 = replace(cluster2,clusters == 3,"Tail strong")) ->
  minor.bulk.final.cluster.data

minor.bulk.final.cluster.data%>%
  ggplot( aes(y = log2FoldChange, fill = as.factor(cluster2), x = Tissue)) + geom_boxplot()+
  theme_light()+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_fill_manual(values = BulkZones)
  ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/figures_new/Figure3/foldChange.pdf",height = 2, width = 4)

  
 

```
#### Whole tissue

```{r}


bulk.data.merged.QC.tpm %>%
  inner_join(geneInfo) %>%
  filter(geneNameProtein %in% rownames(mat_DEseq)) %>%
  inner_join(sampleInfo.QC.merged)%>%
  filter(Tissue %in% c("Minor","Major","Head")) %>%
  group_by(geneNameProtein,Tissue) %>%
  summarise(mean = mean(tpm))  %>%
  pivot_wider(names_from = "Tissue", values_from = "mean") %>%
  column_to_rownames("geneNameProtein") %>%
  select(Head,Major,Minor) %>%
  as.matrix() -> mat_tissue

mat_tissue = log10(mat_tissue)
mat_tissue = mat_tissue[rownames(mat_DEseq),c("Minor","Major","Head")]
border_colors <- c("Up" = "blue", "Down" = "red","None"="black")


DEseq.minor.gland.HM.results %>%
  select(geneNameProtein,Tissue,sigLevel) %>%
  pivot_wider(names_from = "Tissue", values_from = "sigLevel") %>%
  column_to_rownames("geneNameProtein") %>%
  mutate(Minor = "None") %>%
  as.matrix() -> border_data

border_data = border_data[rownames(mat_DEseq),c("Minor","Major","Head")]


DEseq.minor.gland.HM.results %>%
  select(geneNameProtein,Tissue,dir) %>%
  pivot_wider(names_from = "Tissue", values_from = "dir") %>%
  column_to_rownames("geneNameProtein") %>%
    mutate(Minor = "None") %>%
  as.matrix() -> border_dir
border_dir = border_dir[rownames(mat_DEseq),c("Minor","Major","Head")]



# Create the heatmap


Heatmap(mat_tissue,
        col = col_fun,
        cluster_rows = row_dend, # Add the row dendrogram
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          border_lty <- border_styles[border_data[i, j]]
          border_col <- border_colors[border_dir[i, j]]
          grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                    gp = gpar(col = border_col, lty = border_lty, lwd = 2), just = "center")
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                              gp = gpar(fill = col_fun(mat_tissue[i,j]), col = NA), just = "center") # Remove black outline
 #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "DE",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
      cluster_columns = FALSE,split = 3) ->
  HM1_tissue

HM1_tissue



row_dend$labels = ordered_genes


```


### Protein expression


```{r}

minorProteomics.QC %>%
  inner_join(geneInfo) %>%
  filter(geneNameProtein %in% rownames(mat_DEseq)) %>%
  select(geneNameProtein,Part,averageIntensity) %>%
  pivot_wider(names_from = "Part", values_from = "averageIntensity")  ->
  minorProteomics.wide

minorProteomics.wide  %>%
  filter(geneNameProtein %in% rownames(mat_DEseq)) %>%
  select(geneNameProtein,Gland,Dope) %>%
  column_to_rownames("geneNameProtein") ->
  prot.express.HM

prot.express.HM = prot.express.HM[rownames(mat_DEseq),]

prot.express.HM = log10(prot.express.HM)
# Color palettes
col_fun_Protein <- colorRamp2(c(min(prot.express.HM, na.rm = TRUE),
                        -3 ,
                        max(prot.express.HM, na.rm = TRUE)),
                      c("blue", "white", "red")) # Corrected color mapping
border_colors <- c("None" = "lightgrey", "***" = "darkgrey", "****" = "black")
border_styles <- c("None" = "blank","**" = "dotted", "***" = "dashed", "****" = "solid")


# Create the heatmap


prot.express.HM = prot.express.HM[rownames(mat_DEseq),]

Heatmap(prot.express.HM,
        col = col_fun_Protein,
        na_col = NA,
        rect_gp = gpar(type = "none"), #Important to remove the default drawing of the cells
        cell_fun = function(j, i, x, y, width, height, fill) {
          if (!is.na(prot.express.HM[i, j])) {
            grid.rect(x = x, y = y, width = width - unit(2, "mm"), height = height - unit(2, "mm"),
                    gp = gpar(col = "black",  lwd = 2), just = "center")
          }
          grid.rect(x = x, y = y, width = width - unit(2.5, "mm"), height = height - unit(2.5, "mm"), # Cell gap
                              gp = gpar(fill = col_fun_Protein(prot.express.HM[i,j]), col = NA), just = "center") # Remove black outline
 #Fill the cells with the color gradient
        },
        row_names_side = "right",
        column_names_side = "bottom",
        name = "Protein expression",
        row_gap = unit(0.5, "mm"),
        column_gap = unit(0.5,"mm"),
        cluster_columns = FALSE) ->
  HM2




HM1+HM1_tissue+HM2 -> HM_all


HM_all
library(ComplexHeatmap)
library(grid)

# Draw heatmap in rotated grid


pdf("~/Google Drive/My Drive/Jobb/Minor gland/paper2/figures_new/Figure3/Figure3.heatmap.pdf", width = 10, height = 10)
HM_all
dev.off()

```






# Compare protein and RNA levels

```{r}
    

minorProteomics.QC %>%
  inner_join(geneInfo) %>%
  select(geneNameProtein,Part,averageIntensity) %>%
  pivot_wider(names_from = "Part", values_from = "averageIntensity")  ->
  minorProteomics.wide.all


bulk.data.merged.QC.tpm %>% inner_join(sampleInfo.QC.merged.DE) %>%
  filter(Tissue == "Minor" & part == "Gland") %>%
  group_by(gene_id) %>%
  summarise(tpm = mean(tpm)) %>% inner_join(geneInfo) %>%
 left_join(minorProteomics.wide.all) %>%
   mutate(Term = factor(Term, levels = c("MiSp","AmSp-like","Spidroins","SpiCE","SignalPeptide","RibosomalProteins","NoSignalPeptide")) ) ->
  minor.RNA.protein 


minor.RNA.protein %>%
  mutate(proteinDetected = "No") %>%
  mutate(proteinDetected = replace(proteinDetected, !is.na(Gland),"Yes")) ->
  minor.RNA.protein
  
minor.RNA.protein %>% 
  filter(tpm > 0) %>%
  ggplot(aes(x = log2(tpm+1) , color = proteinDetected)) +
  geom_vline(xintercept = 2, color = "black")+
  geom_density()+
   scale_color_brewer(palette = "Set1")  +
  theme_light() +
  labs(
    x = "log2(tpm+1)",
    y = "Density",
    title = "Distribution of RNA expression in the Minor gland and protein detection status"
  ) ->
  ProteinRNAexpression.minor.1

minor.RNA.protein %>%
  filter(geneNameProtein %in% rownames(mat_DEseq)) ->
  selected

minor.RNA.protein %>%
  filter(tpm>2^2) %>%
  filter(Gland >0.00009) ->
  background





selected %>%
  ggplot(aes(x = log10(tpm), y = log10(Gland), color = Term, label = geneNameProtein)) +
  geom_point(data = background, alpha = 0.7, size = 0.8)+
  geom_point(alpha = 1, size = 4, shape = 2)+
  geom_smooth(data =background, color = "black", method = "lm")+
  ylab("log 10( protein level)") +
  ylab("Fraction of Intensity") +
  theme_light() +
  scale_color_brewer(palette = "Set2")  

minor.RNA.protein %>%
  filter(!is.na(Gland)) %>%
  filter(tpm > 3)->
  test
cor.test(x = log10(test$tpm), y = log10(test$Gland))  


DEseq.minor.gland.HM.results %>% filter(dir == "Up" ) %>% filter(sigLevel %in% c("***","****")) %>% filter(Tissue == "Major") -> upMinor

selected %>% filter(geneNameProtein %in% upMinor$geneNameProtein) ->selected.up

selected %>%
  ggplot(aes(x = log2(tpm+1), y = log2(Gland*100), color = Term, label = geneNameProtein)) +
  geom_point(data = background, alpha = 0.7, size = 0.6)+
  geom_point(alpha = 1, size = 3)+
  geom_smooth(data =background, color = "black", method = "lm")+
  geom_text_repel(data =selected.up, aes(label = geneNameProtein) , color = "black" )+
    labs(
    x = "log2(tpm+1)",
    y = "Percent of Intensity",
    title = "RNA and protein levels in the Minor gland\n of genes with idetified protein expression"
  )+ 
  scale_color_brewer(palette = "Set2") +
  theme_light()+ 
  xlim(0,18)->
  ProteinRNAexpression.minor.2



ProteinRNAexpression.minor.1 + ProteinRNAexpression.minor.2 + plot_layout(ncol = 1) -> RNAprotein.disitribution

ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/SuppFigure6/ProteinRNAConsentrations.pdf",
       height = 8, width = 6,plot = RNAprotein.disitribution)


cor.test(log2(background$tpm),log2(background$Gland))

```



#### Compare glands

```{r , message=FALSE, warning=FALSE}

geneInfo.QC %>% inner_join(minorProteomics.expressed)-> geneInfo.QC.minor.prot


sampleInfo.QC.merged %>%
  mutate(Tissue = as.character(Tissue)) %>%
  mutate(sample = as.character(sample)) %>%
  filter(Tissue %in% c("Head","Major","Minor"))-> 
  sampleInfo.QC.merged.DE



dds.Gland = getDEseq2Object(count.df = bulk.data.merged.QC.long,
                 md =sampleInfo.QC.merged.DE,
                 gi =geneInfo.QC.minor.prot, 
                 design =  "Tissue" )
dds.Gland = DESeq(dds.Gland)


resultsNames(dds.Gland)
#Sac_center_vs_Tail_proximal

resMiHe <-results(dds.Gland, 
               contrast =
                 list(c("designMinor"),c("designHead")), 
               listValues = c(1,-1))

as.data.frame(resMiHe) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Head") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.MiHe.DEseq2


resMiMa <-results(dds.Gland, 
               contrast =
                 list(c("designMinor"),c("designMajor")), 
               listValues = c(1,-1))

as.data.frame(resMiMa) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Major") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.MiMa.DEseq2



resMaHe <-results(dds.Gland, 
               contrast =
                 list(c("designMajor"),c("designHead")), 
               listValues = c(1,-1))

as.data.frame(resMaHe) %>%
  rownames_to_column(var = "gene_id") %>%
  mutate(Tissue = "Major_Head") %>%
  select(gene_id,Tissue,baseMean,log2FoldChange,padj) %>%
  arrange(padj) %>% inner_join(geneInfo) %>%
  filter(!is.na(padj))->  
  geneInfo.MiMa.DEseq2


bind_rows(geneInfo.MiMa.DEseq2,geneInfo.MiHe.DEseq2)->
  DEseq.minor.gland.results

DEseq.minor.gland.results %>%
  group_by(gene_id) %>%
  summarise(minPadj = min(padj), maxFoldChange = max(log2FoldChange)) %>%
  arrange(minPadj) %>%
  inner_join(DEseq.minor.gland.results) %>%
  filter(minPadj < 0.05,maxFoldChange >log2(5) ) %>%
  filter(signalP == "Yes")->
  DEseq.minor.gland.pruned.results


DEseq.minor.gland.results %>%
  filter(geneNameProtein %in%  rownames(DEseq.parts.A.results.pruned.signalP.wide)) %>%
  mutate(sigLevel = "None") %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.1, "**")) %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.05, "***")) %>%
  mutate(sigLevel = replace(sigLevel, padj < 0.0001, "****")) %>%
  mutate(dir = "Up") %>%
  mutate(dir = replace(dir, log2FoldChange < 0, "Down")) %>%
  select(geneNameProtein,Tissue,log2FoldChange,sigLevel,dir)->
  DEseq.minor.gland.HM.results



#DEseq.parts.A.results.pruned.signalP.wide = DEseq.parts.A.results.pruned.signalP.wide[DEseq.minor.gland.results.2$geneNameProtein,] 

#DEseq.parts.A.results.pruned.signalP.wide


geneInfo.MiHe.DEseq2%>%
  filter(padj < 0.05,log2FoldChange >log2(5) ) %>%
  inner_join(geneInfo) %>%
  select(geneNameProtein,Term) -> miHe.final


DEseq.minor.parts.A.results.pruned %>% inner_join(geneInfo) %>%
  distinct(geneNameProtein, Term) %>% inner_join(miHe.final) %>%
  group_by(Term) %>% summarise(n = n())



DEseq.minor.gland.results %>%
  filter(Tissue == "Head")  %>%
  filter(padj < 0.05) %>%
  filter(log2FoldChange >  log2(5)) ->
  DEseq.minor.head.results



DEseq.minor.head.results %>% inner_join(interpro5) %>%
  group_by(interpro_accession,interpro_description) %>%
  summarise(n = n()) %>%
  arrange(-n)
  

GOI_common = c("LASC007034","LASC0011995","lman2l",
        "x975_2","SpiCE-LMa5","AmSp-like1","AmSp-like2",
        "SpiCE-LMa3","SpiCE-LMa1",
        "SpiCE-LMa2","SpiCE-LMa4",
        "LASC009784","spa",
        "LASC007099","LASC017532")
geneInfo %>% 
  left_join(interpro5) %>% 
  filter(geneNameProtein %in%  GOI_common)


geneInfo %>% filter(grepl("SpiCE",geneNameProtein2 )) %>% select(geneNameProtein, geneNameProtein2) %>%
  arrange(geneNameProtein2)

interpro5  %>%
   filter(interpro_accession == "IPR043070") %>%
   group_by(geneName ) %>%
   summarise(n = n())  %>%
  arrange(-n)  -> IPR043070



```
