---
title: "Add gland Metadata to slides"
author: "Johan"
date: "11/11/2022"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    df_print: paged
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(patchwork)

source("../customFunctions.R")
source("../colorSchemes.R")




  
```


##Geneinfo

```{r load geneInfo}

source("../loadGeneInfo.R")
# Get gene info
geneInfo = getGeneInfoFull()

# add transcript info
geneInfo.transcript = addTranscriptInfo(geneInfo)


# get Major Ampulate silk genes
proteinsInMAGsilk = read_csv( "~/Box/bridgeSpider/annotation/silk.proteins.MAG.csv")


```

## Region info 

```{r }
order =   c( "Tail","ZoneA", "ZoneA_MaSp2" ,"ZoneA_MaSp1", "ZoneA_SpiCE-LMa3",
     "Sac", "ZoneB","ZoneB_MaSp3","ZoneBC","ZoneC", "ZoneC_SpiCE-LMa1",
     "Duct","Duct_1","Duct_2", "Not DE") 

bulkOrder =   c( "Tail","Sac","Duct") 

scOrder =   c("ZoneA_MaSp2" ,"ZoneA_MaSp1", "ZoneA_SpiCE-LMa3",
     "ZoneB_MaSp3","ZoneBC","ZoneC_SpiCE-LMa1",
     "Duct_1","Duct_2") 

spatialOrder =   c( "ZoneA","ZoneB","ZoneC") 

```



# Marker genes 
```{r}
numberOfTopGenes = 50

```
### Single cell 
```{r }
markerGenes.sc = read_csv("~/Box/bridgeSpider/annotation/sc.clusters.MAG.csv")

markerGenes.sc %>% 
  filter((pct.1 - pct.2 > 0.05)) ->
  markerGenes.sc

# Top marker genes
markerGenes.sc %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  arrange(p_val_adj, avg_log2FC) %>%
  slice_head(n =numberOfTopGenes) ->
  markerGenes.sc.top

# Only in the best cluster
markerGenes.sc.top %>% 
  group_by(gene) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 1) ->
  markerGenes.sc.top

markerGenes.sc.top %>%
  filter(avg_log2FC > 3) %>%
  arrange(cluster,p_val_adj) %>%
  group_by(cluster) %>%
  slice_head(n = 10)



```

### Spatial Data
```{r }
#write_csv(silk.proteins.marker.1.2, "~/Box/bridgeSpider/annotation/silk.proteins.marker.1.2.csv")
 markerGenes.spatial = read_csv("~/Box/bridgeSpider/annotation/spatial.clusters.MAG.csv")
markerGenes.spatial %>% 
  filter((pct.1 - pct.2 > 0.00)) ->
  markerGenes.spatial

 
markerGenes.spatial %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  arrange(p_val_adj, avg_log2FC) %>%
  slice_head(n =numberOfTopGenes) ->
  markerGenes.spatial.top

# Only in the best cluster
markerGenes.spatial.top %>% 
  group_by(gene) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 1) ->
  markerGenes.spatial.top

markerGenes.spatial.top %>%
  filter(avg_log2FC > 1) %>%
  arrange(cluster,p_val_adj)


l.spatial = split(markerGenes.spatial.top$gene, markerGenes.spatial.top$cluster)
l.spatial = l.spatial[c("ZoneA","ZoneB","ZoneC")]

```
### Bulk RNA data 
```{r}
#read_tsv("~/Box/bridgeSpider/bulkRNA/PCAanalysis.sigGenes.2.txt") %>%
#  select(geneName,Major) ->bulk.major

#bulk.major %>% filter(!is.na(Major)) %>%
#  filter(Major != "All")

bulk.DE.genes = read_csv(file = "../../results/bulkRNA/DEseq.sigGenes.csv")
bulk.DE.genes %>%
  rename(p_val_adj = "padj") %>%
  rename(avg_log2FC = "log2FoldChange") %>%
  rename(gene = "scGeneName") %>%
  rename(cluster = "part") %>%
  select(gene,p_val_adj,avg_log2FC,cluster ) ->
  bulk.DE.genes


bulk.DE.genes %>%
  filter(p_val_adj < 0.01) ->
  bulk.DE.genes.top
bulk.DE.genes %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  arrange(p_val_adj, avg_log2FC) %>%
  slice_head(n =numberOfTopGenes) ->
  bulk.DE.genes.top

# Only in the best cluster
bulk.DE.genes.top %>% 
  group_by(gene) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 1) ->
  bulk.DE.genes.top

bulk.DE.genes.top %>%
  filter(avg_log2FC > 1) %>%
  arrange(cluster,p_val_adj)

l.bulk = split(bulk.DE.genes.top$gene, bulk.DE.genes.top$cluster)
l.bulk = l.bulk[c("Tail","Sac","Duct")]


  

```


### All Marker genes combined
```{r }

markerGenes.spatial %>% select(gene,p_val_adj,avg_log2FC,cluster ) %>% mutate(sequence = "Spatial") -> markerGenes.spatial.top.small
markerGenes.sc %>% select(gene,p_val_adj,avg_log2FC,cluster )%>% mutate(sequence = "SingleCell") -> markerGenes.sc.top.small
bulk.DE.genes.top %>% select(gene,p_val_adj,avg_log2FC,cluster )%>% mutate(sequence = "BulkRNA") -> bulk.DE.genes.top.small

bind_rows(markerGenes.spatial.top.small,markerGenes.sc.top.small,bulk.DE.genes.top.small) -> all.top.small


all.top.small %>%
  group_by(cluster) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 10) %>%
  as.data.frame() %>%
  distinct(gene) ->
  all.top.small.10.genes


  
  
all.top.small %>% 
  group_by(gene) %>% 
  summarise(nrOfClusters = n()) %>%
  filter(nrOfClusters > 1) %>%
  mutate(present = 1) %>%
  inner_join(all.top.small) ->
  all.top.small.pruned 

  all.top.small.pruned %>%
#  filter(gene %in% all.top.small.10.genes$gene) %>%
  # all.top.small %>%
  mutate(present = 1) %>%
  select(gene,cluster,present) %>%
  pivot_wider(names_from = "cluster",values_from = "present",values_fill = 0) %>%
  column_to_rownames(var = "gene")->
  all.top.small.boolean

UpSetR::upset(all.top.small.boolean, nsets =14)
colnames(all.top.small.boolean)
DataOrder = c("")
all.top.small.boolean[c()]



         


```



### Silk protein with marker info
```{r}



all.top.small %>% arrange(p_val_adj) %>%
  filter(gene %in% proteinsInSilk$geneName) %>%
  rename(geneName = "gene")%>%
  right_join(proteinsInSilk) %>% 
  mutate(cluster = replace(cluster,is.na(cluster), "Not DE")) %>%
  mutate(sequence = replace(sequence,is.na(sequence), "None")) %>%
  mutate(cluster = replace(cluster, grepl("Sac_1", cluster), "ZoneBC")) %>%
  mutate(cluster = factor(cluster,levels = order)) %>%
  arrange(cluster,-mean ) ->
  spiderSilkGenes.annotated

spiderSilkGenes.annotated %>% 
  filter(sequence != "BulkRNA" ) %>%
  select(geneNameProtein, cluster,p_val_adj) %>%
  pivot_wider(names_from = cluster, values_from = p_val_adj, values_fill = 1)  ->
  spiderSilkGenes.annotated.wide



spiderSilkGenes.annotated.wide %>%
  mutate(geneNameProtein = factor(geneNameProtein,geneOrder)) %>%
  arrange(geneNameProtein) %>%
  mutate(Duct_2 = 1) -> spiderSilkGenes.annotated.wide



spiderSilkGenes.annotated.wide %>% select(c(geneNameProtein,scOrder))  %>%
  column_to_rownames("geneNameProtein") -> heatmap.data
  
 
spiderSilkGenes.annotated.wide  %>%
  pivot_longer(cols = -geneNameProtein) %>%
  mutate( name = factor(name, levels =order)) %>%
  ggplot(aes(x = geneNameProtein, y = name, fill = log10(value))) + 
  geom_tile(height = 0.4,width = 0.4 )+
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust = 1))+
  scale_fill_gradientn(colors = c("blue", "white", "red")) 

ggsave(filename = "~/Box/bridgeSpider/paper1/main/figures/figure6/markerGenes.sc.spatial.pdf", 
       width = 8, height = 4)




spiderSilkGenes.annotated  %>% 
  filter(sequence == "SingleCell" ) %>%
  mutate( cluster = factor(cluster, levels =order)) %>%
  mutate(geneNameProtein = factor(geneNameProtein,geneOrder)) %>%
  arrange(geneNameProtein, cluster) %>%
  ggplot(aes(x = geneNameProtein, y = cluster, fill = log10(p_val_adj))) + 
  geom_tile(height = 0.8,width = 0.8,color = "black" ,lwd = 0.2,
            linetype = 1 )+
  scale_fill_gradientn(colors = c("red", "white"))+ theme_light()+scale_x_discrete(drop=FALSE)+
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust = 1)) -> sc  

spiderSilkGenes.annotated  %>% 
  filter(sequence == "Spatial" ) %>%
  mutate( cluster = factor(cluster, levels =order)) %>%
  mutate(geneNameProtein = factor(geneNameProtein,geneOrder)) %>%
  arrange(geneNameProtein, cluster) %>%
  ggplot(aes(x = geneNameProtein, y = cluster, fill = log10(p_val_adj))) + 
  geom_tile(height = 0.8,width = 0.8,color = "black" ,lwd = 0.2,
            linetype = 1 ,labs.row =geneOrder) +
  scale_fill_gradientn(colors = c("red", "white"))+ theme_light()+scale_x_discrete(drop=FALSE)+
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust = 1)) -> spatial  

spiderSilkGenes.annotated  %>% 
  filter(sequence == "BulkRNA" ) %>%
  mutate( cluster = factor(cluster, levels =order)) %>%
  mutate(geneNameProtein = factor(geneNameProtein,geneOrder)) %>%
  arrange(geneNameProtein, cluster) %>%
  ggplot(aes(x = geneNameProtein, y = cluster, fill = log10(p_val_adj))) + 
  geom_tile(height = 0.8,width = 0.8,color = "black" ,lwd = 0.2,
            linetype = 1 ,labs.row =geneOrder) +
  scale_fill_gradientn(colors = c("red", "white"))+ theme_light()+scale_x_discrete(drop=FALSE)+
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5, hjust = 1)) -> bulk  


layout <- "
AAA
AAA
AAA
AAA
AAA
AAA
AAA
BBB
BBB
BBB
CCC
CCC
"
(sc + theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank()#remove x axis ticks
        )+ ylab("sc")) +
(spatial+ theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.title.x = element_blank())
        + ylab("spatial"))+ 
  (bulk + ylab("bulk")) + 
   plot_layout(design = layout,guides = 'collect')& theme(legend.position = 'none')

ggsave(filename = "~/Box/bridgeSpider/paper1/main/figures/figure6/markerGenes.sc.spatial.bulk.protein.pdf", 
       width = 5, height = 3)

```



##Protein expression

```{r Loading the proetin layer data}
# Retrieving the data

file = "Samples Report With Clusters for PB223_Urmi_Silk_and_factions_normalized_total_spectra.xls"

path = paste("../../data/proteomics/",file, sep  = "/")

proteomicsData.info = read_tsv("../../data/proteomics/Samples Report With Clusters for PB223_Urmi_Silk_and_factions_normalized_total_spectra.xls" ,skip = 63) %>%
  select( 10:33)

sampleInfo = as.data.frame(t(proteomicsData.info[1,]))  %>%
  rownames_to_column(var = "sampleInfo") %>%
  rename(Sample = "V1") 

proteomicsData = read_tsv("../../data/proteomics/Samples Report With Clusters for PB223_Urmi_Silk_and_factions_normalized_total_spectra.xls" ,skip = 64) %>%
  rename(tx_name = "Identified Proteins (4769/9420)") %>%
  rename(Ambiguity = "Protein Grouping Ambiguity")  %>%
  select(tx_name,Ambiguity, 10:33) %>%
  pivot_longer(cols = -c("tx_name","Ambiguity"), 
               names_to = "Sample", 
               values_to = "Normalized_Total_Spectra")  ->
  proteomicsData 


geneInfo.transcript %>% select(geneName,tx_name) %>%
  inner_join(proteomicsData) %>%
  mutate(Normalized_Total_Spectra = gsub(pattern = ",",
                                         replacement = ".",
                                         Normalized_Total_Spectra))%>% 
  mutate(Normalized_Total_Spectra =
           as.numeric(Normalized_Total_Spectra))%>%
  arrange(-Normalized_Total_Spectra) ->
  proteomicsData.df 


  
sampleInfo %>%
  separate(sampleInfo, into = c("Fraction","replicate"), sep = "\\.\\.\\.") %>%
  arrange(Fraction) %>%
  mutate(replicate  = rep(1:3,8)) %>%
  mutate(experiment = c(rep("Layer",15), rep("Urea",9))) %>%
  inner_join(proteomicsData.df) ->
  proteomicsData.df.info 

proteomicsData.df.info%>%
  group_by(Fraction,experiment ,replicate) %>%
  summarise( sum = sum(Normalized_Total_Spectra)) %>%
  inner_join(proteomicsData.df.info) %>%
  mutate(relativeFraction = Normalized_Total_Spectra / sum) ->
  proteomicsData.df.info

proteomicsData.df.info

proteomicsData.df.info %>%
  group_by(Sample,Fraction) %>%
  dplyr::summarise(n =length(unique(tx_name))) 




read_csv( file = "../../data/proteomics/proteomics.concentration.csv") %>% 
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder)) ->
  protein.initial.data

protein.initial.data %>%
  mutate(total = sum(mean)) %>%
  mutate(fraction = mean/total) %>%
  mutate(geneNameProtein =gsub(pattern = ("Spice"),replacement = ("SpiCE"),geneNameProtein)) ->
 protein.initial.data

#write_csv(protein.initial.data,  file = "../../data/proteomics/proteomics.concentration.csv")

protein.initial.data%>%
  mutate(group = "Protein") %>%
  mutate(percent = fraction *100) %>%
  arrange(-percent) %>%
  ggplot(aes(x = geneNameProtein, y = group, fill = percent ) )+
  geom_tile(width = 0.8, height = 0.8)+
  geom_text(aes(label = round(percent, 2)),size=4, ,angle=90,vjust=0.5, hjust = 0.5) +
  theme_minimal()  + 
  scale_fill_gradientn(trans = "log10", colors = c("blue", "white", "red"))+
  theme_light()+
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5))  -> concentration

protein.initial.data%>%
  mutate(group = "Protein") %>%
  mutate(percent = fraction *100) %>%
  arrange(-percent) %>%
  slice_head(n = 17) %>%
  ggplot(aes(x = geneNameProtein, y = group, fill = percent ) )+
  geom_tile(width = 0.8, height = 0.8)+
  geom_text(aes(label = round(percent, 2)),size=4, ,angle=90,vjust=0.5, hjust = 0.5) +
  theme_minimal()  + 
  scale_fill_gradientn(trans = "log10", colors = c("blue", "white", "red"))+
  theme_light()+
  theme(axis.text.x=element_text(angle = 30, hjust =1 ,vjust = 1))  -> concentration.17
 concentration.17

 protein.initial.data%>%
  mutate(group = "Protein") %>%
  mutate(percent = fraction *100) %>%
  arrange(-percent) %>%
  slice_head(n = 17) %>%
  ggplot(aes(x = geneNameProtein, y = group, fill = percent ) )+
  geom_tile(width = 0.8, height = 0.8)+
  theme_minimal()  + 
  scale_fill_gradientn(trans = "log10", colors = c("blue", "white", "red"))+
  theme_light()+
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5))  -> concentration.17.no.text
 concentration.17.no.text

 
 
bulk.major.expression = read_csv("../../data/bulkRNA/normExpression.protein.silk.csv") 


bulk.major.expression %>%
  mutate(group = "RNA") %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder)) ->
  bulk.major.expression

bulk.major.expression %>%
  ggplot(aes(x = geneNameProtein, y = group, fill = norm_exp ) )+
  geom_tile(width = 0.8, height = 0.8)+
  geom_text(aes(label = round(log2(norm_exp), 2)),size=4, ,angle=90,vjust=0.5, hjust = 0.5) +
  theme_minimal()  + 
  scale_fill_gradientn(trans = "log10", colors = c("blue", "white", "red"),values = c(0,0.5,1))+
  theme_light()+
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) -> expression


bulk.major.expression %>%
  arrange(-mean) %>% 
  slice_head(n = 17)->
  bulk.major.expression.17

bulk.major.expression.17 %>%
  ggplot(aes(x = geneNameProtein, y = group, fill = norm_exp ) )+
  geom_tile(width = 0.8, height = 0.8)+
  geom_text(aes(label = round(log2(norm_exp), 2)),size=4, ,angle=90,vjust=0.5, hjust = 0.5) +
  theme_minimal()  + 
  scale_fill_gradientn( trans = "log10",colors = c("blue", "white", "red"),values = c(0,0.5,1))+
  theme_light()+
  theme(axis.text.x=element_text( hjust = 0,vjust = 0.5)) -> expression.17


expression.17/concentration.17.no.text
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/protein.RNA.levels.withLegend.pdf", width = 10, height = 10)

concentration.17.no.text+ NoLegend() -> conc.2

ggsave(plot = conc.2,filename = "~/Box/bridgeSpider/paper1/main/figures/figure7/protein.17.levels.noLegend.pdf",
       width = 2.9,
       height =1.5)



```
#### Expression of spidersilk genes


```{r}


proteomicsData.df.info %>%
  filter(experiment == "Urea") %>%
  as.data.frame() %>%
  select(geneName) %>%
  filter(geneName %in% proteinsInSilk$geneName) %>%
  distinct(geneName ) ->
    proteomicsData.df.info.urea.geneNames
  
  proteomicsData.df.info %>%
    filter(experiment == "Urea") %>%
    filter(geneName %in% proteomicsData.df.info.urea.geneNames$geneName) ->
  proteomicsData.df.info.urea  
  
  
  proteomicsData.df.info.urea %>%
  select(geneName , Normalized_Total_Spectra, Fraction, replicate)%>%
    group_by(Fraction, replicate) %>%
    summarise(totalFraction = sum(Normalized_Total_Spectra)) %>%
    inner_join(proteomicsData.df.info.urea) %>%
    mutate(percentFraction = Normalized_Total_Spectra/totalFraction*100) ->
  proteomicsData.df.info.urea.df 



    
    
proteomicsData.df.info.urea %>%
  select(geneName , Normalized_Total_Spectra, Fraction, replicate) %>%
  group_by(geneName,Fraction) %>%
  summarise(Normalized_Total_Spectra = mean(Normalized_Total_Spectra)) ->
  proteinDistribution.urea.relative.df


  proteomicsData.df.info.urea %>%
  select(geneName , Normalized_Total_Spectra, Fraction, replicate) %>%
    group_by(geneName) %>%
    summarise(TotalGeneSpectra = sum(Normalized_Total_Spectra)) %>%
    inner_join(proteomicsData.df.info.urea) %>%
    mutate(percentGene = Normalized_Total_Spectra/TotalGeneSpectra*100) ->
  proteinDistribution.urea.relative.df 
  
  proteinDistribution.urea.relative.df %>% 
    select(geneName, Fraction, Normalized_Total_Spectra, percentGene) ->
    proteinDistribution.urea.relative.df
  

  spiderSilkGenes.annotated %>%
  group_by(geneName,sequence) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 1) %>%
  arrange(geneNameProtein) %>%
  filter(cluster != "Not DE") ->
  spiderSilkGenes.annotated.2
  
proteinDistribution.urea.relative.df %>% inner_join(spiderSilkGenes.annotated.2) ->
  proteinDistribution.urea.relative.df.markers
  



proteomicsData.df.info.urea.df %>%
  inner_join(protein.info) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder)) %>%
  mutate(Fraction = factor(Fraction, levels = c("Silk in 8M Urea","Silk in 4M Urea","Silk in 2M Urea"))) %>%
  ggplot(aes(y = percentFraction, fill = Fraction, x = geneNameProtein)) + geom_boxplot() + coord_flip()+scale_fill_manual(values = AllColors2)

ggsave("~/Box/bridgeSpider/paper1/supplementary/SupplFig/suppFigure38.UREA.replicates.pdf")


proteomicsData.df.info.urea.df %>%
  inner_join(protein.info) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder)) %>% group_by(Fraction,geneNameProtein ) %>%
  summarise(sd = sd(percentFraction),
            percentFraction = mean(percentFraction),
            n = n()) %>%
  mutate(Fraction = factor(Fraction, levels = c("Silk in 8M Urea","Silk in 4M Urea","Silk in 2M Urea"))) ->
  barPlotInfo

ggplot(barPlotInfo, aes(x=geneNameProtein, y=percentFraction, fill=Fraction)) + 
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar(aes(ymin=percentFraction, ymax=percentFraction+sd), width=.4,
                 position=position_dodge(.9))+
  theme_light()+
  coord_flip()+
  scale_fill_manual(values = AllColors2)   

ggsave("~/Box/bridgeSpider/paper1/supplementary/SupplFig/suppFigure38.UREA.replicates.pdf" ,width = 4, height = 10)


unique(proteomicsData.df.info.urea.df$Fraction)

sd(c(2,3,4))
```
  
   
```{r}  


order.present= intersect(proteinDistribution.urea.relative.df.markers$cluster,order)

proteinDistribution.urea.relative.df.markers %>% 
  distinct(sequence, cluster) -> clusterInfo


proteinDistribution.urea.relative.df.markers %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder))  %>%
  distinct(geneNameProtein,cluster) %>% 
  mutate(present = 1) %>% 
  inner_join(clusterInfo) %>%
  mutate(cluster = factor(cluster, levels = order)) %>%
  mutate(sequence = factor(sequence, 
                           levels = c("SingleCell",
                                      "Spatial",
                                      "BulkRNA",
                                      "None")))->
  proteinDistribution.urea.relative.df.markers  
  
  
proteinDistribution.urea.relative.df.markers %>%
  select(cluster,geneNameProtein ) %>%
  mutate(present = 1 ) %>%
  pivot_wider(names_from = "cluster", values_from = "present",values_fill = 1) %>%
  pivot_longer(cols = -geneNameProtein, 
               names_to = "cluster",
               values_to = "present")  ->
  background

proteinDistribution.urea.relative.df.markers %>%
  distinct(cluster, sequence) %>%
  inner_join(background) %>%
  mutate(sequence = factor(sequence, 
                           levels = c("SingleCell",
                                      "Spatial",
                                      "BulkRNA",
                                      "None")))%>%
  mutate(cluster = factor(cluster, levels = order)) ->
  background

#library(ggnewscale)


proteinDistribution.urea.relative.df.markers %>%
  filter(sequence != "None") ->
  proteinDistribution.urea.relative.df.markers.2

ggplot() + 
  geom_tile(data = proteinDistribution.urea.relative.df.markers.2,
            mapping = aes(x = geneNameProtein, 
                          y = sequence, fill = cluster ),
            width = 0.8, height = 0.8,)+
  theme_minimal()  + 
  scale_fill_manual(values = AllColors2)+ theme_light()+scale_x_discrete(drop=FALSE)+
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5))  ->
  markers

markers+ theme(legend.position="bottom")

ggsave("~/Box/bridgeSpider/paper1/main/figures/figure6/markers.pdf", height = 3, width = 6)


proteinDistribution.urea.relative.df.markers %>%
  filter(sequence == "SingleCell") %>%
  filter(grepl(pattern = "ZoneA", cluster))->
  zoneA_markers

zoneA_markers %>% inner_join(proteinsInSilk)%>%
  inner_join(select(geneInfo,geneName, scGeneName)) %>%
  select(scGeneName, geneNameProtein,cluster) ->zoneA_markers 
  

ggplot() + 
  geom_tile(data = proteinDistribution.urea.relative.df.markers.2,
            mapping = aes(x = geneNameProtein, 
                          y = sequence, fill = cluster ),
            width = 0.8, height = 0.8,)+
  theme_minimal()  + 
  scale_fill_manual(values = AllColors2)+ theme_light()+scale_x_discrete(drop=FALSE)+
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5))  ->
  markers

markers+ theme(legend.position="bottom")



#write_csv(protein.markers , "~/Box/bridgeSpider/annotation/silk.proteins.marker.1.1.csv")



#protein.markers[c(2,3,1,6,4,5,7,9,10,8,27,23,22,24,28,25,26,11,12,13,16,15,14,21,18,20,17,19),] ->
#  protein.markers

#write_csv(protein.markers , "~/Box/bridgeSpider/annotation/silk.proteins.marker.1.1.csv")


#protein.markers$geneNameProtein

```




```{r}  
  
  
protein = proteinDistribution.urea.relative.df%>%
  filter(geneName %in% proteinsInSilk$geneName) %>%
  mutate(geneName = factor(geneName, levels = proteinsInSilk$geneName)) %>%
  ggplot(aes(x = geneName,
             y = Normalized_Total_Spectra,
             fill =Fraction )) +
  geom_boxplot() + 
  theme_minimal()+
  theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
  theme(legend.position = "none") +ylab("proteomics") +
  scale_fill_manual(values = UreaColors)


proteinDistribution.urea.relative.df %>%
  inner_join(proteinsInSilk) %>%
  group_by(geneNameProtein) %>%
  summarise(mean = mean(Normalized_Total_Spectra),sd = sd(Normalized_Total_Spectra)) ->
  p.stat

proteinDistribution.urea.relative.df  %>%
  inner_join(proteinsInSilk) %>%
  group_by(geneNameProtein, Fraction) %>%
  summarise(expression = mean(Normalized_Total_Spectra)) %>%
  inner_join(p.stat) %>%
  mutate(zscore = (expression-mean)/ sd) %>%
  mutate(geneNameProtein = factor(geneNameProtein, geneOrder ))  %>%
  mutate(Fraction = factor(Fraction,  levels = c("Silk in 8M Urea","Silk in 4M Urea","Silk in 2M Urea") ))-> 
  proteinDistribution.zscores
  
  
  proteinDistribution.zscores %>%
    ggplot(aes(x = geneNameProtein,y =  Fraction, fill = zscore)) + 
    geom_tile(width=0.9,height = 0.9,
            lwd = 1,
            linetype = 1) +
  theme_minimal()+ theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) + 
  scale_fill_gradientn(colors = c("blue", "white", "red")) +scale_x_discrete(drop=FALSE)



  proteinDistribution.zscores %>% 
    ggplot(aes(x = geneNameProtein,y =  Fraction, fill = zscore)) + 
    geom_tile(width=0.9,height = 0.9,
            lwd = 1,
            linetype = 1) +
  theme_minimal()+ theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) + 
  scale_fill_gradientn(colors = c("blue", "white", "red"))

ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/relativeProteinConentration.withLegends.pdf")

```

#### Statistical t-test
```{r }
proteinDistribution.urea.relative.df %>% 
  group_by(geneName) %>%
  summarise(n = n()) ->
  names

## Greater
 proteinDistribution.urea.relative.df %>% 
  group_by(geneName) %>%
  group_map(~ pairwise.t.test(.x$Normalized_Total_Spectra, .x$Fraction,pool.sd = FALSE,alternative = "greater",p.adjust.method = "BH" )) ->
 proteinDistribution.urea.relative.df.t.test  

names(proteinDistribution.urea.relative.df.t.test) = names$geneName

ttest.list = list()

for(name in names$geneName){
  proteinDistribution.urea.relative.df.t.test[[name]]$p.value %>%
    as.data.frame() %>%
    rownames_to_column(var = "class2") %>%
    pivot_longer(cols = c("Silk in 2M Urea","Silk in 4M Urea"), names_to = "class1", values_to = "pvalue") %>%
    mutate(direction = "greater") %>%
    mutate(geneName = name)->
    masspec.greater
  ttest.list[[name]] = masspec.greater
}
bind_rows(ttest.list) -> masspec.greater


## less
 proteinDistribution.urea.relative.df %>% 
  group_by(geneName) %>%
  group_map(~ pairwise.t.test(.x$Normalized_Total_Spectra, .x$Fraction,pool.sd = FALSE,alternative = "less",p.adjust.method = "BH" )) ->
 proteinDistribution.urea.relative.df.t.test  

names(proteinDistribution.urea.relative.df.t.test) = names$geneName

ttest.list = list()

for(name in names$geneName){
  proteinDistribution.urea.relative.df.t.test[[name]]$p.value %>%
    as.data.frame() %>%
    rownames_to_column(var = "class2") %>%
    pivot_longer(cols = c("Silk in 2M Urea","Silk in 4M Urea"), names_to = "class1", values_to = "pvalue") %>%
    mutate(direction = "less") %>%
    mutate(geneName = name)->
    masspec.less
  ttest.list[[name]] = masspec.less
}

bind_rows(ttest.list) -> masspec.less

bind_rows(masspec.less, masspec.greater) -> masspec.ttest


masspec.ttest %>% 
  filter(class2 == "Silk in 8M Urea")  %>%
  filter(class1 == "Silk in 2M Urea") ->
  masspec.ttest
  
masspec.ttest %>%
  filter(pvalue < 0.1) %>%
  mutate(sig = "p-value < 0.1") %>%
  mutate(sig = replace(sig, pvalue < 0.05, "p-value < 0.05")) %>%
  mutate(sig = replace(sig, pvalue < 0.001, "p-value < 0.001")) %>%
  mutate(Fraction = "Silk in 8M Urea") %>%
  mutate(Fraction = replace(Fraction, direction == "less","Silk in 2M Urea" )) %>%
  select(geneName,Fraction,pvalue, sig)%>% 
  inner_join( proteinsInSilk)%>%
  select(geneNameProtein,Fraction,pvalue, sig)  ->
  masspec.ttest.sig
  

```



```{r}  


  proteinDistribution.urea.relative.df %>%
  inner_join(proteinsInSilk) %>%
  distinct(geneNameProtein,Fraction,percentGene) %>%
  group_by(geneNameProtein,Fraction) %>%
  summarise(percentGene = sum(percentGene))%>% 
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder)) %>%
  left_join(masspec.ttest.sig) %>%
  mutate(sig = replace(sig, is.na(sig), "None")) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder)) ->
  proteinDistribution.urea.relative.df.summary



proteinDistribution.urea.relative.df.summary%>%
  inner_join(proteinDistribution.zscores) %>%
  
  mutate(Fraction = factor(Fraction, levels = c("Silk in 8M Urea","Silk in 4M Urea","Silk in 2M Urea") )) %>%
  ggplot(mapping = aes(x = geneNameProtein, 
                       y = Fraction,
                       fill = zscore)) + 
  geom_tile(width=0.8,height = 0.8,color = "black",
            lwd = 0.5, aes(linetype = sig)) +
  theme_light()+ theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) + 
  scale_fill_gradientn(colors = c("blue", "white", "red")) + 
  scale_linetype_manual(values=c(0,1,2))->
  proteinDistribution


proteinDistribution.urea.relative.df.summary%>%
  inner_join(proteinDistribution.zscores) %>%
  filter(geneNameProtein %in% proteinsInSilk.17$geneNameProtein) %>%
  mutate(Fraction = factor(Fraction, levels = c("Silk in 8M Urea","Silk in 4M Urea","Silk in 2M Urea") )) %>%
  ggplot(mapping = aes(x = geneNameProtein, 
                       y = Fraction,
                       fill = zscore)) + 
  geom_tile(width=0.8,height = 0.8,color = "black",
            lwd = 0.5, aes(linetype = sig)) +
  theme_light()+ theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) + 
  scale_fill_gradientn(colors = c("blue", "white", "red")) + 
  scale_linetype_manual(values=c(0,1,2))->
  proteinDistribution.17





  
  proteinDistribution + theme(legend.position="bottom")
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/protein.levels.distribution.pdf", height = 3, width = 6)


  

```






# Single cell 

```{r }

major.sc.combined.integrated.sct = readRDS( 
    file = "~/Box/bridgeSpider/scRNA/LarSco.combined_filtered.integrated.MajorSpecific.SCT.2.rds")

clusterRelationship = read_csv("../../results/scRNA/majorGland/clusterRelationship.csv")

clusterRelationship %>%
  mutate(clusterZones = clusterMaAG3) %>%
  mutate(clusterZones = gsub(pattern ="Tail" , replacement ="ZoneA" , x = clusterZones)) %>%
  mutate(clusterZones = gsub(pattern ="Sac_LASC013329" , replacement ="ZoneC_Spice-LMa1" , x = clusterZones) )%>%
  mutate(clusterZones = gsub(pattern ="Sac_MaSp3" , replacement ="ZoneB_MaSp3" , x = clusterZones)) ->
  clusterRelationship

major.sc.combined.integrated.sct = addMetaDataToSeruatObject(major.sc.combined.integrated.sct, clusterRelationship)

major.sc.combined.integrated.sct = SetIdent(major.sc.combined.integrated.sct, value = "clusterZones")



AverageExpression(major.sc.combined.integrated.sct)   ->
  scE 



scE$SCT %>%
  as.data.frame() %>%
  rownames_to_column(var = "scGeneName") %>%
  pivot_longer(-scGeneName,
               names_to = "cluster",
               values_to = "expression") %>%
  inner_join(proteinsInSilk) %>%
  select(geneNameProtein,cluster,expression) %>%
  mutate(cluster =  gsub("ZoneA_Spice8","ZoneA_SpiCE-LMa3", cluster)) %>%
  mutate(cluster =  gsub("ZoneC_Spice-LMa1","ZoneC_SpiCE-LMa1",cluster)) %>%
  mutate(cluster =  gsub("Sac_1","ZoneBC",cluster)) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder))  %>%
  mutate(cluster = factor(cluster, levels = rev(scOrder))) ->
  sc.sct.expression

spiderSilkGenes.annotated %>%
  filter(sequence == "SingleCell") %>%
  select(geneNameProtein,cluster) %>%
  mutate(MarkerGene = "Yes") %>%
  right_join(sc.sct.expression) %>%
  mutate(MarkerGene = replace(MarkerGene,is.na(MarkerGene),"No")) ->
sc.sct.expression  

  
table(sc.sct.expression$cluster)

sc.sct.expression %>%
  mutate(lg2Exp = log2(expression)) %>%
  group_by(geneNameProtein) %>%
  summarise(mean = mean(expression), std = sd(expression)) %>%
  inner_join(sc.sct.expression) %>%
  mutate(lg2Exp = log2(expression)) %>%
  mutate(zscore = (expression-mean)/std) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder))  %>%
  mutate(cluster = factor(cluster, levels = scOrder)) %>%
  mutate(MarkerGene = factor(MarkerGene, levels = c("Yes","No"))) ->
  sc.sct.expression


MarkerColors = c("Yes" = "#111111ff", "No" = "#ffffff00") %>% as.vector()

sc.sct.expression %>% 
  filter(cluster %in% c("ZoneA_MaSp2" ,"ZoneA_MaSp1", "ZoneA_SpiCE-LMa3",
     "ZoneB_MaSp3","ZoneC_SpiCE-LMa1","ZoneBC") ) -> sc.sct.expression

ggplot(data = sc.sct.expression,
       mapping = aes(x = geneNameProtein, 
                     y = cluster, 
                     fill = zscore,
                     color = MarkerGene )) + 
  geom_tile(lwd = 0.5,
            linetype = 1,width=0.8,height = 0.8)+
  theme_light() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) +
  scale_fill_gradientn(colors = c("blue", "white", "red"))  +
  scale_colour_manual(values =c("#111111","#ffffff00")) ->
  scExpressionGraf

sc.sct.expression %>% 
  filter(geneNameProtein %in% proteinsInSilk.17$geneNameProtein) %>%
  ggplot(
       mapping = aes(x = geneNameProtein, 
                     y = cluster, 
                     fill = zscore,
                     color = MarkerGene )) + 
  geom_tile(lwd = 0.5,
            linetype = 1,width=0.8,height = 0.8)+
  theme_light() +
  theme(axis.text.x=element_text(angle = -90, hjust = 0,vjust = 0.5)) +
  scale_fill_gradientn(colors = c("blue", "white", "red"))  +
  scale_colour_manual(values =c("#111111","#ffffff00")) ->
  scExpressionGraf.17


scExpressionGraf.17
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/sc.relative.expression.withlegend.pdf")

```

## Combine cluster, RNA and protein levels
```{r}  

layout <- "
AAA
AAA
AAA
BBB
BBB
BBB
BBB
BBB
CCC
CCC
CCC"
(markers + theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank()#remove x axis ticks
        )+ ylab("RNA Markers")) +
(scExpressionGraf+ theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.title.x = element_blank())
        + ylab("scRNA"))+ 
  (proteinDistribution + ylab("mass-spec")) + 
   plot_layout(design = layout,guides = 'collect')& theme(legend.position = 'none')
  
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/urea.expression.pdf", width = 6, height = 4)









```

### Combine RNA and protein top 17
```{r}




layout <- "
AAA
AAA
AAA
BBB
BBB
BBB
BBB
BBB
CCC
CCC
CCC
DDD
DDD
DDD"



(expression.17 + theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank()#remove x axis ticks
        )+ ylab("")) +
(scExpressionGraf.17+ theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.title.x = element_blank())
        + ylab("scRNA"))+ 
(proteinDistribution.17+ theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.title.x = element_blank())
        + ylab("mass-spec"))+ 
  (concentration.17 + ylab("")) + 
   plot_layout(design = layout,guides = 'collect')& theme(legend.position = 'none')
  
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/protein.RNA.expression.17.pdf", width = 4, height = 4)



(expression + theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank()#remove x axis ticks
        )+ ylab("")) +
(scExpressionGraf+ theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.title.x = element_blank())
        + ylab("scRNA"))+coord_flip()+ 
(proteinDistribution+ theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.title.x = element_blank())
        + ylab("mass-spec"))+coord_flip()+ 
  (concentration + ylab("")+coord_flip()) + 
   plot_layout(design = layout,guides = 'collect')& theme(legend.position = 'none')
  
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/protein.RNA.expression.28.pdf", width = 6, height = 6)



expression/concentration
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/protein.RNA.17.withLegend.pdf")




```
## Combine cluster,  protein levels
```{r}  

layout <- "
AAA
AAA
AAA
CCC
CCC
CCC
"
(markers + theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank()#remove x axis ticks
        )+ ylab("RNA Markers")) +
  (proteinDistribution + ylab("mass-spec")) + 
   plot_layout(design = layout,guides = 'collect')& theme(legend.position = 'none')
  
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure8/urea.markers.pdf", width = 4, height = 2.7)





```




```{r}

all.top.small %>% 
  group_by(gene) %>% 
  summarise(nrOfClusters = n()) %>%
  filter(nrOfClusters > 1) %>%
  filter(gene %in% all.top.small.10.genes$gene) %>%
  inner_join(all.top.small) %>%
  select(gene,cluster,avg_log2FC) %>%
  pivot_wider(names_from = "cluster",values_from = "avg_log2FC",values_fill = 0) %>%
  column_to_rownames(var = "gene")->
  all.top.small.l2f




all.top.small.l2f = all.top.small.l2f[,c("Tail","ZoneA","ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
                             "Sac","Sac_1" ,
                             "ZoneB","ZoneB_MaSp3",
                             "ZoneC","ZoneC_SpiCE-LMa1",
                             "Duct","Duct_1","Duct_2")] 

all.top.small.l2f[all.top.small.l2f>5] = 5


all.top.small.l2f %>% rownames_to_column() %>%
  arrange(ZoneA_MaSp2,ZoneA_SpiCE8,ZoneA_MaSp1,Tail,
                             Sac_1 ,ZoneB_MaSp3,
                             `ZoneC_SpiCE-LMa1`,Sac,
                             Duct_1,Duct_2) %>%
  column_to_rownames()->all.top.small.l2f

#pdf("~/Box/bridgeSpider/paper1/main/figures/figure5/ExpressionComparison.heatmap.pdf")
ComplexHeatmap::Heatmap(all.top.small.l2f, cluster_columns = FALSE, cluster_rows = FALSE )
#dev.off()


all.top.small %>%
  filter(gene %in% rownames(all.top.small.l2f) ) %>%
  mutate(gene = factor(gene, levels = rownames(all.top.small.l2f))) %>%
  mutate(cluster = factor(cluster, levels = c("Tail","Sac","Duct",
                                              "ZoneA","ZoneB","ZoneC",
                                              "ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
                                              "Sac_1","ZoneB_MaSp3","ZoneC_SpiCE-LMa1",
                                              "Duct_1","Duct_2"))) ->
  all.top.small.l2f.long

ggplot(all.top.small.l2f.long, aes(x = gene, y =cluster, size = avg_log2FC,  color = sequence, fill = cluster)) +
    geom_point(shape=21) +
  theme_light()+ theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
  scale_fill_manual(values = AllColors)+guides(fill = FALSE) 

#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure5/ExpressionComparison.dotplot.pdf", width = 10)
```



#### Dependent of proteins


```{r }

all.top.small.l2f.long
scE.integrated =scE$integrated
scE.sct =scE$SCT

class(scE.integrated)




genes = intersect(rownames(scE.),unique(proteinDistribution.urea.relative.df$geneName)) 
markerGenes.sc %>% 
  filter(p_val_adj < 0.0001) %>%
  filter(avg_log2FC > 3) %>%
  select(gene, cluster,avg_log2FC) %>%
  filter(gene %in% genes) %>%
  rename( geneName = "gene") ->
  genesInCommon


genesInCommon %>%
  arrange(geneName, -avg_log2FC) %>%
  group_by(geneName) %>%
  slice_head(n = 1) ->
  order



scE.sct[genesOrder$geneName, ]
genesOrder


scE.integrated[scE$integrated<0 ] = 0 

scE.integrated[genesOrder$geneName,]

proteinsInSilk %>%
  select(geneName, geneNameProtein) %>%
  inner_join(order) %>%
  arrange(cluster,geneName) ->
  genesOrder


proteinsInSilk %>%
  select(geneName, geneNameProtein) %>%
  inner_join(genesInCommon) %>%
  arrange(cluster,geneName) ->
  genesInCommon


  
proteinDistribution.urea.relative.df  %>%
  group_by(geneName) %>%
  summarise(mean = mean(Normalized_Total_Spectra), sd = sd(Normalized_Total_Spectra)) %>%
  ungroup %>%
  inner_join(proteinDistribution.urea.relative.df) %>%
  inner_join(genesInCommon) %>%
  mutate(zscore = (Normalized_Total_Spectra - mean)/sd) %>%
  mutate(cluster = factor(cluster, levels = zoneOrder)) %>%
     mutate(cluster = factor(cluster, levels = c("ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
                                              "Sac_1","ZoneB_MaSp3","ZoneC_SpiCE-LMa1"))) %>% 
     filter(!is.na(cluster))  ->
  proteinDistribution.urea.relative.df.2


proteinDistribution.urea.relative.df.2 %>%
  distinct(cluster,geneNameProtein) %>%
  arrange(cluster) ->
  geneOrder

geneOrder[-9,] -> geneOrder

proteinDistribution.urea.relative.df.2 %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = geneOrder$geneNameProtein)) ->
  proteinDistribution.urea.relative.df.2

  library(ggforce)
  proteinDistribution.urea.relative.df.2%>%
     ggplot(aes(x = geneNameProtein, y = zscore, fill =Fraction )) +
     geom_boxplot() + 
    theme_minimal()+
    theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
   ylab("proteomics") +
    scale_fill_manual(values = AllColors)+ 
  facet_row(facets = "cluster", scales = "free_x", space = "free")
  

    proteinDistribution.urea.relative.df.2%>%
      group_by(geneNameProtein,Fraction,cluster) %>%
      summarise(zscore = mean(zscore)) %>%
     ggplot(aes(x = geneNameProtein, y = Fraction, size =zscore, fill = cluster )) +
     geom_point(shape=21) + 
    theme_minimal()+
    theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
   ylab("proteomics") +
    scale_fill_manual(values = AllColors)+ 
  facet_row(facets = "cluster", scales = "free_x", space = "free") -> proteomics

    
  
  intersect(proteinDistribution.urea.relative.df.2$cluster,names(bulkClustersZone))
```

  
```{r}
proteinDistribution.urea.relative.df.2%>%
     ggplot(aes(x = geneNameProtein, y = zscore, fill = cluster )) +
     geom_boxplot() + 
    theme_minimal()+
    theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
   ylab("proteomics") +
    scale_fill_manual(values = bulkClustersZone)+ 
  facet_row(facets = "Fraction", scales = "free_x", space = "free") 
   

proteinsInSilk %>%
  select(geneName, geneNameProtein) %>%
  inner_join(genesInCommon) %>%
  arrange(cluster,geneName) ->
  genesInCommon



unique(genesInCommon$cluster)
genesOrder


```


```{r }

all.top.small.l2f.long %>% filter(gene %in% proteinInfo$geneName) ->
  all.top.small.l2f.long.protein 

all.top.small.l2f.long %>%as.data.frame()%>% distinct(cluster,sequence) -> cluster.relationship
all.top.small.l2f.long.protein %>%
  select(gene,cluster,avg_log2FC) %>%
  pivot_wider(values_from = "avg_log2FC" , names_from = "cluster", values_fill = 0) %>%
  pivot_longer(cols = unique(all.top.small.l2f.long.protein$cluster),values_to ="avg_log2FC", names_to = "cluster" ) %>%
  inner_join(cluster.relationship) ->
  all.top.small.l2f.long.protein
  
all.top.small.l2f.long.protein %>%  filter(sequence == "BulkRNA") %>%
   mutate(cluster = factor(cluster, levels = c("Tail","Sac","Duct",
                                              "ZoneA","ZoneB","ZoneC",
                                              "ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
                                              "Sac_1","ZoneB_MaSp3","ZoneC_SpiCE-LMa1",
                                              "Duct_1","Duct_2"))) %>%
ggplot( aes(x = gene, y =cluster, size = avg_log2FC,  color = sequence, fill = cluster)) +
    geom_point(shape=21) +
  theme_minimal()+ theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
  scale_fill_manual(values = AllColors)+guides(fill = FALSE)+ 
  theme(legend.position = "none", axis.text.x=element_blank(),axis.title.x = element_blank() )+ylab("bulk")   -> bulk.plot

all.top.small.l2f.long.protein %>%
mutate(cluster = factor(cluster, levels = c("Tail","Sac","Duct",
                                              "ZoneA","ZoneB","ZoneC",
                                              "ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
                                              "Sac_1","ZoneB_MaSp3","ZoneC_SpiCE-LMa1",
                                              "Duct_1","Duct_2"))) %>%
  filter(sequence == "SingleCell") %>%
ggplot( aes(x = gene, y =cluster, size = avg_log2FC,  color = sequence, fill = cluster)) +
    geom_point(shape=21) +
  theme_minimal()+ theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
  scale_fill_manual(values = AllColors)+ 
  theme(legend.position = "none", axis.text.x=element_blank() ,axis.title.x = element_blank() )+ylab("singleCell")  -> sc.plot

all.top.small.l2f.long.protein %>%
  mutate(cluster = factor(cluster, levels = c("Tail","Sac","Duct",
                                              "ZoneA","ZoneB","ZoneC",
                                              "ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
                                              "Sac_1","ZoneB_MaSp3","ZoneC_SpiCE-LMa1",
                                              "Duct_1","Duct_2"))) %>%
  filter(sequence == "Spatial") %>%
ggplot( aes(x = gene, y =cluster, size = avg_log2FC,  color = sequence, fill = cluster)) +
    geom_point(shape=21) +
  theme_minimal()+ theme(axis.text.x=element_text(angle = -90, hjust = 0)) + 
  scale_fill_manual(values = AllColors)+ theme(legend.position = "none") + 
  theme(legend.position = "none", axis.text.x=element_blank() ,axis.title.x = element_blank() )+ylab("spatial") -> spatial.plot


bulk.plot+ sc.plot+spatial.plot +plot_layout(ncol = 1, guides = "collect")


ggsave("~/Box/bridgeSpider/paper1/main/figures/figure5/ExpressionComparison.dotplot.proteinonly.pdf")


```

```{r }
bulk.DE.genes %>%
  group_by(part) %>%
  arrange(padj, log2FoldChange) %>%
  #  slice_min(order_by = p_val_adj,n =numberOfTopGenes) ->
  slice_head(n =numberOfTopGenes) ->
  marker.bulk.filtered.top25

marker.bulk.filtered.top25 %>% 
  group_by(geneName) %>%
  arrange(padj) %>%
  slice_head(n = 1) ->
  marker.bulk.filtered.top25


l.bulk = split(marker.bulk.filtered.top25$geneName, marker.bulk.filtered.top25$part)

l.sc = l.sc[c("ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
              "ZoneB_MaSp3","Sac_1","ZoneC_SpiCE-LMa1",
              "Duct_1","Duct_2")]

title2 = paste("Top",numberOfTopGenes,"marker genes", sep = " ")
o = overlap_phyper2(l.sc,l.bulk,title = title2, remove.diag = F,with.total = T)

```

#### Dependent of proteins
```{r }

numberOfTopGenes = 50
bulk.DE.genes %>%
  filter(avg_log2FC > 1) %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  arrange(p_val_adj, avg_log2FC) %>%
  #  slice_min(order_by = p_val_adj,n =numberOfTopGenes) ->
  slice_head(n =numberOfTopGenes) ->
  markerGenes.sc.top25

markerGenes.sc.top25 %>% 
  group_by(gene) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 1) ->
  markerGenes.sc.top25

bulk.DE.genes %>%
  filter(avg_log2FC > 1) %>%
  filter(p_val_adj < 0.05) %>%
  group_by(cluster) %>%
  arrange(p_val_adj, avg_log2FC) %>%
  #  slice_min(order_by = p_val_adj,n =numberOfTopGenes) ->
  slice_head(n =numberOfTopGenes) ->
  marker.zones.filtered.top25

marker.zones.filtered.top25 %>% 
  group_by(gene) %>%
  arrange(p_val_adj) %>%
  slice_head(n = 1) ->
  marker.zones.filtered.top25



marker.zones.filtered.top25 %>%
  select(gene, cluster) %>%
  rename(spatialCluster = "cluster") %>%
  inner_join(markerGenes.sc.top25) ->
  marker.combined

marker.combined %>% arrange(gene,p_val_adj) 


l.sc = split(markerGenes.sc.top25$gene, markerGenes.sc.top25$cluster)
l.spatial = split(marker.zones.filtered.top25$gene, marker.zones.filtered.top25$cluster)

l.spatial = l.spatial[c("ZoneA","ZoneB","ZoneC")]
l.sc = l.sc[c("ZoneA_MaSp2","ZoneA_SpiCE8","ZoneA_MaSp1",
              "ZoneB_MaSp3","Sac_1","ZoneC_SpiCE-LMa1",
              "Duct_1","Duct_2")]

title2 = paste("Top",numberOfTopGenes,"marker genes", sep = " ")

o = overlap_phyper2(l.sc,l.spatial,title = title2, remove.diag = F,with.total = T)





```









```{r}
DimPlot(major.sc.combined.integrated.sct,
        group.by = "clusterMaAG3",
        cols = bulkClustersGenes2)

DimPlot(major.sc.combined.integrated.sct,
        group.by = "clusterZones",
        cols = bulkClustersZones3)
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure5/dimPlot.zones.pdf")

bulkClustersGenes3

DoHeatmap(major.sc.combined.integrated.sct,features = $scGeneName,
          assay = "integrated", group.by ="clusterZones" , raster = FALSE, group.colors = bulkClustersZones3) + scale_fill_gradientn(colors = c("blue", "white", "red"))



test = BuildClusterTree(major.sc.combined.integrated.sct, dims =1:4,reduction ="pca")
PlotClusterTree(test, direction = "downwards")



```

```{r}
library(SingleCellExperiment)

counts = major.sc.combined.integrated.sct@assays$RNA@counts

pdata = major.sc.combined.integrated.sct@meta.data

geneInfo = getGeneInfoFull() %>%filter(scGeneName %in% fdata)
rownames(geneInfo ) = geneInfo$scGeneName

geneInfo = geneInfo[rownames(counts),]
pdata = pdata[colnames(counts),]


sce.Major = SingleCellExperiment(list(counts=counts),
    colData=pdata,
    rowData=geneInfo,
    metadata=list(study="GSE111111"))






```




## Expression pattern single cell

```{r }
DefaultAssay(spatial.combined.sct) <- "Spatial"
spatial.combined.sct = ScaleData(spatial.combined.sct)


DoHeatmap(spatial.combined.sct,features = proteinsInSilk$scGeneName,
          assay = "Spatial", group.by ="majorAmpullateZones" , raster = FALSE) + scale_fill_gradientn(colors = c("blue", "white", "red"))

#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure4/Heatmap.zones.genes.spatial.pdf",width = 8, height = 8)


VlnPlot(spatial.combined.sct,features = proteinInfo.filtered$scGeneName,
          assay = "Spatial", group.by ="majorAmpullateZones",pt.size = 0)
#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure4/Violin.zones.genes.spatial.pdf",width = 8, height = 16)

VlnPlot(spatial.combined.sct,features = proteinInfo.filtered$scGeneName,
          assay = "Spatial", group.by ="majorAmpullateZones")
#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure4/Violin.zones.genes.spatial.withPoints.pdf",width = 8, height = 16)

```




## Subset to get better gene selection
```{r subset o}



Lsc.spa.major.annotated@meta.data %>%
  filter(!is.na(majorAmpullateZones)) %>%
  group_by(majorAmpullateZones) %>%
  slice_sample(n =100 ) -> Lsc.spa.major.annotated.subsetmd

Lsc.spa.major.annotated.subset = 
  subset(Lsc.spa.major.annotated,SpotID %in%  Lsc.spa.major.annotated.subsetmd$SpotID)

Lsc.spa.major.annotated.subset <- SCTransform(Lsc.spa.major.annotated.subset, verbose = FALSE, assay = "Spatial")

Lsc.spa.major.annotated.subset <- RunPCA(Lsc.spa.major.annotated.subset, verbose = FALSE)
Lsc.spa.major.annotated.subset <- RunUMAP(Lsc.spa.major.annotated.subset, dims = 1:30, verbose = FALSE)




GlandAnnotation = DimPlot(Lsc.spa.major.annotated.subset, reduction = "umap", group.by = "majorAmpullateZones" , label = TRUE)

DoHeatmap(Lsc.spa.major.annotated.subset,features = proteinInfo.filtered$scGeneName, group.by ="majorAmpullateZones", raster = FALSE ) + scale_fill_gradientn(colors = c("blue", "white", "red")) 
#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure4/Heatmap.zones.genes.spatial.pdf",width = 8, height = 8)


VlnPlot(Lsc.spa.major.annotated.subset,features = proteinInfo.filtered$scGeneName,
          assay = "SCT", group.by ="majorAmpullateZones")
#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure4/Violin.zones.genes.SCT.withPoints.pdf",width = 8, height = 16)


DotPlot(Lsc.spa.major.annotated.subset,features = proteinInfo.filtered$scGeneName,
          assay = "SCT", group.by ="majorAmpullateZones", dot.min = 0.1) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#ggsave("~/Box/bridgeSpider/paper1/main/figures/figure4/Dotplot.zones.genes.SCT.withPoints.pdf",width = 8, height = 16)

FeaturePlot(spatial.combined.sct,features = proteinInfo.filtered$geneName )
```



# Expression analysis 



# Visualise expression of spidroin genes.


# Single slides

```{r adding data to the analysis, include=FALSE}

samplesSpatial = read_rds("~/Box/bridgeSpider/spatialRNA/seurat/EightSamples.individual.RDS")




bulkGeneLists = list()
for(bulkCluster in unique(proteinInfo.filtered$hierarchicalCluster2)){
  proteinInfo.filtered %>% filter(hierarchicalCluster2 == bulkCluster)->
  proteinInfo.filtered.g
  genes = list(proteinInfo.filtered.g$geneName)
  bulkGeneLists[bulkCluster] = genes
}

proteinInfo.filtered$scGeneName
```
## Slide4  figure
```{r}



slideN = samplesSpatial[["Slide4"]]

slideN$SpotID = paste(rownames(slideN@meta.data)) 


md_major %>% inner_join(positions) %>%
  filter(Sample == "Slide4") %>%
  mutate(SpotID = gsub(pattern = paste("Slide4","_", sep = ""),
                       replacement = "",SpotID)) ->
  ampulateInfoSlide


ampulateInfoSlide %>% distinct(SilkGlands)  %>%
  filter(SilkGlands != "unsure") -> glands

slideN = addMetaDataToSpatialData(  slideN ,ampulateInfoSlide )

slideN.major = subset(slideN, majorAmpullateZones %in% c("ZoneA","ZoneC","ZoneB"))

 SpatialPlot(slideN.major,group.by =  "majorAmpullateZones", alpha = 0.9, pt.size.factor = 2, cols = spatialZone)+NoLegend()
 
  ggsave(filename = paste("~/Box/bridgeSpider/paper1/main/figures/",
  slide,"/spatial.slide4.annotaion.geom.point.0.9.pdf", sep = ""),
  width = 5, height = 5)

ggplot(slideN.major@meta.data, aes( x = x))+ geom_density()
 slideN.major.left =  subset(slideN.major, x < 22500)   
 SpatialPlot(slideN.major.left,group.by =  "majorAmpullateZones", alpha = 0.9, pt.size.factor = 5, cols = spatialZone)+NoLegend()
 
SpatialFeaturePlot(slideN.major.left, features = c("MaSp3a") )
 
  ggsave(filename = "~/Box/bridgeSpider/paper1/main/figures/spatial.slide4.left.annotaion.geom.point.0.9.pdf",width = 5, height = 5)
  

for(bulkCluster in names(bulkGeneLists)){ 
slide4.major <- AddModuleScore(object = slideN.major.left, 
                               features = bulkGeneLists[bulkCluster],
                               name =bulkCluster,nbin = 12)

}
sac_bulk_11 = SpatialFeaturePlot(object = slide4.major,
                   features = proteinInfo.filtered$scGeneName,min.cutoff = 'q3')



sac_bulk_11
 ggsave(filename = "~/Box/bridgeSpider/paper1/main/figures/figure4/spatial.slide4.geneExpression.pdf", width = 20, height = 20)
 sac_bulk_12 = SpatialFeaturePlot(object = slide4.major,
                   features = "sac_bulk_21")

sac_tail_1 = SpatialFeaturePlot(object = slide4.major,
                   features = "tail_bulk_11")

sac_tail_2 = SpatialFeaturePlot(object = slide4.major,
                   features = "tail_bulk_21")

sac_tail_4 = SpatialFeaturePlot(object = slide4.major,
                   features = "tail_bulk_41")

sac_tail_1+sac_tail_2+sac_tail_4
proteinInfo.filtered

  slide4.major$sac_bulk_11
 names(bulkGeneLists)  
 
```

#Deconvolution
## Bullk RNA
```{r}
sampleInfo.QC.merged.MAG = read_csv(file = "~/Box/bridgeSpider/bulkRNA/sampleInfo.csv") %>%
  filter(grepl(pattern = "major", x = Tissue))

bulk.data.merged = read_csv(file = "../../data/bulkRNA/All.samples.count.csv") %>%
  column_to_rownames(var = "Geneid") %>%
  select(all_of(sampleInfo.QC.merged.MAG$sample))




bulk.data.merged = bulk.data.merged[geneInfo$gene_id, ]
rownames(bulk.data.merged) = geneInfo$scGeneName

sampleInfo.QC.merged.MAG %>%column_to_rownames("sample") ->
  pdata

eset.bulk <- getESET(as.matrix(bulk.data.merged), 
                     fdata = rownames(bulk.data.merged), 
                     pdata = pdata)




```
### Music
#### Without cluster
```{r}


bulk.mtx = exprs(eset.bulk)

dim(bulk.mtx)
dim(Mousesub.basis$M.theta)
# Estimate cell type proportions
Major.prop = music_prop(bulk.mtx = bulk.mtx,sc.sce = sce.Major,
                        clusters = 'clusterMaAG3',
                        samples = 'orig.ident', verbose = F)

pdf("~/Box/bridgeSpider/paper1/main/figures/figure5/bulk.convolution.MuSiC.prop.pdf")
ComplexHeatmap::Heatmap(Major.prop$Est.prop.weighted, cluster_rows = FALSE)
dev.off()

```

#### With cluster
##### Find markers for subparts 

```{r}
major.tail = subset( major.sc.combined.integrated.sct,  clusterMaAG3 %in% c("Tail_MaSp1","Tail_MaSp2","Tail_SpiCE8"))
major.tail = SetIdent(major.tail, value ="clusterMaAG3")
TailMarkers = FindAllMarkers(object = major.tail, only.pos = TRUE, max.cells.per.ident = 500)

major.sac = subset( major.sc.combined.integrated.sct,  clusterMaAG3 %in% c("Sac_MaSp3","Sac_LASC013329", "Sac_1"))
major.sac = SetIdent(major.sac, value ="clusterMaAG3")
SacMarkers = FindAllMarkers(object = major.sac, only.pos = TRUE, max.cells.per.ident = 500)

major.duct = subset( major.sc.combined.integrated.sct,
                    clusterMaAG3 %in% c('Duct_1', 'Duct_2'))
major.duct = SetIdent(major.duct, value ="clusterMaAG3")
DuctMarkers = FindAllMarkers(object = major.duct, only.pos = TRUE, max.cells.per.ident = 500)
```


```{r  }
Mousesub.basis = music_basis(sce.Major, clusters = 'clusterMaAG3', samples = 'orig.ident', non.zero = TRUE)

d <- dist(t(log(Mousesub.basis$Disgn.mtx + 1e-6)), method = "euclidean")
hc1 <- hclust(d, method = "complete" )

pdf("~/Box/bridgeSpider/paper1/main/figures/figure5/bulk.convolution.MuSiC.designMatrix.pdf")
plot(hc1, cex = 0.6, hang = -1, main = 'Cluster log(Design Matrix)')
dev.off()

d <- dist(t(log(Mousesub.basis$M.theta + 1e-8)), method = "euclidean")
# Hierarchical clustering using Complete Linkage
# hc2 <- hclust(d, method = "complete" )
hc2 <- hclust(d, method ="ward.D2")
# Plot the obtained dendrogram

pdf("~/Box/bridgeSpider/paper1/main/figures/figure5/bulk.convolution.MuSiC.designMatrix.pdf")
plot(hc2, cex = 0.6, hang = -1, main = 'Cluster log(Mean of RA)')
dev.off()
plot(hc2, cex = 0.6, hang = -1, main = 'Cluster log(Mean of RA)')

unique(sce.Major$clusterMaAG3)

clusters.type = list(C1 = c("Tail_MaSp1","Tail_MaSp2","Tail_SpiCE8"),
                     C2 = c("Sac_MaSp3","Sac_LASC013329", "Sac_1"),
                     C3 = c('Duct_1', 'Duct_2'))

cl.type = as.character(sce.Major$clusterMaAG3)
for(cl in 1:length(clusters.type)){
  cl.type[cl.type %in% clusters.type[[cl]]] = names(clusters.type)[cl]
}

sce.Major$clusterType = factor(cl.type, levels = c(names(clusters.type)))

s.mouse = unlist(clusters.type)
s.mouse


IEmarkers = list(unique(TailMarkers$gene),unique(SacMarkers$gene),unique(DuctMarkers$gene))
names(IEmarkers) = c('C1', 'C2', 'C3')



Est.mouse.bulk = music_prop.cluster(bulk.mtx = bulk.mtx,
                                    sc.sce = sce.Major,
                                    group.markers = IEmarkers,
                                    clusters = 'clusterMaAG3',
                                    groups = 'clusterType',
                                    samples = 'orig.ident',
                                    clusters.type = clusters.type)


names(Est.mouse.bulk)

pdf("~/Box/bridgeSpider/paper1/main/figures/figure5/bulk.convolution.MuSiC.cluster.prop.pdf")

ComplexHeatmap::Heatmap(Est.mouse.bulk$Est.prop.weighted.cluster, cluster_rows = FALSE)
dev.off()
```
### SCDC


```{r}

rownames(bulk.data.merged) = geneInfo$scGeneName

bulk.data.merged.norm = getCPM0(bulk.data.merged)

deconv = deconv_simple(
  count.filter.norm =bulk.data.merged.norm ,
  basis.norm = sc.sdcs.basis$basis,
  iter.max = 5000 )


ComplexHeatmap::Heatmap(deconv$prop.est.mvw)

```




## Spatial RNA 
```{r}



spatial.combined.sct = readRDS("~/Box/bridgeSpider/spatialRNA/spatial.all.major.zones.RDS")
DefaultAssay(spatial.combined.sct) <- "integrated"

spatial.combined.sct = SetIdent(spatial.combined.sct, value = "majorAmpullateZones")


levels(spatial.combined.sct)


DimPlot(spatial.combined.sct, reduction = "umap", group.by = "majorAmpullateZones" , label = TRUE)

#markers = FindAllMarkers(object = spatial.combined.sct, assay = "Spatial",only.pos = TRUE) 


#markers  %>%
#  filter(p_val_adj < 0.05)   ->
#  markers.zones
  
#  write_csv(x =markers.zones, file = #"../../results/spatial/zones.markerGenes.csv" )


  
    
  
  
```






### Load probabilites 
```{r}

probabilities = read_csv("../../results/deconvolution/major/CARD.50.csv")
ct.select = colnames(probabilities)[2:9]

probabilities %>% pivot_longer(col = ct.select,names_to = "clusterMaAG3", 
                               values_to = "prob") %>% inner_join(clusterRelationship) ->
  prob.df

singleCellClusters = clusterRelationship$clusterZones

```

#### Overall distrubutions 
```{r }

prob.df %>% select(SpotID,clusterZones,Sample, majorAmpullateZones,prob) %>%
  pivot_wider(names_from = "clusterZones",values_from = "prob") ->
  probabilities



umapCoord <- as.data.frame(Embeddings(object = spatial.combined.sct[["umap"]]))  %>% 
  rownames_to_column(var = "SpotID")

probabilities = inner_join(umapCoord,probabilities)

library(scatterpie)
clusterRelationship

clusterRelationship %>% filter()


radius = 0.1
ggplot() + geom_scatterpie(data = probabilities,  aes(x = UMAP_1, y = UMAP_2, r = radius), cols = singleCellClusters, color = NA) + 
    coord_fixed(ratio = 1) + scale_fill_manual(values = AllColors) 
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure5/spatial.umap.convolution.pieChart.pdf")


probabilities %>% 
  select(SpotID, all_of(singleCellClusters)) %>%
  pivot_longer(cols = -SpotID, values_to = "fraction",names_to = "Cluster") %>%
  inner_join(spatial.combined.sct@meta.data) %>%
  mutate(Cluster = factor(Cluster, levels = singleCellClusters)) ->
  probabilities.annotated

probabilities.annotated %>%
  mutate(Cluster = gsub("ZoneA_Spice8", "ZoneA_SpiCE-LMa3",Cluster)) %>%
  mutate(Cluster = gsub("ZoneC_Spice-LMa1", "ZoneC_SpiCE-LMa1",Cluster)) %>%
  mutate(Cluster = gsub("Sac_1", "ZoneBC",Cluster)) %>%
  mutate(Cluster = factor(Cluster,levels =rev(scOrder))) %>%
  group_by(Cluster, majorAmpullateZones) %>%
  summarise(meanFraction = mean(fraction)) %>%
  ggplot(aes(x =majorAmpullateZones, y = meanFraction , fill = Cluster)) + geom_col() +
  scale_fill_manual(values = AllColors2) +
  theme_light()+ coord_flip()
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure6/spatial.convolution.barplot.pdf")

probabilities.annotated %>%
  mutate(Cluster = gsub("ZoneA_Spice8", "ZoneA_SpiCE-LMa3",Cluster)) %>%
  mutate(Cluster = gsub("ZoneC_Spice-LMa1", "ZoneC_SpiCE-LMa1",Cluster)) %>%
  mutate(Cluster = gsub("Sac_1", "ZoneBC",Cluster)) %>%
  mutate(Cluster = factor(Cluster,levels =rev(scOrder))) %>%
  group_by(Cluster, majorAmpullateZones) %>%
  summarise(meanFraction = mean(fraction)) %>%
  filter(Cluster %in% c("ZoneBC","ZoneB_MaSp3","ZoneC_SpiCE-LMa1")) %>%
  filter(majorAmpullateZones == "ZoneB") %>%
  ggplot(aes(x =majorAmpullateZones, y = meanFraction , fill = Cluster)) + geom_col() +
  scale_fill_manual(values = AllColors2) +
  theme_light()
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/distribution.zoneB.pdf",
        width = 3, height = 1.5)

probabilities.annotated %>%
  mutate(Cluster = gsub("ZoneA_Spice8", "ZoneA_SpiCE-LMa3",Cluster)) %>%
  mutate(Cluster = gsub("ZoneC_Spice-LMa1", "ZoneC_SpiCE-LMa1",Cluster)) %>%
  mutate(Cluster = gsub("Sac_1", "ZoneBC",Cluster)) %>%
  mutate(Cluster = factor(Cluster,levels =rev(scOrder))) %>%
  group_by(Cluster, majorAmpullateZones) %>%
  summarise(meanFraction = mean(fraction)) %>%
  filter(majorAmpullateZones == "ZoneC") %>%
  filter(Cluster %in% c("ZoneBC","ZoneB_MaSp3","ZoneC_SpiCE-LMa1")) %>%
  ggplot(aes(x =majorAmpullateZones, y = meanFraction , fill = Cluster)) + geom_col(position="fill") +
  scale_fill_manual(values = AllColors2) +
  theme_light()
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure7/distribution.zoneC.pdf",
        width = 3, height = 1.5)


probabilities.annotated %>%
  mutate(Cluster = gsub("ZoneA_Spice8", "ZoneA_SpiCE-LMa3",Cluster)) %>%
  mutate(Cluster = gsub("ZoneC_Spice-LMa1", "ZoneC_SpiCE-LMa1",Cluster)) %>%
  mutate(Cluster = gsub("Sac_1", "ZoneBC",Cluster)) %>%
  mutate(Cluster = factor(Cluster,levels =rev(scOrder))) %>%
  group_by(Cluster, majorAmpullateZones) %>%
  mutate(majorAmpullateZones = factor(majorAmpullateZones,c("ZoneC","ZoneB","ZoneA") )) %>%
  summarise(meanFraction = mean(fraction)) %>%
  ggplot(aes(x =majorAmpullateZones, y = meanFraction , fill = Cluster)) + geom_col(width = 0.7) +
  scale_fill_manual(values = AllColors2) +
  theme_light()+ coord_flip() + NoLegend()


ggsave("~/Box/bridgeSpider/paper1/main/figures/figure6/spatial.convolution.barplot.2.pdf", width = 4, height = 2)




  


```

#### Per slide 
```{r}


prob.df %>% select(SpotID,clusterZones,Sample, majorAmpullateZones,prob) %>%
  pivot_wider(names_from = "clusterZones",values_from = "prob") ->
  probabilities
locationID = read_csv(file = "../../annotation/spatialData/coordinates.x.y.all.csv")



left_join(locationID, probabilities) %>%
  filter(!is.na(ZoneA_Spice8)) ->
  probabilities

left_join(locationID, probabilities) %>%
  filter(is.na(ZoneA_Spice8)) ->
  structure
  


library(scatterpie)



radius = 250
slidePlots = list()
for(sample in unique(probabilities$Sample) ){
  probabilities %>% filter(Sample == sample ) ->
    pr.sample
  structure %>% filter(Sample == sample ) ->
    lID.sample
  ggplot(lID.sample, aes(x = x, y = -y, r = radius/5)) +
    geom_point( color = "grey", size =0.1)+
  geom_scatterpie(data = pr.sample,  aes(x = x, y = -y, r = radius), cols = singleCellClusters, color = NA) + 
    coord_fixed(ratio = 1) +
  scale_fill_manual(values = AllColors) + theme_minimal() ->
    sample.plot
  sample.plot
  slidePlots[[sample]] =sample.plot 
  fileName = paste("~/Box/bridgeSpider/paper1/main/figures/figure6/deconvolution",
                   paste(sample,"xycoord.prob.piechart.pdf", sep ="."),
                   sep = "/")
  print(sample.plot)
  ggsave(fileName)
  
}

ggsave("~/Box/bridgeSpider/paper1/main/figures/figure5/spatial.umap.convolution.pieChart.pdf")


probabilities %>% 
  select(SpotID, all_of(ct.select)) %>%
  pivot_longer(cols = -SpotID, values_to = "fraction",names_to = "Cluster") %>%
  inner_join(spatial.combined.sct@meta.data) %>%
  mutate(Cluster = factor(Cluster, levels = singleCellClusters)) ->
  probabilities.annotated

probabilities.annotated %>%
  group_by(Cluster, majorAmpullateZones) %>%
  summarise(meanFraction = mean(fraction)) %>%
  ggplot(aes(x =majorAmpullateZones, y = meanFraction , fill = Cluster)) + geom_col() +
  scale_fill_manual(values = bulkClustersGenes3)
ggsave("~/Box/bridgeSpider/paper1/main/figures/figure5/spatial.convolution.barplot.pdf")


  


left_join(locationID, probabilities) %>%
  filter(!is.na(Tail_SpiCE8))


```
