---
title: "Ampullate glands show a distinct set of SpiCE proteins"
output: html_document
date: "2023-10-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Spider silk, distinguished by its exceptional mechanical properties, is a product of renewable resources, synthesized within fractions of a second under ambient conditions1. Its unique combination of rapid production, sustainability, and superior performance characteristics positions spider silk as a compelling environmentally friendly material for high-performance fiber applications2. To unlock its potential, the development of artificial replicas of the pristine silk fiber is imperative, which necessitates a comprehensive understanding of the bioprocessing of the spider silk fiber.

During the past 400 million years spiders have refined their silk spinning abilities, and today, individual spiders of some species can spin seven different types of silks from specialized glands located in the abdomen (opisthosoma)3(Fig. 1a). These include major ampullate silk for making the radial threads and frame of the web, minor ampullate silk for making the inner spiral, flagelliform silk for making the extensible capture spiral that is coated by sticky droplets of aggregate silk, pyriform silk for making attachment discs, tubuliform silk for making the outer layer of the egg sac, and aciniform silk for making the inner part of the egg sacs and/or swathing prey4. The different silk types are produced in separate sets of glands which have secretory ducts connected to spigots located caudoventrally in the abdomen5. The spider silks are primarily composed of spider silk proteins (spidroins), characterized by a distinctive architectural arrangement that features non-repetitive N- and C-terminal domains flanking an extensive repetitive region6-13. The terminal domains are evolutionarily conserved14 while the nature of the repetitive regions differ between the different spidroin types7-11.

The major ampullate silk is the strongest of the different silks, and the main components of this fiber are called major ampullate spidroins (MaSps)7-13. The repetitive regions of the MaSps are rich in glycine and alanine that organize into typical motifs, which together with clustering of the terminal domains in phylogenetic analyses have led to the classification of MaSps into 5 types (MaSp1-5). MaSp115 typically have poly-Ala blocks interspersed with Gly-rich repeats. MaSp216 repeats are rich in proline and glutamine with abundant Gly-Pro-Gly and Gln-Gln motifs that MaSp1 lacks. MaSp317 is similar to MaSp1 but is void of the Gly-Gly motifs commonly found in MaSp1. In contrast, MaSp4 and MaSp518 repeats lack poly-Ala blocks, but MaSp4 contains Gly-Pro-Gly motifs common to MaSp2 whereas MaSp5 have Gly-Gly motifs like MaSp1. Although recent work has extended our understanding of the spider genome7-11,13,19, the spidroin genes are difficult to assemble because they contain extended stretches of repeats and are often found in tandem arrays in the genome. It therefore appears that previous descriptions of spidroin gene repertoires are incomplete.

The major ampullate silk is produced in paired glands, each consisting of a long tubular tail, a wide sac, and a narrow s-shaped duct20 (Fig. 1b). The lumen of the duct is lined by a cuticular intima, a hard and chitin-rich structure believed to aid in removal of water during the spinning process and to protect against tissue laceration during spinning20,21. Histological studies of the gland have shown that the tail and sac are composed of three types of columnar epithelial cells confined to three different zones (A, B and C)22. Zone A cells are found in the tail and proximal parts of the sac, the middle part of the sac harbors zone B cells, while the most distal part of the sac next to the duct is composed of zone C cells (Fig. 1b). According to immuno-TEM studies performed using an antibody that identified the conserved spidroin N-terminal domain, zone A and B cells produce spidroins while the nature of the secretion produced by zone C cells remains uncharacterized22. Investigations of the major ampullate silk have shown that the fiber has a core-skin structure22-29, wherein the central core contains the spidroins23,30 while the surrounding skin layer is composed of unknown proteins and other components like carbohydrates, glycoproteins and lipids23,26,27. Consequently, the compositions of the secretions from zone A, B and C are largely unknown and the link between the epithelial zones the architecture of the fiber remains to be revealed.

Here, we present a complete spidroin repertoire from an orb weaving spider, Larinioides sclopetarius. Spatial transcriptomics and single-cell RNA (scRNA) sequencing showed that the tail and sac of the major ampullate gland are composed of six silk-producing cell types found in the three distinct zones (A, B and C). This, combined with image analyses of histological sections and proteomics analyses of sequentially dissolved fibers, indicate that the silk fiber is a three-layered structure in which the core is composed of proteins belonging to the MaSp1, 2 and 4 families, the middle layer is dominated by MaSp3, and the outmost layer contains several uncharacterized non-spidroin proteins. Thus, we present a detailed description of the structure and composition of the major ampullate silk gland and fiber, in which we successfully link the layered epithelial secretions in the gland to those in the silk fiber. Our results bridge the long-standing knowledge gaps necessary for the production of biomimetic artificial silk fibers.

```{r cars}
summary(cars)
```

## Results


### Major and minor ampullate gland zones have common gene sets 
#### 


  