---
title: "Bulk RNA analysis"
author: "Johan"
date: "05/26/2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    df_print: paged

params:
  RfunctionsDirectory: /Users/johanreimegard/git/RNAmappingPipeline/R
  
  
---






```{r setup, include=FALSE}

# if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")#

# BiocManager::install("ropls")

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
#install.packages("kableExtra")
#library("viridis")     

library(RColorBrewer)
#library(GenomicFeatures)

#library(ropls)

library(ggforce)
library(patchwork)


#install.packages("PTXQC")
library(PTXQC)
rename <- dplyr::rename


source("../loadGeneInfo.R")
source("../colorSchemes.R")
summarise = dplyr::summarise               
select = dplyr::select
group_by = dplyr::group_by
rename = dplyr::rename
library(ggrepel)

#install.packages("GGally")
library("GGally")


repositoryPath = "~/git/Rising-Lab/MajorGland_paper"


```


## Functions


## Parameters
```{r, message=FALSE, warning=FALSE}

minExpressionCutoff =20
DataGenerated = TRUE

```
## Reference
```{r}

geneInfo = read_csv("../../annotation/Lsc.v1.0.3.geneInfo.csv.gz")


geneInfo %>% select(geneName,geneNameProtein) %>%
  filter(grepl("MaSp", geneName)) %>%
  mutate(geneNameProtein = geneName) %>%
  mutate(class = "MaSp") -> MaSpProteins

geneInfo %>% select(geneName,geneNameProtein) %>%
  filter(grepl("MiSp", geneName)) %>%
  mutate(geneNameProtein = geneName) %>%
  mutate(class = "MiSp") -> MiSpProteins

geneInfo %>% select(geneName,geneNameProtein) %>%
  filter(grepl("FlSp", geneName)) %>%
  mutate(geneNameProtein = geneName) %>%
  mutate(class = "FlSp") -> FlSpProteins

geneInfo %>% select(geneName,geneNameProtein) %>%
  filter(grepl("SpiCE", geneNameProtein)) %>%
  mutate(class = "SpiCE") -> spiceProteins

geneInfo %>% select(geneName,geneNameProtein) %>%
  filter(grepl("AmSp-like", geneName)) %>%
  mutate(geneNameProtein = geneName) %>%
  mutate(class = "AmSp-like") -> AmSpProteins


silkGenes = bind_rows(spiceProteins, AmSpProteins,MaSpProteins,MiSpProteins)



GeneOntolyAnnotation = read_tsv("../../annotation/Lsc.v1.0.Makerids.GOterm.tsv.gz") %>%
  filter(GO != "NIL") %>%
  separate_rows(GO, sep = ",") %>%
  rename(GOterm = "GO") %>%
  rename(gene_id = "makerid") %>%
  inner_join(read_tsv("../../annotation/functional/GOterms.description.tsv.gz")) 

    



geneInfo %>%
 mutate(Term = factor(Term, levels = c("MiSp","AmSp-like","Spidroins","SpiCE","SignalPeptide","RibosomalProteins","NoSignalPeptide")) ) ->
  geneInfo



  GOI_minor = data.frame(geneNameProtein2 = c("MiSpa","MiSpb","MiSpc","MiSpd",
                                             "LASC010012","sp-1339#prd34368.1_3"), 
                         specificity ="Minor",
                         zone =c("ZoneA","ZoneA","ZoneB",
                                 "ZoneA","ZoneC","ZoneC"))
  
  GOI_major = data.frame(geneNameProtein2 = c("MaSp1a","MaSp1b","MaSp1c",
                                             "MaSp2a","MaSp2b","MaSp2c",
                                             "MaSp2e","MaSp2f",
                                             "MaSp3a","MaSp3b",
                                              "SpiCE-LMa10","SpiCE-LMa8"),
                          specificity ="Major",
                         zone =c("ZoneA","ZoneA","ZoneA",
                                 "ZoneA","ZoneA","ZoneA",
                                 "ZoneA","ZoneA",
                                 "ZoneB","ZoneB",
                                 "ZoneC","ZoneC"))     
GOI_common = data.frame(geneNameProtein2 = c(
  "LASC007034","lman2l","SpiCE-LMa13","SpiCE-LMa5",
  "AmSp-like1","AmSp-like2","SpiCE-LMa3",
  "SpiCE-LMa1","SpiCE-LMa2","SpiCE-LMa4",
  "SpiCE-LMa11","spa","LASC007099","SpiCE-LMa9","mttp_2"),
                        specificity ="Common", 
  zone = c("ZoneA","ZoneA","ZoneA","ZoneA",
               "ZoneB","ZoneB","ZoneB",
               "ZoneC","ZoneC","ZoneC",
               "ZoneC","ZoneC","ZoneC","ZoneC","ZoneC"))
  

  
GOIs = bind_rows(GOI_minor, GOI_major, GOI_common) %>%
  left_join(geneInfo)
  
geneInfo %>% filter(grepl("SpiCE-LMa", geneNameProtein2))



```



## Proteomics


### Read in the data

#### Minor

```{r}

MQ_data.minor = read_tsv("../../data/proteomics/minorSilk/txt/proteinGroups.txt")


colnames(MQ_data.minor)

MQ_data.minor %>% select(1,2,106:117) -> MQ_data.minor.intensity 


MQ_data.minor.intensity %>%
  rename(Mi_Dope_1 = "LFQ intensity Dope_spider_1") %>%
  rename(Mi_Dope_2 = "LFQ intensity Dope_spider_2") %>%
  rename(Mi_Dope_3 = "LFQ intensity Dope_spider_3") %>%
  rename(Mi_Dope_4 = "LFQ intensity Dope_spider_4") %>%
  rename(Mi_Gland_1 = "LFQ intensity Gland_spider_1") %>%
  rename(Mi_Gland_2 = "LFQ intensity Gland_spider_2") %>%
  rename(Mi_Gland_3 = "LFQ intensity Gland_spider_3") %>%
  rename(Mi_Gland_4 = "LFQ intensity Gland_spider_4") %>%
  rename(Mi_Gland_5 = "LFQ intensity Intact_Mi_Gland_spider_5") %>%
  rename(Mi_Gland_6 = "LFQ intensity Intact_Mi_Gland_spider_6") %>%
  rename(Mi_Silk_1 = "LFQ intensity Silk_Mi_1") %>%
  rename(Mi_Silk_2 = "LFQ intensity Silk_Mi_2") %>%
  rename(geneName = "Protein IDs") ->
  MQ_data.minor.intensity
colnames(MQ_data.minor.intensity)



MQ_data.minor.intensity %>% filter(Mi_Silk_2 != 0 | Mi_Silk_1 != 0)  %>%
  select(1,13,14) %>% 
 inner_join(geneInfo) %>%
  select(geneName,Mi_Silk_2) %>%
  mutate(sum = sum(Mi_Silk_2)) %>%
  mutate(relative = Mi_Silk_2 / sum) %>%
  arrange(-relative*100)

  




```

x#### Major

```{r}

MQ_data.major = read_tsv("../../data/proteomics/majorSilk/txt/proteinGroups.txt")


colnames(MQ_data.major)

MQ_data.major %>% select(1,2,120:133) %>%
  rename(Ma_Dope_2 = "LFQ intensity MA_dope_2") %>%
  rename(Ma_Dope_3 = "LFQ intensity MA_dope_3") %>%
  rename(Ma_Gland_1 = "LFQ intensity MA_gland_1") %>%
  rename(Ma_Gland_2 = "LFQ intensity MA_gland_2") %>%
  rename(Ma_Gland_3 = "LFQ intensity MA_gland_3") %>%
  rename(Ma_SilkHFIP_1 = "LFQ intensity Native_Silk_HFIP_1") %>%
  rename(Ma_SilkHFIP_2 = "LFQ intensity Native_Silk_HFIP_2") %>%
  rename(Ma_SilkHFIP_3 = "LFQ intensity Native_Silk_HFIP_3") %>%
  rename(Ma_SilkLiBr_1 = "LFQ intensity Native_Silk_LiBr_1") %>%
  rename(Ma_SilkLiBr_2 = "LFQ intensity Native_Silk_LiBr_2") %>%
  rename(Ma_SilkLiBr_3 = "LFQ intensity Native_Silk_LiBr_3") %>%
  rename(Ma_Silk_1 = "LFQ intensity Native_Silk_Urea_1") %>%
  rename(Ma_Silk_2 = "LFQ intensity Native_Silk_Urea_2") %>%
  rename(Ma_Silk_3 = "LFQ intensity Native_Silk_Urea_3") %>%
  rename(geneName = "Protein IDs") %>%
  select(geneName,Ma_Dope_2,Ma_Dope_3,
         Ma_Gland_1,Ma_Gland_2,Ma_Gland_3,
         Ma_SilkHFIP_1,Ma_SilkHFIP_2,Ma_SilkHFIP_3,
         Ma_SilkLiBr_1,Ma_SilkLiBr_2,Ma_SilkLiBr_3,
         Ma_Silk_1,Ma_Silk_2,Ma_Silk_3,)->
  MQ_data.major.intensity


MQ_data.major.intensity  %>%
  inner_join(geneInfo) %>%
  right_join(GOIs) 


MQ_data.minor.intensity  %>%
  inner_join(geneInfo) %>%
  right_join(GOIs) %>% filter(specificity != "Major") 


```


### QC 



#### Minor 

removing the contaminants and low intensity proteins and one bad sample. 
Converting to relative abundance
Averaging the intensity for the genenames that comes from multiple proteins. 


```{r}

MQ_data.minor.intensity %>% select(1,3,4,5,6,7,8,9,10,11,12,13,14) %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE)->
  MQ_data.MiSp


MQ_data.MiSp%>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Part")



MQ_data.MiSp %>% mutate(Type = "Other") %>%
  mutate(Type = replace(Type, grepl("MiSp",geneName), "MiSp")) %>%
  mutate(Type = replace(Type, grepl("CON_",geneName), "Contaminants")) %>%
  mutate(Type = replace(Type, grepl("REV_",geneName), "FALSE"))->
  MQ_data.MiSp

MQ_data.MiSp %>%
  filter(intensity > 10000) %>%
  filter(Type %in% c("Other", "MiSp")) %>%
  filter(Sample != "Mi_Silk_1")->
  MQ_data.MiSp.QC.all
  
MQ_data.MiSp.QC.all %>% filter(spider <=4) -> MQ_data.MiSp.QC 

MQ_data.MiSp.QC %>% group_by(Gland,spider,Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.MiSp.QC) %>%
  mutate(relativeIntensity = intensity / TotalIntensity) ->
  MQ_data.MiSp.QC


MQ_data.MiSp.QC %>% 
  ungroup() %>%
  distinct(geneName) %>%
  inner_join(MQ_data.MiSp.QC) %>%
  mutate(geneID = 1:n()) %>%
  separate_rows(geneName, sep = ";") %>%
  select(geneID,geneName,Gland,spider,Part,Sample,Type,relativeIntensity)->
  MQ_data.MiSp.QC.temp

MQ_data.MiSp.QC.temp %>%
  group_by(geneID) %>%
  summarise(nrOfGenes= n(), factor = 1/nrOfGenes) %>%
  inner_join(MQ_data.MiSp.QC.temp) %>%
  mutate(relativeIntensity = relativeIntensity * factor) ->
  MQ_data.MiSp.QC


  

```




#### Summary of the sample

```{r}


MQ_data.MiSp.QC  %>% ggplot(aes(x = Part, y = relativeIntensity, fill = Type)) + geom_col() +facet_wrap("spider")



MQ_data.MiSp.QC %>%
  mutate(Type2 = Type) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpa","MiSpa")) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpb","MiSpb")) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpc","MiSpc")) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpd","MiSpd"))->
  MQ_data.MiSp.QC




MQ_data.MiSp.QC %>% 
  left_join(geneInfo) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope","Silk"))) ->
  MQ_data.MiSp.QC.summary.full



MQ_data.MiSp.QC %>% ungroup() %>% 
  mutate(log10RI = log10(relativeIntensity)) %>%
  left_join(geneInfo) %>%
  select(Sample,geneName,log10RI,Term) %>%
  pivot_wider(names_from = "Sample", values_from = "log10RI") %>%
  column_to_rownames("geneName")->
  MQ_data.MiSp.QC.wide

```


##### Sfig prot distribution
```{r}

MQ_data.MiSp.QC.summary.full %>% filter(spider == 2) %>% 
  select(Sample, geneNameProtein,relativeIntensity, Term) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity", values_fill = 0.000001) %>%
  filter(Mi_Gland_2>0.000001 | Mi_Dope_2>0.000001) ->
  spider2.data
  spider2.data %>% filter(Mi_Gland_2>0.000001 & Mi_Dope_2>0.000001) %>%
    filter(!Term %in% c("SignalPeptide","NoSignalPeptide")) %>%
    filter(geneNameProtein != "SpiCE-LMa13",
           geneNameProtein != "SpiCE-LMa7") ->
    spider2.data.common
  
  ggplot(data = spider2.data,aes(x = Mi_Gland_2, y = Mi_Dope_2, label = geneNameProtein, color = Term )) +
    geom_point()+
    geom_text_repel(data = spider2.data.common)+
    ggtitle("Sample2 ")+
    xlab("Gland")+ ylab("Dope")+
    scale_x_log10() +  # Apply log scale to x-axis
    scale_y_log10()+
    theme_light() +
    scale_color_brewer(palette = "Set2")+theme(legend.position = "none")  -> sample2
  
 MQ_data.MiSp.QC.summary.full %>%
    filter(Part != "Silk") %>%
  ggplot(aes(x = spider, y = relativeIntensity,fill = Term)) +
  geom_bar(stat = "identity",position = "fill") + facet_row("Part", shrink = TRUE) +
   ylab("Fraction of Intensity") +
   theme_light() +
   scale_fill_brewer(palette = "Set2") ->
  figure1.supp
 
 (figure1.supp +theme(legend.position = "none"))+ sample2 + plot_layout(widths = c(1.5,2)) + plot_annotation(tag_levels = "A")  -> suppFig2
 
 suppFig2
 
  
ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/supplementaryFigureProt2.pdf",width = 160,height = 140,units = "mm",plot = suppFig2)




```
##### S.fig ggpairs
```{r}
 
samples = c("Mi_Gland_1","Mi_Gland_2","Mi_Gland_3","Mi_Gland_4","Mi_Dope_2","Mi_Dope_3","Mi_Dope_4")

samples.gland = c("Mi_Gland_1","Mi_Gland_2","Mi_Gland_3","Mi_Gland_4")

ggpairs(MQ_data.MiSp.QC.wide[,samples.gland],
        mapping = aes(color = MQ_data.MiSp.QC.wide$Term),
        upper = "blank",diag = "blank") +
            theme_light() +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2")   -> suppFigProt1.gland
  
ggpairs(MQ_data.MiSp.QC.wide[,samples.gland],
        lower = "blank")+
    theme_light() +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2") -> suppFigProt2.gland

suppFigProt1<- suppFigProt1.gland + theme(
  axis.text = element_text(size = 7),  # Axis labels
  axis.title = element_text(size = 8), # Axis titles
  legend.text = element_text(size = 7), # Legend labels
  legend.title = element_text(size = 8), # Legend title
  strip.text = element_text(size = 8) # Panel labels
)

suppFigProt2<- suppFigProt2.gland + theme(
  axis.text = element_text(size = 7),  # Axis labels
  axis.title = element_text(size = 8), # Axis titles
  legend.text = element_text(size = 7), # Legend labels
  legend.title = element_text(size = 8), # Legend title
  strip.text = element_text(size = 8) # Panel labels
)

ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/protGGpairs1.gland.pdf",width = 160,height = 160,units = "mm",plot = suppFigProt1.gland)
ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/protGGpairs2.gland.pdf",width = 160,height = 160,units = "mm",plot = suppFigProt2.gland)

samples.dope = c("Mi_Dope_1","Mi_Dope_2","Mi_Dope_3","Mi_Dope_4")


MQ_data.MiSp.QC.wide %>% filter(!Term %in% c("NoSignalPeptide"))  ->
  MQ_data.MiSp.QC.wide.dope.signelPeptide
 
ggpairs(MQ_data.MiSp.QC.wide[,samples.dope],
        mapping = aes(color = MQ_data.MiSp.QC.wide$Term),
        upper = "blank",
        diag = "blank") +
    theme_light() +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2")   -> suppFigProt1.dope
  
ggpairs(MQ_data.MiSp.QC.wide[,samples.dope],
        lower = "blank",
        diag = list(continuous = wrap("densityDiag",adjust = 0.5, alpha = 0.7)) )+
    theme_light() +

    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2") -> suppFigProt2.dope

suppFigProt1.dope<- suppFigProt1.dope + theme(
  axis.text = element_text(size = 7),  # Axis labels
  axis.title = element_text(size = 8), # Axis titles
  legend.text = element_text(size = 7), # Legend labels
  legend.title = element_text(size = 8), # Legend title
  strip.text = element_text(size = 8) # Panel labels
)

suppFigProt2.dope<- suppFigProt2.dope + theme(
  axis.text = element_text(size = 7),  # Axis labels
  axis.title = element_text(size = 8), # Axis titles
  legend.text = element_text(size = 7), # Legend labels
  legend.title = element_text(size = 8), # Legend title
  strip.text = element_text(size = 8) # Panel labels
)

ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/protGGpairs1.dope.pdf",
       width = 160,height = 160,units = "mm",
       plot = suppFigProt1.dope)
ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/protGGpairs2.dope.pdf",
       width = 160,height = 160,units = "mm",
       plot = suppFigProt2.dope)


```


```{r}
  
  

MQ_data.MiSp.QC %>%
  group_by(Part,geneName)  %>%
  summarise(numberOfSamples = n(), averageIntensity = sum(relativeIntensity)/numberOfSamples)%>%
  left_join(geneInfo) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope","Silk"))) %>%
  arrange(-averageIntensity)  ->
  MQ_data.MiSp.QC.summary


 MQ_data.MiSp.QC.summary %>%
   group_by(Part,Term) %>%
   summarise(averageIntensity= sum(averageIntensity)) %>%
  ggplot(aes(x = Part, y = averageIntensity,fill = Term)) +
  geom_bar(stat = "identity",position = "fill") +
   ylab("Fraction of Intensity") +
   theme_light() +
   scale_fill_brewer(palette = "Set2") ->
  figure1a
 

  
 
 

 
MQ_data.MiSp.QC.wide %>%filter(!Term %in% c("NoSignalPeptide","RibosomalProteins")) %>%
  select(all_of(samples)) %>%
  ggpairs()+
  ggtitle("Only proteins from genes with an signal peptide")+
    theme_light() +
    scale_color_brewer(palette = "Set2") +
    scale_fill_brewer(palette = "Set2") -> suppFigProt3

ggsave("figures/supplementaryFigures/supplementaryFigureProt3.pdf",width = 160,height = 160,units = "mm",plot = suppFigProt3)


MQ_data.MiSp.QC.summary.full %>%
  ggplot(aes(x = Part, y = relativeIntensity,fill = Term)) +
  geom_bar(stat = "identity",position = "fill") +
   ylab("Fraction of Intensity") +
   theme_light() +
   scale_fill_brewer(palette = "Set2") ->
  figure1.supp


  figure1a
  ggsave("figures/figure2/A.protDist.pdf",height = 2, width = 4)

  
MQ_data.MiSp.QC.summary.full %>%
  filter(Sample != "Mi_Dope_1") %>%
  group_by(Part,geneName,Gland) %>%
  summarise(numberOfSamples = n(), averageIntensity = sum(relativeIntensity)/3)  %>%
  filter( numberOfSamples >= 2) %>% filter(Part == "Dope") %>%
  inner_join(geneInfo) %>%
  filter(signalP == "Yes") %>%
  arrange(-averageIntensity)
  
MQ_data.MiSp.QC.summary.full %>%
  group_by(Part,geneName,Gland) %>%
  summarise(numberOfSamples = n(), averageIntensity = sum(relativeIntensity)/4)  %>%
  filter( numberOfSamples >= 1) %>% filter(Part != "Dope") %>%
  inner_join(geneInfo) %>%
  filter(signalP == "Yes") %>%
  arrange(-averageIntensity)

MQ_data.MiSp.QC.summary.full %>% filter(geneName %in% c("MiSpc")) 
  
```

#### Major
```{r}

MQ_data.major.intensity %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE)->
  MQ_data.MaSp


MQ_data.MaSp%>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Part")



MQ_data.MaSp %>% mutate(Type = "Other") %>%
  mutate(Type = replace(Type, grepl("MaSp1",geneName), "MaSp1")) %>%
  mutate(Type = replace(Type, grepl("MaSp2",geneName), "MaSp2")) %>%
  mutate(Type = replace(Type, grepl("MaSp3",geneName), "MaSp3")) %>%
  mutate(Type = replace(Type, grepl("MaSp4",geneName), "MaSp4")) %>%
  mutate(Type = replace(Type, grepl("CON_",geneName), "Contaminants")) %>%
  mutate(Type = replace(Type, grepl("REV_",geneName), "FALSE"))->
  MQ_data.MaSp

MQ_data.MaSp %>% filter(intensity > 10000) %>%
  filter(Type != "Contaminants") %>%
  filter(Type != "FALSE") %>%
  group_by(Gland,spider,Part,Type) %>%
  summarise(n = n()) %>%
  pivot_wider(values_from = "n", names_from = "Type")





MQ_data.MaSp %>%
  filter(intensity > 1000000) %>%
  filter(Type != "Contaminants") %>%
  filter(Type != "FALSE")->
  MQ_data.MaSp.QC




MQ_data.MaSp.QC %>% group_by(Gland,spider,Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.MaSp.QC) %>%
  mutate(relativeIntensity = intensity / TotalIntensity) ->
  MQ_data.MaSp.QC

MQ_data.MaSp.QC %>% ggplot(aes(x = Part, y = relativeIntensity, fill = Type)) + geom_col() +facet_wrap("spider")



MQ_data.MaSp.QC %>% ungroup() %>%
  select(Sample, geneName,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity") %>% column_to_rownames("geneName")   ->MQ_data.MaSp.QC.wide




samples = MQ_data.MaSp.QC %>% distinct(Sample) %>% arrange(Sample)
ggpairs(log10(MQ_data.MaSp.QC.wide[,samples$Sample[1:5]]))

MaGlands = c("Ma_Gland_1","Ma_Gland_3")
ggpairs(log10(MQ_data.MaSp.QC.wide[,samples$Sample[1:5]]))
ggpairs(log10(MQ_data.MaSp.QC.wide[,MaGlands]))

MQ_data.MaSp.QC.wide

ggpairs(log10(MQ_data.MaSp.QC.wide[,samples$Sample[6:14]]))

```

##### Check Major silk for different proteins identified
```{r check for silk proteins} 

MQ_data.MaSp.QC %>%
  filter(grepl("Silk",Part)) %>%
  group_by(geneName) %>%
  summarise(nrOfSamples = n(), averageIntensity = sum(relativeIntensity)/9)  %>%
    separate_rows(geneName, sep = ";") %>%
  inner_join(geneInfo) ->
 MQ_data.MaSp.QC.silk.summary 
  


MQ_data.MaSp.QC.silk.summary %>%
  filter(geneNameProtein2 %in% GOI_common$geneNameProtein2) %>%
  arrange(-averageIntensity)%>% 
    mutate(silkGroup = "Common") -> silk.common.summary
 
  MQ_data.MaSp.QC.silk.summary %>%
  filter(geneNameProtein2 %in% GOI_minor$geneNameProtein2 ) %>%
  arrange(-averageIntensity)%>% 
    mutate(silkGroup = "Minor") -> silk.minor.summary

    MQ_data.MaSp.QC.silk.summary %>%
  filter(geneNameProtein2 %in% GOI_major$geneNameProtein2) %>%
  arrange(-averageIntensity) %>% 
    mutate(silkGroup = "Major") -> silk.major.summary

    MQ_data.MaSp.QC.silk.summary %>%
  filter(grepl("MaSp",geneNameProtein)) %>%
  arrange(-averageIntensity) %>% 
    mutate(silkGroup = "Major") -> silk.major.masp.summary
 
          
    
bind_rows(silk.common.summary,silk.minor.summary,silk.major.summary,silk.major.masp.summary) %>%
  filter(nrOfSamples>2) %>%
  select(geneName,geneNameProtein,averageIntensity,silkGroup) %>%
  arrange(-averageIntensity) %>% arrange(silkGroup) ->
  MQ_data.MaSp.QC.silk.summary.all
  
MQ_data.MaSp.QC.silk.summary %>%
  filter(!geneNameProtein %in%  MQ_data.MaSp.QC.silk.summary.all$geneNameProtein) %>%
  filter(signalP == "Yes" )%>%
  filter(nrOfSamples>2) %>%
  arrange(lascID) %>% 
    mutate(silkGroup = "Other") %>%
  select(geneName,geneNameProtein,averageIntensity,silkGroup) -> silk.other.summary

bind_rows(MQ_data.MaSp.QC.silk.summary.all,silk.other.summary) %>%
  group_by(silkGroup) %>%
  summarise(averageIntensity = sum(averageIntensity)) 
    
MQ_data.MaSp.QC %>%
  filter(grepl("Silk",Part)) %>%
  group_by(geneName,Part) %>%
  summarise(nrOfSamples = n(), averageIntensity = sum(relativeIntensity)/3)  %>%
    separate_rows(geneName, sep = ";") %>%
  filter(nrOfSamples > 1) %>%
  inner_join(geneInfo) %>%
  filter(geneNameProtein %in% GOIs$geneNameProtein) %>%
  ungroup() %>%
  distinct(geneNameProtein,Term,geneName)  %>%
  arrange(geneNameProtein)


geneInfo %>% filter(grepl("SpiCE",geneNameProtein))
  
MQ_data.MaSp.QC %>% group_by(geneName,Part) %>%
  summarise(relativeIntensity = mean(relativeIntensity)) %>%
  filter(relativeIntensity > 0.0001) %>% mutate(gland = "Major") -> 
MQ_data.MaSp.QC.summary

MQ_data.MaSp.QC %>% 
  ungroup() %>%
  distinct(geneName) %>%
  inner_join(MQ_data.MaSp.QC) %>%
  mutate(geneID = 1:n()) %>%
  separate_rows(geneName, sep = ";") %>%
  select(geneID,geneName,Gland,spider,Part,Sample,Type,relativeIntensity)->
  MQ_data.MaSp.QC.temp

MQ_data.MaSp.QC.temp %>%
  group_by(geneID) %>%
  summarise(nrOfGenes= n(), factor = 1/nrOfGenes) %>%
  inner_join(MQ_data.MaSp.QC.temp) %>%
  mutate(relativeIntensity = relativeIntensity * factor) ->
  MQ_data.MaSp.QC



MQ_data.MaSp.QC %>%
  filter(grepl("Silk",Part)) %>%
  group_by(geneName) %>%
  summarise(nrOfSamples = n(), averageIntensity = sum(relativeIntensity)/9)  %>%
    separate_rows(geneName, sep = ";") %>%
  inner_join(geneInfo) ->
 MQ_data.MaSp.QC.silk.summary

MQ_data.MaSp.QC.silk.summary %>% arrange(-nrOfSamples)



MQ_data.MaSp.QC.silk.summary.all %>%
  group_by(silkGroup) %>%
  summarise(averageIntensity = sum(averageIntensity)) %>%
  mutate(percent = averageIntensity / sum(averageIntensity)*100) %>%
  select(silkGroup,percent)


MQ_data.MaSp.QC.silk.summary.all %>%
  filter(silkGroup == "Minor") %>%
  arrange(-averageIntensity) %>%
  mutate(percent = averageIntensity / sum(averageIntensity)*100) %>%
  select(geneNameProtein,percent)

```



#### Major ampullate gland 

```{r}

MQ_data.MaSp.QC %>%
  filter(Sample %in% c("Ma_Gland_1","Ma_Gland_3")) %>%
  separate_rows(geneName, sep = ";") %>%
  inner_join(geneInfo) %>%
  group_by(geneName) %>%
  summarise(nrOfSamples = n(), averageIntensity = sum(relativeIntensity)/2) %>%
  mutate(averageIntensity = round(digits = 5)) ->
 MQ_data.MaSp.QC.Masp.summary 


  

MQ_data.MaSp.QC.Masp.summary %>% 
  filter(averageIntensity > 0.00005)  %>%
  arrange(-averageIntensity) ->
  MQ_data.MaSp.QC.gland.QC.summary


MQ_data.MaSp.QC %>%
  filter(Sample %in% c("Ma_Gland_1","Ma_Gland_3")) %>%
  filter(geneName %in% MQ_data.MaSp.QC.gland.QC.summary$geneName) %>%
  inner_join(geneInfo) %>%
  select(geneNameProtein,Sample,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity", values_fill = 0) ->
  MQ_data.MaSp.QC.gland.QC.wide

MQ_data.MaSp.QC.gland.QC.summary %>%
  select(geneNameProtein,averageIntensity) %>%
  inner_join(MQ_data.MaSp.QC.gland.QC.wide) %>%
  select(geneNameProtein, averageIntensity, Ma_Gland_1,Ma_Gland_3) ->
  Supp.table.7


MQ_data.MaSp.QC.gland.QC.summary %>%
  averageIntensity > 0.0001 %>%
    inner_join(geneInfo) %>%
  filter(grepl("SpiCE",geneNameProtein))

```





 
## Saved for analysis with RNA data 
```{r }
MQ_data.MiSp.QC.summary


MQ_data.MiSp.QC %>%
  group_by(Part,geneName)  %>%
  summarise(numberOfSamples = n(), averageIntensity = sum(relativeIntensity)/numberOfSamples) %>%
  group_by(geneName) %>%
  summarise(max = max(averageIntensity)) %>%
  filter(max > 0.0001) %>% arrange(-max) ->
  MQ_data.MiSp.QC.expressed



write_csv(MQ_data.MiSp.QC.expressed,"../../data/proteomics/minor/MQ_data.MiSp.QC.expressed.csv")

write_csv(MQ_data.MiSp.QC,"../../results/proteomics/minor/minorAmpullateEpxression.csv")


MQ_data.MiSp.QC.expressed %>%
  inner_join(geneInfo ) %>% right_join(GOIs)



```







# Comparison


## Using Normalised intensities
```{r}

Proteomics.data = bind_rows( MQ_data.MaSp.QC,MQ_data.MiSp.QC) %>%
  inner_join(geneInfo)%>% select(geneNameProtein, Gland,Part,spider,relativeIntensity)
Proteomics.data %>% ggplot(aes(x = log10(relativeIntensity), color = Part)) + geom_density()+facet_wrap("Gland")


GOIs %>%
  mutate(specificity = factor(specificity, levels = c("Major","Common","Minor")))  %>%
  arrange(zone, specificity,geneNameProtein2)->
  GOIs.order
```

### Supp fig. Prot 3 


```{r}


Proteomics.data %>%
  inner_join(GOIs.order) %>%
  filter(Part %in% c("Gland","Dope")) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope"))) %>%
    select(geneNameProtein, Gland, Part, spider, relativeIntensity, specificity, zone) %>%
  pivot_wider(names_from = c("Gland","spider"), values_from = "relativeIntensity", values_fill = 0.00001) %>%
  pivot_longer(cols = starts_with("Mi") | starts_with("Ma"),
    names_to = c("Sample"),
    values_to =  "relativeIntensity"
  ) %>%
  separate(Sample, into = c( "Gland", "spider"), remove = FALSE)%>%
  mutate(specificity = factor(specificity, levels = c("Major","Common","Minor"))) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = rev(GOIs.order$geneNameProtein))) %>%
  ggplot(aes(x = geneNameProtein, y = relativeIntensity * 100,  color = Gland)) +
 geom_jitter(
    aes(color = Gland), # Inherit color from Gland for consistency
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), # Use jitterdodge
    size = 1.5, # Adjust point size
    alpha = 0.7, # Adjust point transparency
  ) +# Rotate the coordinate system by 90 degrees
  coord_flip() +
  facet_grid(specificity~Part, scales = "free_y",space = "free_y") +
  scale_y_continuous(trans = "log10") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    y = "Percent abundance (%)",
    fill = "Gland"
  ) -> 
  proteomics.dope.plot


ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/supplementaryFigureProt3.png",width = 160,height = 180,units = "mm",plot = proteomics.dope.plot)


```

### Supp fig. Prot 4 


```{r}
Proteomics.data %>%
  inner_join(GOIs.order) %>%
  filter(!Part %in% c("Gland","Dope")) %>%

    select(geneNameProtein, Gland, Part, spider, relativeIntensity, specificity, zone) %>%
  pivot_wider(names_from = c("Gland","spider"), 
              values_from = "relativeIntensity", values_fill = 0.00001) %>%
  pivot_longer(cols = starts_with("Mi") | starts_with("Ma"),
    names_to = c("Sample"),
    values_to =  "relativeIntensity"
  ) %>%
  separate(Sample, into = c( "Gland", "spider"), remove = FALSE)%>%
  mutate(specificity = factor(specificity, levels = c("Major","Common","Minor"))) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = rev(GOIs.order$geneNameProtein))) %>%
  filter(Gland == "Ma")  ->
  Proteomics.data.MA.silk


Proteomics.data.MA.silk %>% 
  filter(Part == "Silk") %>%
  select(-Sample) %>%
   right_join(GOIs) %>%
  select(geneNameProtein, specificity,Gland,Part,spider,relativeIntensity) %>%
  mutate(Gland = "Ma", Part = "Silk") %>%
  mutate(spider = replace(spider, is.na(spider),1)) %>%
  mutate(relativeIntensity = replace(relativeIntensity, is.na(relativeIntensity),0.00001)) %>%
    pivot_wider(names_from = c("Gland","spider"), values_from = "relativeIntensity", values_fill = 0.00001) %>%
  select(geneNameProtein, specificity,Part,Ma_1,Ma_2,Ma_3) %>%
  pivot_longer(cols = starts_with("Mi") | starts_with("Ma"),
    names_to = c("Sample"),
    values_to =  "relativeIntensity"
  ) %>%
  separate(Sample, into = c( "Gland", "spider"), remove = FALSE)  %>%
  mutate(geneNameProtein = factor(geneNameProtein,
                                  levels = rev(GOIs.order$geneNameProtein)))  %>%
  mutate(specificity = factor(specificity, levels = c("Major","Common","Minor")))->
  Proteomics.data.MA.silk

  
Proteomics.data.MA.silk %>%ggplot(aes(x = geneNameProtein, y = relativeIntensity * 100,  color = Gland)) +
 geom_jitter(
    aes(color = Gland), # Inherit color from Gland for consistency
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), # Use jitterdodge
    size = 1.5, # Adjust point size
    alpha = 0.7, # Adjust point transparency
  ) +# Rotate the coordinate system by 90 degrees
  coord_flip() +
  facet_grid(specificity~Part, scales = "free_y",space = "free_y") +
  scale_y_continuous(trans = "log10", labels = scales::label_number()) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    y = "Relative intensity (%) (log10)",
    fill = "Gland"
  ) ->
  proteomics.silk.plot

proteomics.silk.plot
ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/supplementary/figures/protein/supplementaryFigureProt4.png",width = 160,height = 140,units = "mm",plot = proteomics.silk.plot)


```


### Figure 8B
```{r}
Proteomics.data %>%
  inner_join(GOIs.order) %>%
  filter(Part %in% c("Gland","Dope")) %>%
  mutate(Part = factor(Part, levels = c("Gland","Dope"))) %>%
    select(geneNameProtein, Gland, Part, spider, relativeIntensity, specificity, zone) %>%
  pivot_wider(names_from = c("Gland","spider"), values_from = "relativeIntensity", values_fill = 0.00001) %>%
  pivot_longer(cols = starts_with("Mi") | starts_with("Ma"),
    names_to = c("Sample"),
    values_to =  "relativeIntensity"
  ) %>%
  separate(Sample, into = c( "Gland", "spider"), remove = FALSE)%>%
  mutate(specificity = factor(specificity, levels = c("Major","Common","Minor"))) %>%
  mutate(geneNameProtein = factor(geneNameProtein, levels = rev(GOIs.order$geneNameProtein))) %>%
  ggplot(aes(x = geneNameProtein, y = relativeIntensity * 100,  color = Gland)) +
 geom_jitter(
    aes(color = Gland), # Inherit color from Gland for consistency
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), # Use jitterdodge
    size = 1.5, # Adjust point size
    alpha = 0.7, # Adjust point transparency
  ) +# Rotate the coordinate system by 90 degrees
  coord_flip() +
  facet_grid(specificity~Part, scales = "free_y",space = "free_y") +
  scale_y_continuous(trans = "log10") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    y = "Relative intensity (%) (log10)",
    fill = "Gland"
  )

Proteomics.data %>%
  distinct(Gland,Part,spider) %>%
  group_by(Gland, Part) %>%
  summarise(nrOfSamples = n()) %>%
  inner_join(Proteomics.data) %>%
  inner_join(GOIs) %>%
  ungroup() %>%
  group_by(Gland, Part, specificity, zone, geneNameProtein) %>%
  summarise(relativeIntensity = sum(relativeIntensity)/nrOfSamples) %>%
  distinct()%>%
  pivot_wider(names_from = "Gland", values_from = "relativeIntensity", values_fill = 0.000001) ->
  proteomics.wide.summary



proteomics.wide.summary %>%
  filter(Part %in% c("Gland")) %>%
  ggplot(aes(x = Mi*100, y = Ma*100, fill = specificity,
              label = geneNameProtein)) +
  # Add shaded 2-fold band
  geom_line(
    data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
      mutate(y = 0.1 * x),
    aes(x = x, y = y),
    inherit.aes = FALSE,
    linetype = "dotted",
    color = "grey40"
  ) +
  
  geom_line(
    data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
      mutate(y = 10 * x),
    aes(x = x, y = y),
    inherit.aes = FALSE,
    linetype = "dotted",
    color = "grey40"
  )  +
  geom_ribbon(data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
                     mutate(ymin = 0.5 * x, ymax = 2 * x),
                   aes(x = x, ymin = ymin, ymax = ymax),
                   inherit.aes = FALSE,
                   fill = "grey50", alpha = 0.3) +  

  geom_point(position = "jitter", shape = 22) +  
 geom_label_repel(aes(fill = specificity), color="black",
  label.r = unit(0, "lines"),           # square corners
  label.padding = unit(0.1, "lines"),   # tighter label
  box.padding = unit(0.15, "lines"),    # closer to points
  max.overlaps = 20,
  size = 3,
  label.size = 0.3)+# varwidth adjusts width by sample size
  facet_wrap(~Part, scales = "free") +
  scale_y_continuous(trans = "log10", limits = c(NA, 100)) +
  scale_x_continuous(trans = "log10", limits = c(NA, 10)) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    y = "Relative intensity (%) (log10)",
    fill = "Gland"
  )+ 
  scale_color_manual(values = c(Minor = "#E2D799",Major = "#E2D799",Common ="#CDDCFB"))+
  scale_fill_manual(values = c(Minor = "#E2D799",Major = "#E2D799",Common ="#CDDCFB"))->
  Figure8.next 

Figure8.next





ggsave("~/Google Drive/My Drive/Jobb/Minor gland/paper2/figures_new/figure8/figure8B.pdf",width = 160,height = 160,units = "mm",plot = Figure8.next)
proteomics.wide.summary %>%
  filter(Part %in% c("Gland")) %>%
  ggplot(aes(x = Mi*100, y = Ma*100, color = specificity,
             label = geneNameProtein)) +
# Add shaded 2-fold band
  geom_ribbon(data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
                mutate(ymin = 0.1 * x, ymax = 10 * x),
              aes(x = x, ymin = ymin, ymax = ymax),
              inherit.aes = FALSE,
              fill = "grey80", alpha = 0.3) +  
    geom_ribbon(data = data.frame(x = 10^seq(-4, 1, length.out = 1000)) %>%
                mutate(ymin = 0.5 * x, ymax = 2 * x),
              aes(x = x, ymin = ymin, ymax = ymax),
              inherit.aes = FALSE,
              fill = "grey50", alpha = 0.3) +  

  geom_point(position = "jitter", size = 2) +  
  geom_label_repel(aes(fill = specificity), color="black")+# varwidth adjusts width by sample size
  facet_wrap(~Part, scales = "free") +
  scale_y_continuous(trans = "log10", limits = c(NA, 100)) +
  scale_x_continuous(trans = "log10", limits = c(NA, 10)) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL,
    y = "Relative intensity (%) (log10)",
    fill = "Gland"
  )+ 
  # Manually specify the shapes (21, 22, 23 for our three levels)
  scale_shape_manual(values = c(21, 22, 23))  + 
  scale_color_manual(values = c(Minor = "#E2D799",Major = "#E2D799",Common ="#CDDCFB"))+
  scale_fill_manual(values = c(Minor = "#E2D799",Major = "#E2D799",Common ="#CDDCFB"))




  
proteomics.wide.summary %>%
  filter(specificity== "Common") %>%
  filter(Part == "Gland")->
gland.commmon  

cor.test(gland.commmon$Ma,gland.commmon$Mi)

proteomics.wide.summary %>%
  filter(specificity== "Major") %>%
  filter(Part == "Gland")->
gland.major  

cor.test(gland.major$Ma,gland.major$Mi)

proteomics.wide.summary %>%
  filter(specificity== "Minor") %>%
  filter(Part == "Gland")->
gland.minor  

cor.test(gland.minor$Ma,gland.minor$Mi)



proteomics.wide.summary %>%
  filter(specificity== "Common") %>%
  filter(Part == "Dope")->
gland.commmon  

cor.test(gland.commmon$Ma,gland.commmon$Mi)

proteomics.wide.summary %>%
  filter(specificity== "Major") %>%
  filter(Part == "Dope")->
gland.major  

cor.test(gland.major$Ma,gland.major$Mi)

proteomics.wide.summary %>%
  filter(specificity== "Minor") %>%
  filter(Part == "Dope")->
gland.minor  

cor.test(gland.minor$Ma,gland.minor$Mi)


proteomics.wide.summary %>%
  filter(Part != "Gland" & Part != "Dope") %>%
  filter(Ma > 0.0001) %>%
  ungroup() %>%
  distinct(geneNameProtein) %>%
  mutate(presentInSilk = "Yes") %>%
  right_join(GOIs) %>%
  mutate(presentInSilk= replace(presentInSilk,is.na(presentInSilk),"No"))  %>%
  group_by(specificity, presentInSilk) %>%
  summarise(n = n())
  


```

