---
title: "Bulk RNA analysis"
author: "Johan"
date: "05/26/2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    df_print: paged

params:
  RfunctionsDirectory: /Users/johanreimegard/git/RNAmappingPipeline/R
  
  
---






```{r setup, include=FALSE}

# if (!require("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")#

# BiocManager::install("ropls")

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
#install.packages("kableExtra")
#library("viridis")     

library(RColorBrewer)
#library(GenomicFeatures)

#library(ropls)

library(ggforce)
library(patchwork)


#install.packages("PTXQC")
library(PTXQC)
rename <- dplyr::rename


source("../loadGeneInfo.R")
source("../colorSchemes.R")
summarise = dplyr::summarise               
select = dplyr::select
group_by = dplyr::group_by
rename = dplyr::rename
library(ggrepel)

#install.packages("GGally")
library("GGally")


repositoryPath = "~/git/Rising-Lab/MajorGland_paper"


```


## Functions


## Parameters
```{r, message=FALSE, warning=FALSE}

minExpressionCutoff =20
DataGenerated = TRUE

```
## Reference
```{r}

geneInfo = getGeneInfoFull(repositoryPath = repositoryPath) 
majorProteins = read_csv("~/Box/bridgeSpider/annotation/silk.proteins.1.1.csv") %>%
  select(geneName,geneNameProtein,mean) %>% rename(masp_concentration = "mean")

geneInfo = geneInfo %>% left_join(majorProteins)
transcriptInfo = addTranscriptInfo(geneInfo,repositoryPath = repositoryPath) %>%
  select(gene_id,geneName, tx_name) %>% left_join(majorProteins) 




geneInfo %>% filter(!is.na(geneNameProtein))

signalPeptideInfo = getSignalPeptideInfo(repositoryPath = repositoryPath,geneInfo.transcript =  transcriptInfo)%>%
  add_row(Term = "signalP", geneName = "MiSpc")

geneInfo %>% left_join(signalPeptideInfo) ->geneInfo


majorProteins %>% select(-masp_concentration) %>%
  slice_head(n = 17)  %>% filter(grepl("SpiCE", geneNameProtein)) %>% mutate(class = "SpiCE-LMa") -> spiceProteins 
majorProteins %>%select(-masp_concentration) %>%
  slice_head(n = 17)  %>% filter(grepl("AmSp-like", geneNameProtein)) %>% mutate(class = "AmSp-like") -> AmSpProteins
geneInfo %>% select(geneName) %>% filter(grepl("MaSp", geneName)) %>% mutate(geneNameProtein = geneName) %>% mutate(class = "MaSp") -> MaSpProteins
geneInfo %>% select(geneName) %>% filter(grepl("MiSp", geneName)) %>% mutate(geneNameProtein = geneName) %>% mutate(class = "MiSp") -> MiSpProteins
geneInfo %>% select(geneName) %>% filter(grepl("FlSp", geneName)) %>% mutate(geneNameProtein = geneName) %>% mutate(class = "FlSp") -> FlSpProteins



geneInfo %>% filter(grepl("MiSp", geneName)) 
silkGenes = bind_rows(spiceProteins, AmSpProteins,MaSpProteins,MiSpProteins,FlSpProteins)


```



## Proteomics


### Read in the data
#### Major

```{r}

MQ_data.major = read_tsv("../../data/proteomics/majorSilk/txt/proteinGroups.txt")


colnames(MQ_data.major)

MQ_data.major %>% select(1,2,120:133) %>%
  rename(Ma_Dope_2 = "LFQ intensity MA_dope_2") %>%
  rename(Ma_Dope_3 = "LFQ intensity MA_dope_3") %>%
  rename(Ma_Gland_1 = "LFQ intensity MA_gland_1") %>%
  rename(Ma_Gland_2 = "LFQ intensity MA_gland_2") %>%
  rename(Ma_Gland_3 = "LFQ intensity MA_gland_3") %>%
  rename(Ma_SilkHFIP_1 = "LFQ intensity Native_Silk_HFIP_1") %>%
  rename(Ma_SilkHFIP_2 = "LFQ intensity Native_Silk_HFIP_2") %>%
  rename(Ma_SilkHFIP_3 = "LFQ intensity Native_Silk_HFIP_3") %>%
  rename(Ma_SilkLiBr_1 = "LFQ intensity Native_Silk_LiBr_1") %>%
  rename(Ma_SilkLiBr_2 = "LFQ intensity Native_Silk_LiBr_2") %>%
  rename(Ma_SilkLiBr_3 = "LFQ intensity Native_Silk_LiBr_3") %>%
  rename(Ma_Silk_1 = "LFQ intensity Native_Silk_Urea_1") %>%
  rename(Ma_Silk_2 = "LFQ intensity Native_Silk_Urea_2") %>%
  rename(Ma_Silk_3 = "LFQ intensity Native_Silk_Urea_3") %>%
  rename(geneName = "Protein IDs") %>%
  select(geneName,Ma_Dope_2,Ma_Dope_3,Ma_Gland_1,Ma_Gland_2,Ma_Gland_3,Ma_Silk_1,Ma_Silk_2,Ma_Silk_3)->
  MQ_data.major.intensity



colnames(MQ_data.major.intensity)
```

#### Minor



```{r}

MQ_data.minor = read_tsv("../../data/proteomics/minorSilk/txt/proteinGroups.txt")


colnames(MQ_data.minor)

MQ_data.minor %>% select(1,2,106:117) -> MQ_data.minor.intensity 


MQ_data.minor.intensity %>%
  rename(Mi_Dope_1 = "LFQ intensity Dope_spider_1") %>%
  rename(Mi_Dope_2 = "LFQ intensity Dope_spider_2") %>%
  rename(Mi_Dope_3 = "LFQ intensity Dope_spider_3") %>%
  rename(Mi_Dope_4 = "LFQ intensity Dope_spider_4") %>%
  rename(Mi_Gland_1 = "LFQ intensity Gland_spider_1") %>%
  rename(Mi_Gland_2 = "LFQ intensity Gland_spider_2") %>%
  rename(Mi_Gland_3 = "LFQ intensity Gland_spider_3") %>%
  rename(Mi_Gland_4 = "LFQ intensity Gland_spider_4") %>%
  rename(Mi_Gland_5 = "LFQ intensity Intact_Mi_Gland_spider_5") %>%
  rename(Mi_Gland_6 = "LFQ intensity Intact_Mi_Gland_spider_6") %>%
  rename(Mi_Silk_1 = "LFQ intensity Silk_Mi_1") %>%
  rename(Mi_Silk_2 = "LFQ intensity Silk_Mi_2") %>%
  rename(geneName = "Protein IDs") ->
  MQ_data.minor.intensity
colnames(MQ_data.minor.intensity)


MQ_data.minor.intensity %>% filter(geneName %in% c("LASC001730","LASC00038","smoc1","LASC010012"))
```


#### Flageliform

```{r}

MQ_data = read_tsv("../../data/proteomics/flageliform/proteinGroups.txt")


MQ_data %>% select(1,2,50:53) -> MQ_data.intensity.Fl 


MQ_data.intensity.Fl %>%
  rename(Fl_Gland_2 = "LFQ intensity Flagelli_Gland_spider_2") %>%
  rename(Fl_Gland_3 = "LFQ intensity Flagelli_Gland_spider_3") %>%
  rename(Fl_Gland_4 = "LFQ intensity Flagelli_Gland_spider_4") %>%
  rename(Fl_Gland_5 = "LFQ intensity Intact_F_Gland_spider_5") %>%
  rename(geneName = "Protein IDs") ->
  MQ_data.intensity.Fl
colnames(MQ_data.intensity.Fl)

write_csv(MQ_data.intensity.Fl,"../../data/proteomics/flageliform_LFQ_int.csv.gz")


```
#### Aggregate

```{r}

MQ_data = read_tsv("../../data/proteomics/aggregateGland/txt/proteinGroups.txt")

#colnames(MQ_data)
MQ_data %>% select(1,2,106:117) -> MQ_data.intensity.Ag  
gsub("LFQ intensity ","", colnames(MQ_data.intensity.Ag)) -> 
  newCN
gsub("TF1_","",newCN)->
  newCN
colnames(MQ_data.intensity.Ag) = newCN


MQ_data.intensity.Ag %>%
  rename(geneName = "Protein IDs") ->
  MQ_data.intensity.Ag
colnames(MQ_data.intensity.Ag)
```

### QC 

Not using this right now since it is not working properly with the data. Something that I do not understand. But will keep it if I need to do the analysis later. 

```{r, include=FALSE}


#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("DEP")
#BiocManager::install("limma")


#library(DEP)


# Preprocessing: filtering, normalization, imputation
#data_clean <- filter_proteins(data)
#data_normalized <- normalize_vsn(data_clean)
#data_imputed <- impute_missing(data_normalized)

# Perform DE analysis
#contrast_matrix <- make_contrast_matrix(experimental_design)
#results <- test_diff(data_imputed, contrast = contrast_matrix)

# Visualize results
#plot_volcano(results)
#plot_heatmap(results)


```




#### Major
```{r}

MQ_data.major.intensity %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE)->
  MQ_data.MaSp


MQ_data.MaSp%>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Part")



MQ_data.MaSp %>% mutate(Type = "Other") %>%
  mutate(Type = replace(Type, grepl("MaSp1",geneName), "MaSp1")) %>%
  mutate(Type = replace(Type, grepl("MaSp2",geneName), "MaSp2")) %>%
  mutate(Type = replace(Type, grepl("MaSp3",geneName), "MaSp3")) %>%
  mutate(Type = replace(Type, grepl("MaSp4",geneName), "MaSp4")) %>%
  mutate(Type = replace(Type, grepl("CON_",geneName), "Contaminants")) %>%
  mutate(Type = replace(Type, grepl("REV_",geneName), "FALSE"))->
  MQ_data.MaSp

MQ_data.MaSp %>% filter(intensity > 10000) %>%
  filter(Type != "Contaminants") %>%
  filter(Type != "FALSE") %>%
  group_by(Gland,spider,Part,Type) %>%
  summarise(n = n()) %>%
  pivot_wider(values_from = "n", names_from = "Type")





MQ_data.MaSp %>%
  filter(intensity > 1000000) %>%
  filter(Type != "Contaminants") %>%
  filter(Type != "FALSE")->
  MQ_data.MaSp.QC




MQ_data.MaSp.QC %>% group_by(Gland,spider,Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.MaSp.QC) %>%
  mutate(relativeIntensity = intensity / TotalIntensity) ->
  MQ_data.MaSp.QC

MQ_data.MaSp.QC %>% 
  ggplot(aes(x = Part, y = relativeIntensity, fill = Type)) + geom_col() +facet_wrap("spider")



MQ_data.MaSp.QC %>% ungroup() %>%
  select(Sample, geneName,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity") %>% column_to_rownames("geneName")   ->MQ_data.MaSp.QC.wide




samples = MQ_data.MaSp.QC %>% distinct(Sample) %>% arrange(Sample)
ggpairs(log10(MQ_data.MaSp.QC.wide[,samples$Sample]))


MQ_data.MaSp.QC %>% filter(Sample != "Ma_Gland2")->MQ_data.MaSp.QC



MQ_data.MaSp.QC %>% group_by(geneName,Part) %>%
  summarise(relativeIntensity = mean(relativeIntensity)) %>%
  filter(relativeIntensity > 0.0001) %>% mutate(gland = "Major") -> 
MQ_data.MaSp.QC.summary


```








#### Minor 

```{r}

MQ_data.minor.intensity %>% select(1,3,4,5,6,7,8,9,10,11,12,13,14) %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE)->
  MQ_data.MiSp


MQ_data.MiSp%>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Part")



MQ_data.MiSp %>% mutate(Type = "Other") %>%
  mutate(Type = replace(Type, grepl("MiSp",geneName), "MiSp")) %>%
  mutate(Type = replace(Type, grepl("CON_",geneName), "Contaminants")) %>%
  mutate(Type = replace(Type, grepl("REV_",geneName), "FALSE"))->
  MQ_data.MiSp

MQ_data.MiSp %>%
  filter(intensity > 10000) %>%
  filter(Type %in% c("Other", "MiSp")) %>%
  filter(Sample != "Mi_Silk_1") ->
  MQ_data.MiSp.QC



MQ_data.MiSp.QC %>% group_by(Gland,spider,Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.MiSp.QC) %>%
  mutate(relativeIntensity = intensity / TotalIntensity) ->
  MQ_data.MiSp.QC

MQ_data.MiSp.QC  %>% ggplot(aes(x = Part, y = relativeIntensity, fill = Type)) + geom_col() +facet_wrap("spider")

MQ_data.MiSp.QC %>% filter(spider == 2) %>% 
  select(Sample, geneName,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity", values_fill = 0.00001) %>%
  filter(Mi_Gland_2>0.00001 | Mi_Silk_2>0.00001) %>%
  ggplot(aes(x = Mi_Gland_2, y = Mi_Silk_2, label = geneName)) + geom_point()+ geom_text_repel()+ ggtitle("Comparing Dope and silk common ")+
  scale_x_log10() +  # Apply log scale to x-axis
  scale_y_log10()


MQ_data.MiSp.QC %>%
  mutate(Type2 = Type) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpa","MiSpa")) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpb","MiSpb")) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpc","MiSpc")) %>%
  mutate(Type2 = replace(Type2,geneName == "MiSpd","MiSpd"))->
  MQ_data.MiSp.QC



MQ_data.MiSp.QC %>% ggplot(aes(x = Part, y = relativeIntensity, fill = Type2)) + geom_col() +facet_wrap("spider")



MQ_data.MiSp.QC %>% ungroup() %>%
  select(Sample, geneName,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity") %>% column_to_rownames("geneName")   -> MQ_data.MiSp.QC.wide

MQ_data.MiSp.QC %>% group_by(geneName,Part) %>%
  summarise(relativeIntensity = mean(relativeIntensity)) %>%
  filter(relativeIntensity > 0.0001) %>% mutate(gland = "Minor") -> 
MQ_data.MiSp.QC.summary



samples = MQ_data.MiSp.QC %>% distinct(Sample) %>% arrange(Sample)
ggpairs(log10(MQ_data.MiSp.QC.wide[,samples$Sample]))

bind_rows(MQ_data.MiSp.QC.summary,MQ_data.MaSp.QC.summary) ->
  MQ_data.AmpGl.QC.summary

write_csv(MQ_data.AmpGl.QC.summary,"../../data/proteomics/MQ_data.AmpGl.QC.summary.csv")

```


#### Flageliform

```{r}



MQ_data.intensity.Fl %>% select(1,3,4,5,6) %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE)->
  MQ_data.FlSp


MQ_data.FlSp%>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Part")



MQ_data.FlSp %>% mutate(Type = "Other") %>%
  mutate(Type = replace(Type, grepl("FlSp",geneName), "FlSp")) %>%
  mutate(Type = replace(Type, grepl("FlSp-like",geneName), "FlSp-like")) %>%
  mutate(Type = replace(Type, grepl("AcSp",geneName), "AcSp")) %>%
  
  mutate(Type = replace(Type, grepl("CON_",geneName), "Contaminants")) %>%
  mutate(Type = replace(Type, grepl("REV_",geneName), "FALSE"))->
  MQ_data.FlSp



MQ_data.FlSp %>%
  filter(intensity > 10000) %>%
  filter(Type %in% c("Other", "FlSp","AcSp", "FlSp-like"))->
  MQ_data.FlSp.QC



MQ_data.FlSp.QC %>% group_by(Gland,spider,Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.FlSp.QC) %>%
  mutate(relativeIntensity = intensity / TotalIntensity) ->
  MQ_data.FlSp.QC

MQ_data.FlSp.QC  %>% ggplot(aes(x = Part, y = relativeIntensity, fill = Type)) + geom_col() +facet_wrap("spider")






MQ_data.FlSp.QC %>% ungroup() %>%
  select(Sample, geneName,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity") %>% column_to_rownames("geneName")   -> MQ_data.FlSp.QC.wide




samples = MQ_data.FlSp.QC %>% distinct(Sample) %>% arrange(Sample)
ggpairs(log10(MQ_data.FlSp.QC.wide[,samples$Sample]))



MQ_data.FlSp.QC %>% arrange(-relativeIntensity)


MQ_data.FlSp.QC %>% filter(Sample != "Fl_Gland_5")->
  MQ_data.FlSp.QC
```
#### Aggregate

```{r}



MQ_data.intensity.Ag %>% select(-2) %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("ID","spider","Side","Part"),sep = "_", remove = FALSE)->
  MQ_data.Ag


MQ_data.Ag%>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Part")



MQ_data.Ag %>% mutate(Type = "Other") %>%
  mutate(Type = replace(Type, grepl("AgSp",geneName), "AgSp")) %>%
  mutate(Type = replace(Type, grepl("Spidroin2",geneName), "AgSp-like")) %>%
  mutate(Type = replace(Type, grepl("AcSp",geneName), "AcSp")) %>%
  
  mutate(Type = replace(Type, grepl("CON_",geneName), "Contaminants")) %>%
  mutate(Type = replace(Type, grepl("REV_",geneName), "FALSE"))->
  MQ_data.Ag

MQ_data.Ag%>% filter(Type %in% c("Other","Contaminants","FALSE")) %>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Type")





MQ_data.Ag %>%
  filter(intensity > 10000) %>%
  filter(Type %in% c("Other", "AgSp","AcSp", "AgSp-like"))->
  MQ_data.Ag.QC



MQ_data.Ag.QC %>% group_by(Side,spider,Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.Ag.QC) %>%
  mutate(relativeIntensity = intensity / TotalIntensity) ->
  MQ_data.Ag.QC

MQ_data.Ag.QC  %>% ggplot(aes(x = Part, y = relativeIntensity, fill = Type)) + geom_col() +facet_wrap("spider")

MQ_data.Ag%>% filter(Type %in% c("Other","Contaminants","FALSE")) %>%
  ggplot(aes(x= log10(intensity+1), color = spider)) + geom_density()+ facet_wrap("Type")


MQ_data.Ag.QC %>%group_by(geneName,Part) %>%
  summarise(mean = mean(relativeIntensity)) %>%
  pivot_wider(values_from = "mean",names_from = "Part" ) %>%
  arrange(-P)



MQ_data.Ag.QC %>% ungroup() %>%
  select(Sample, geneName,relativeIntensity) %>%
  pivot_wider(names_from = "Sample", values_from = "relativeIntensity") %>% column_to_rownames("geneName")   -> MQ_data.Ag.QC.wide




ggpairs(log10(MQ_data.Ag.QC.wide))


```






# Comparison


## Using Normalised intensities
```{r}

Proteomics.data = bind_rows(MQ_data.FlSp.QC, MQ_data.MaSp.QC,MQ_data.MiSp.QC)
Proteomics.data %>% ggplot(aes(x = log10(intensity+1), color = Part)) + geom_density()+facet_wrap("Gland")


Proteomics.data  %>%
  mutate(logInt = log10(intensity+1))%>%
  filter(logInt >3) %>%
  group_by(geneName,Gland, Part ) %>%
  summarise(n= n(), intensity = sum(logInt)/n) %>%
  separate_longer_delim(geneName, delim = ";")  -> 
  Proteomics.data.summary

Proteomics.data.summary%>% ggplot(aes(x = intensity, color = Part)) + geom_density()+facet_wrap("Gland")



Proteomics.data.summary %>% left_join(silkGenes)-> Proteomics.data.summary

Proteomics.data.summary %>% group_by(Gland, Part) %>%
  summarise(mean = mean(intensity), sd = sd(intensity)) %>%
  ungroup() %>%
  inner_join(Proteomics.data.summary) %>%
  mutate(int_norm = (intensity - mean)/sd )->
  Proteomics.data.summary

Proteomics.data.summary%>% ggplot(aes(x = int_norm, color = Gland)) + geom_density()+facet_wrap("Part")


Proteomics.data.summary %>% filter(!is.na(geneNameProtein)) -> Proteomics.data.silk.summary
Proteomics.data.summary %>% filter(is.na(geneNameProtein)) -> Proteomics.data.other.summary


Proteomics.data %>%
  ungroup() %>%
  select(geneName,Sample,intensity) %>%
  pivot_wider(names_from = "Sample",values_from = "intensity", values_fill = 1000)  %>%
  pivot_longer(names_to = "Sample", values_to ="intensity" , -geneName ) -> Proteomics.data.intensity

Proteomics.data.intensity %>% distinct()

Proteomics.data %>%ungroup() %>%
  distinct(Gland,spider,Part,Sample ) %>%
  inner_join(Proteomics.data.intensity) ->
  Proteomics.data.full

Proteomics.data%>% ungroup() %>%
  distinct(geneName,Type,Type2) %>%
  arrange(geneName,Type2) %>%
  group_by(geneName ) %>%
  slice_head(n =1)%>%
  inner_join(Proteomics.data.full) ->
  Proteomics.data.full

Proteomics.data%>%
  ungroup() %>%
  distinct(geneName,Type,Type2)
 





```

#### Gland

```{r}

Proteomics.data.silk.summary %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneNameProtein,class, Gland,int_norm) %>%
  pivot_wider(names_from = "Gland",values_from = "int_norm", values_fill = -2) ->
  Proteomics.data.silk.gland.summary

Proteomics.data.other.summary %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneName, Gland,class,int_norm) %>%
  pivot_wider(names_from = "Gland",values_from = "int_norm", values_fill = -2) ->
  Proteomics.data.other.gland.summary


Proteomics.data.other.gland.summary %>% 
  inner_join(signalPeptideInfo)%>%
  mutate(Term = replace(Term,is.na(Term),"NoSignalPeptide"))-> 
  Proteomics.data.other.gland.sp.summary

ggplot(data =Proteomics.data.other.gland.sp.summary,  aes(x = Mi, y = Ma))+ geom_point( aes(size = Term))+
  geom_point(data = Proteomics.data.silk.gland.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.gland.summary,aes(label = geneNameProtein, color = class ),force = 15)+
  geom_abline(slope = 1,intercept = c(0,0))+
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) -> glands.plot.ma.mi

ggplot(data =Proteomics.data.other.gland.sp.summary,  aes(x = Ma, y = Fl))+ geom_point( aes(size = Term))+
  geom_point(data = Proteomics.data.silk.gland.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.gland.summary,aes(label = geneNameProtein, color = class ),force = 15)+
  geom_abline(slope = 1,intercept = c(0,0))+
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) -> glands.plot.ma.fl

ggplot(data =Proteomics.data.other.gland.sp.summary,  aes(x = Mi, y = Fl))+ geom_point( aes(size = Term))+
  geom_point(data = Proteomics.data.silk.gland.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.gland.summary,aes(label = geneNameProtein, color = class ),force = 15)+
  geom_abline(slope = 1,intercept = c(0,0))+
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) -> glands.plot.mi.fl

glands.plot.ma.mi
glands.plot.ma.fl
glands.plot.mi.fl

Proteomics.data.other.gland.sp.summary %>% arrange( -Mi)  

```


```{r  statistics}
Proteomics.data.summary %>%
  filter(Part == "Gland")  %>%
  filter(Gland %in% c("Ma","Mi")) %>%
  filter(n >= 2) %>%
  inner_join(signalPeptideInfo)  %>%
  group_by(geneName) %>%
  summarise(nrOfGlands = n()) %>%
  filter(nrOfGlands ==2) %>%
  inner_join(Proteomics.data) %>%
  filter(Part == "Gland")  %>%
   filter(Gland %in% c("Ma","Mi"))   -> 
  Proteomics.data.ma.mi.common


Proteomics.data.ma.mi.common %>%
  mutate(min = min())







mann_whitney_test <- function(df) {
  wilcox.test(relativeIntensity ~ Gland, data = df)
}

# Apply the test for each gene and return results
results.ttest <- Proteomics.data.ma.mi.common %>%
  group_by(geneName) %>%
  summarize(
    t_p_value = t.test(relativeIntensity ~ Gland, var.equal = FALSE)$p.value,
    t_statistic = t.test(relativeIntensity ~ Gland, var.equal = FALSE)$statistic
  )
# Show the results
results.ttest %>% ggplot(aes(t_p_value)) + geom_density()


# Apply the test for each gene and return results
results.wilcox <- Proteomics.data.ma.mi.common %>%
  group_by(geneName) %>%
  summarize(
    w_p_value = wilcox.test(relativeIntensity ~ Gland, var.equal = FALSE)$p.value,
    w_statistic = wilcox.test(relativeIntensity ~ Gland, var.equal = FALSE)$statistic
  )
# Show the results
results.wilcox %>% ggplot(aes(w_p_value)) + geom_density()

inner_join(results.wilcox,results.ttest) %>%
  ggplot(aes(x =w_p_value,y = t_p_value ))+ geom_point() 


inner_join(results.wilcox,results.ttest) %>%
  filter(w_p_value < 0.1 &t_p_value < 0.1 )
```


#### Dope
```{r}
Proteomics.data.silk.summary %>%
  filter(Part == "Dope") %>%
  ungroup() %>%
  select(geneNameProtein,class, Gland,int_norm) %>%
  pivot_wider(names_from = "Gland",values_from = "int_norm", values_fill = -3) ->
  Proteomics.data.silk.dope.summary

Proteomics.data.other.summary %>%
  filter(Part == "Dope") %>%
  ungroup() %>%
  select(geneName, Gland,int_norm) %>%
  pivot_wider(names_from = "Gland",values_from = "int_norm", values_fill = -3) ->
  Proteomics.data.other.dope.summary


Proteomics.data.other.dope.summary%>% 
  inner_join(signalPeptideInfo)%>%
  mutate(Term = replace(Term,is.na(Term),"NoSignalPeptide"))-> 
  Proteomics.data.other.dope.summary

ggplot(data =Proteomics.data.other.dope.summary,  aes(x = Mi, y = Ma))+ geom_point( aes(size = Term))+
  geom_point(data = Proteomics.data.silk.dope.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.dope.summary,
                  aes(label = geneNameProtein, color = class )) +
  geom_abline(slope = 1,intercept = c(0,0)) +
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) ->
  dope.plot.ma.mi




dope.plot.ma.mi
```

#### Major gland
```{r}

Proteomics.data.silk.summary %>%
  filter(Gland == "Ma") %>%
  ungroup() %>%
  select(geneNameProtein,class, Part,intensity) %>%
  pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) ->
  Proteomics.data.silk.Ma.summary

Proteomics.data.other.summary %>%
  filter(Gland == "Ma") %>%
  ungroup() %>%
  select(geneName, Part,intensity) %>%
  pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) %>%
  filter(Dope > 0) %>% filter(Gland > 0) ->
  Proteomics.data.other.Ma.summary

ggplot(data =Proteomics.data.other.Ma.summary,  aes(x = Dope, y = Gland))+ geom_point( size = 0.1)+ geom_smooth(method = "lm")+
  geom_point(data = Proteomics.data.silk.Ma.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.Ma.summary,aes(label = geneNameProtein, color = class )) ->Ma.plot
```

#### Minor gland
```{r}

Proteomics.data.silk.summary %>%
  filter(Gland == "Mi") %>%
  ungroup() %>%
  select(geneNameProtein,class, Part,intensity) %>%
  pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) ->
  Proteomics.data.silk.Mi.summary

Proteomics.data.other.summary %>%
  filter(Gland == "Mi") %>%
  ungroup() %>%
  select(geneName, Part,intensity) %>%
  pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) %>%
  filter(Dope > 0) %>% filter(Gland > 0) ->
  Proteomics.data.other.Mi.summary

ggplot(data =Proteomics.data.other.Mi.summary,  aes(x = Gland, y = Dope))+ geom_point( size = 0.1)+ 
  geom_point(data = Proteomics.data.silk.Mi.summary, aes(color = class))+
  geom_abline(slope = 1,intercept = 0)+
  geom_text_repel(data = Proteomics.data.silk.Mi.summary,aes(label = geneNameProtein, color = class,
                                                             max.overlaps = Inf ,force = 0.1)) ->Mi.plot.GD


ggplot(data =Proteomics.data.other.Mi.summary,  aes(x = Gland, y = Silk))+ geom_point( size = 0.1)+ 
  geom_point(data = Proteomics.data.silk.Mi.summary, aes(color = class))+
  geom_abline(slope = 1,intercept = 0)+
  geom_text_repel(data = Proteomics.data.silk.Mi.summary,aes(label = geneNameProtein, color = class,
                                                             max.overlaps = Inf ,force = 0.1)) ->Mi.plot.GS


ggplot(data =Proteomics.data.other.Mi.summary,  aes(x = Dope, y = Silk))+ geom_point( size = 0.1)+ 
  geom_point(data = Proteomics.data.silk.Mi.summary, aes(color = class))+
  geom_abline(slope = 1,intercept = 0)+
  geom_text_repel(data = Proteomics.data.silk.Mi.summary,aes(label = geneNameProtein, color = class,
                                                             max.overlaps = Inf ,force = 0.1)) ->Mi.plot.DS


ggplot(data =Proteomics.data.other.Mi.summary,  aes(x = Gland, y = Silk))+ geom_point( size = 0.1)+ 
  geom_point(data = Proteomics.data.silk.Mi.summary, aes(color = class))+
  geom_abline(slope = 1,intercept = 0)+
  geom_text_repel(data = Proteomics.data.silk.Mi.summary,aes(label = geneNameProtein, color = class,
                                                             max.overlaps = Inf ,force = 0.1)) ->Mi.plot.GS

Mi.plot.GD
Mi.plot.GS+Mi.plot.DS+ plot_layout(guides = "collect")

Proteomics.data.other.Mi.summary %>% filter(Silk > 0)

```

## Using relative intensity
```{r}

Proteomics.data = bind_rows(MQ_data.FlSp, MQ_data.MaSp,MQ_data.MiSp)
Proteomics.data %>% ggplot(aes(x = log10(intensity+1), color = Part)) + geom_density()+facet_wrap("Gland")


Proteomics.data  %>%
  filter(intensity > 1000) %>%
  group_by(geneName,Gland, Part ) %>%
  summarise(n= n(), intensity = mean(intensity))->
  Proteomics.data.mean

Proteomics.data.mean %>%
  group_by(Gland, Part ) %>%
  summarise(Total = sum(intensity)) %>%
  inner_join(Proteomics.data.mean) %>%
  mutate(relativeIntensity = intensity/Total) ->
  Proteomics.data.relative

Proteomics.data.relative%>% ggplot(aes(x = log10(relativeIntensity), color = Part)) + geom_density()+facet_wrap("Gland")



Proteomics.data.relative %>% left_join(silkGenes)-> Proteomics.data.relative
Proteomics.data.relative %>% filter(!is.na(geneNameProtein)) -> Proteomics.silk.relative
Proteomics.data.relative %>% filter(is.na(geneNameProtein)) -> Proteomics.other.relative


```


#### Gland

```{r}
min = min(Proteomics.data.relative$relativeIntensity)
Proteomics.silk.relative %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneNameProtein,class, Gland,relativeIntensity) %>%
  pivot_wider(names_from = "Gland",values_from = "relativeIntensity", values_fill = (min)) ->
  Proteomics.silk.gland.relative

Proteomics.other.relative %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneName, Gland,relativeIntensity) %>%
  pivot_wider(names_from = "Gland",values_from = "relativeIntensity", values_fill = (min)) ->
  Proteomics.other.gland.relative

ggplot(data =Proteomics.other.gland.relative,  aes(x = Mi, y = Ma))+ geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.gland.relative, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.gland.relative,aes(label = geneNameProtein, color = class )) -> glands.plot

library(scales)

glands.plot+
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent )+
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent)+
  theme_light()+ ggtitle("Gland Major vs Minor")+ geom_abline(intercept = 0, slope = 1) -> Gland.plot


```

#### Dope
```{r}

Proteomics.silk.relative
Proteomics.silk.relative %>%
  filter(Part == "Dope") %>%
  ungroup() %>%
  select(geneNameProtein,class, Gland,relativeIntensity) %>%
  pivot_wider(names_from = "Gland",values_from = "relativeIntensity", values_fill = min) ->
  Proteomics.silk.relative.dope
Proteomics.other.relative %>%
  filter(Part == "Dope") %>%
  ungroup() %>%
  select(geneName, Gland,relativeIntensity) %>%
  pivot_wider(names_from = "Gland",values_from = "relativeIntensity", values_fill = min) ->
  Proteomics.other.relative.dope

ggplot(data =Proteomics.other.relative.dope,  aes(x = Mi, y = Ma))+ geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.relative.dope, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.relative.dope,aes(label = geneNameProtein, color = class )) ->dope.plot

dope.plot+
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent )+
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent)+
  theme_light()+ ggtitle("Dope Major vs Minor")+ geom_abline(intercept = 0, slope = 1) -> Dope.plot

Dope.plot

```

#### Major gland
```{r}

Proteomics.silk.relative %>%
  filter(Gland == "Ma") %>%
  ungroup() %>%
  select(geneNameProtein,class, Part,relativeIntensity) %>%
  pivot_wider(names_from = "Part",values_from = "relativeIntensity", values_fill = min) ->
  Proteomics.silk.relative.dope

Proteomics.other.relative %>%
  filter(Gland == "Ma") %>%
  ungroup() %>%
  select(geneName, Part,relativeIntensity) %>%
  pivot_wider(names_from = "Part",values_from = "relativeIntensity", values_fill = min) ->
  Proteomics.other.relative.dope

ggplot(data =Proteomics.other.relative.dope,  aes(x = Gland, y = Dope))+ geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.relative.dope, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.relative.dope,aes(label = geneNameProtein, color = class )) ->Major.plot

Major.plot+
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent )+
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent)+
  theme_light()+ ggtitle("Major gland Gland vs Dope")+ geom_abline(intercept = 0, slope = 1) -> Major.plot

```

#### Minor gland
```{r}

Proteomics.silk.relative %>%
  filter(Gland == "Mi") %>%
  ungroup() %>%
  select(geneNameProtein,class, Part,relativeIntensity) %>%
  pivot_wider(names_from = "Part",values_from = "relativeIntensity", values_fill = min) ->
  Proteomics.silk.relative.dope

Proteomics.other.relative %>%
  filter(Gland == "Ma") %>%
  ungroup() %>%
  select(geneName, Part,relativeIntensity) %>%
  pivot_wider(names_from = "Part",values_from = "relativeIntensity", values_fill = min) ->
  Proteomics.other.relative.dope

ggplot(data =Proteomics.other.relative.dope,  aes(x = Gland, y = Dope))+ geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.relative.dope, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.relative.dope,aes(label = geneNameProtein, color = class )) ->dope.plot

dope.plot+
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent )+
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1),labels = percent) +
  theme_light()+ ggtitle("Minor gland Gland vs Dope") + geom_abline(intercept = 0, slope = 1) -> Minor.plot


(Gland.plot+Dope.plot) / (Minor.plot+ Major.plot) + patchwork::plot_layout(guides = "collect")



Proteomics.data.silk.summary %>%
filter(Part == "Dope") %>%
ungroup() %>%
select(geneNameProtein,class, Gland,intensity) %>%
pivot_wider(names_from = "Gland",values_from = "intensity", values_fill = -12) ->
Proteomics.data.silk.dope.summary

Proteomics.data.other.summary %>%
filter(Part == "Dope") %>%
ungroup() %>%
select(geneName, Gland,intensity) %>%
pivot_wider(names_from = "Gland",values_from = "intensity", values_fill = 0) ->
Proteomics.data.other.dope.summary

ggplot(data =Proteomics.data.other.dope.summary,  aes(x = Mi, y = Ma))+ geom_point( size = 0.1)+
geom_point(data = Proteomics.data.silk.dope.summary, aes(color = class))+
geom_text_repel(data = Proteomics.data.silk.dope.summary,aes(label = geneNameProtein, color = class )) ->dope.plot

Proteomics.data.silk.summary %>%
filter(Gland == "Ma") %>%
ungroup() %>%
select(geneNameProtein,class, Part,intensity) %>%
pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) ->
Proteomics.data.silk.Ma.summary

Proteomics.data.other.summary %>%
filter(Gland == "Ma") %>%
ungroup() %>%
select(geneName, Part,intensity) %>%
pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) %>%
filter(Dope > 0) %>% filter(Gland > 0) ->
Proteomics.data.other.Ma.summary

ggplot(data =Proteomics.data.other.Ma.summary,  aes(x = Dope, y = Gland))+ geom_point( size = 0.1)+ geom_smooth(method = "lm")+
geom_point(data = Proteomics.data.silk.Ma.summary, aes(color = class))+
geom_text_repel(data = Proteomics.data.silk.Ma.summary,aes(label = geneNameProtein, color = class )) ->Ma.plot



Proteomics.data.silk.summary %>%
filter(Gland == "Mi") %>%
ungroup() %>%
select(geneNameProtein,class, Part,intensity) %>%
pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) ->
Proteomics.data.silk.Mi.summary

Proteomics.data.other.summary %>%
filter(Gland == "Mi") %>%
ungroup() %>%
select(geneName, Part,intensity) %>%
pivot_wider(names_from = "Part",values_from = "intensity", values_fill = 0) %>%
filter(Dope > 0) %>% filter(Gland > 0) ->
Proteomics.data.other.Mi.summary

ggplot(data =Proteomics.data.other.Mi.summary,  aes(x = Dope, y = Gland))+ geom_point( size = 0.1)+ geom_smooth(method = "lm")+
geom_point(data = Proteomics.data.silk.Mi.summary, aes(color = class))+
geom_text_repel(data = Proteomics.data.silk.Mi.summary,aes(label = geneNameProtein, color = class )) ->Mi.plot





```

