---
title: "Proteomics Analysis of Spider Glands"
subtitle: "Analysis of major, minor, flagelliform, and aggregate gland proteomics"
author: "Johan"
date: "05/26/2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
    df_print: paged

params:
  RfunctionsDirectory: /Users/johanreimegard/git/RNAmappingPipeline/R
  
  
---

## 💻 Setup and Library Loading
This section sets up the R environment by loading necessary libraries for data manipulation, visualization, and proteomics analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load core tidyverse libraries for data manipulation and visualization
library(tidyverse)
library(knitr)
library(RColorBrewer)
library(ggforce)
library(patchwork)
library(ggrepel)
library(GGally)
library(PTXQC) # For proteomics quality control

# Overriding functions to ensure consistency with tidyverse
# This prevents conflicts with other packages
rename <- dplyr::rename
summarise <- dplyr::summarise
select <- dplyr::select
group_by <- dplyr::group_by
rename <- dplyr::rename

# Source custom R scripts for gene information and color schemes
source("../loadGeneInfo.R")
source("../colorSchemes.R")

# Define the repository path for data and scripts
repositoryPath <- "~/git/Rising-Lab/MajorGland_paper"
```

## ⚙️ Analysis Parameters and Reference Data

This section defines global analysis parameters and loads critical reference information, including gene annotations and known silk protein data, which will be used throughout the script.

```{r parameters-and-reference}
# Define analysis parameters
minExpressionCutoff <- 20
DataGenerated <- TRUE

# Load gene and transcript information
geneInfo <- getGeneInfoFull(repositoryPath = repositoryPath) 
majorProteins <- read_csv("~/Box/bridgeSpider/annotation/silk.proteins.1.1.csv") %>%
  select(geneName, geneNameProtein, mean) %>% 
  rename(masp_concentration = "mean")

# Join gene information with protein data and signal peptide information
geneInfo <- geneInfo %>% left_join(majorProteins)
transcriptInfo <- addTranscriptInfo(geneInfo, repositoryPath = repositoryPath) %>%
  select(gene_id, geneName, tx_name) %>% left_join(majorProteins)

# Load signal peptide information and add a specific entry for 'MiSpc'
signalPeptideInfo <- getSignalPeptideInfo(repositoryPath = repositoryPath, geneInfo.transcript = transcriptInfo) %>%
  add_row(Term = "signalP", geneName = "MiSpc")

# Add signal peptide info to the main geneInfo dataframe
geneInfo <- geneInfo %>% left_join(signalPeptideInfo)

# Categorize and bind rows for different silk gene types (MaSp, MiSp, FlSp, etc.)
spiceProteins <- majorProteins %>% 
  select(-masp_concentration) %>%
  slice_head(n = 17)  %>% filter(grepl("SpiCE", geneNameProtein)) %>% 
  mutate(class = "SpiCE-LMa")
AmSpProteins <- majorProteins %>%
  select(-masp_concentration) %>%
  slice_head(n = 17)  %>% filter(grepl("AmSp-like", geneNameProtein)) %>% 
  mutate(class = "AmSp-like")
MaSpProteins <- geneInfo %>% 
  select(geneName) %>% 
  filter(grepl("MaSp", geneName)) %>% 
  mutate(geneNameProtein = geneName) %>% 
  mutate(class = "MaSp")
MiSpProteins <- geneInfo %>% 
  select(geneName) %>% 
  filter(grepl("MiSp", geneName)) %>% 
  mutate(geneNameProtein = geneName) %>% 
  mutate(class = "MiSp")
FlSpProteins <- geneInfo %>% 
  select(geneName) %>% 
  filter(grepl("FlSp", geneName)) %>% 
  mutate(geneNameProtein = geneName) %>% 
  mutate(class = "FlSp")

# Combine all silk gene dataframes into a single table
silkGenes <- bind_rows(spiceProteins, AmSpProteins, MaSpProteins, MiSpProteins, FlSpProteins)
```

## 🧬 Proteomics Data Processing

This section handles the reading, cleaning, and quality control of proteomics data from different spider glands.

### Read Raw Proteomics Data

```{r read-data}
# Read in the raw MaxQuant data for major glands
MQ_data.major <- read_tsv("../../data/proteomics/majorSilk/txt/proteinGroups.txt")
MQ_data.major.intensity <- MQ_data.major %>% 
  select(1,2,120:133) %>%
  rename(Ma_Dope_2 = "LFQ intensity MA_dope_2",
         Ma_Dope_3 = "LFQ intensity MA_dope_3",
         Ma_Gland_1 = "LFQ intensity MA_gland_1",
         Ma_Gland_2 = "LFQ intensity MA_gland_2",
         Ma_Gland_3 = "LFQ intensity MA_gland_3",
         Ma_SilkHFIP_1 = "LFQ intensity Native_Silk_HFIP_1",
         Ma_SilkHFIP_2 = "LFQ intensity Native_Silk_HFIP_2",
         Ma_SilkHFIP_3 = "LFQ intensity Native_Silk_HFIP_3",
         Ma_SilkLiBr_1 = "LFQ intensity Native_Silk_LiBr_1",
         Ma_SilkLiBr_2 = "LFQ intensity Native_Silk_LiBr_2",
         Ma_SilkLiBr_3 = "LFQ intensity Native_Silk_LiBr_3",
         Ma_Silk_1 = "LFQ intensity Native_Silk_Urea_1",
         Ma_Silk_2 = "LFQ intensity Native_Silk_Urea_2",
         Ma_Silk_3 = "LFQ intensity Native_Silk_Urea_3",
         geneName = "Protein IDs") %>%
  select(geneName,Ma_Dope_2,Ma_Dope_3,Ma_Gland_1,Ma_Gland_2,Ma_Gland_3,Ma_Silk_1,Ma_Silk_2,Ma_Silk_3)

# Read in the raw MaxQuant data for minor glands
MQ_data.minor <- read_tsv("../../data/proteomics/minorSilk/txt/proteinGroups.txt")
MQ_data.minor.intensity <- MQ_data.minor %>% 
  select(1,106:117) %>%
  rename(Mi_Dope_1 = "LFQ intensity Dope_spider_1",
         Mi_Dope_2 = "LFQ intensity Dope_spider_2",
         Mi_Dope_3 = "LFQ intensity Dope_spider_3",
         Mi_Dope_4 = "LFQ intensity Dope_spider_4",
         Mi_Gland_1 = "LFQ intensity Gland_spider_1",
         Mi_Gland_2 = "LFQ intensity Gland_spider_2",
         Mi_Gland_3 = "LFQ intensity Gland_spider_3",
         Mi_Gland_4 = "LFQ intensity Gland_spider_4",
         Mi_Gland_5 = "LFQ intensity Intact_Mi_Gland_spider_5",
         Mi_Gland_6 = "LFQ intensity Intact_Mi_Gland_spider_6",
         Mi_Silk_1 = "LFQ intensity Silk_Mi_1",
         Mi_Silk_2 = "LFQ intensity Silk_Mi_2",
         geneName = "Protein IDs")

# Read in the raw MaxQuant data for flagelliform glands
MQ_data.intensity.Fl <- read_tsv("../../data/proteomics/flageliform/proteinGroups.txt") %>%
  select(1, 50:53) %>%
  rename(Fl_Gland_2 = "LFQ intensity Flagelli_Gland_spider_2",
         Fl_Gland_3 = "LFQ intensity Flagelli_Gland_spider_3",
         Fl_Gland_4 = "LFQ intensity Flagelli_Gland_spider_4",
         Fl_Gland_5 = "LFQ intensity Intact_F_Gland_spider_5",
         geneName = "Protein IDs")

write_csv(MQ_data.intensity.Fl,"../../data/proteomics/flageliform_LFQ_int.csv.gz")

# Read in the raw MaxQuant data for aggregate glands
MQ_data.intensity.Ag <- read_tsv("../../data/proteomics/aggregateGland/txt/proteinGroups.txt") %>%
  select(1,  106:117)
newCN <- gsub("LFQ intensity ","", colnames(MQ_data.intensity.Ag))
newCN <- gsub("TF1_","", newCN)
colnames(MQ_data.intensity.Ag) <- newCN
MQ_data.intensity.Ag <- MQ_data.intensity.Ag %>% rename(geneName = "Protein IDs")
```

### Data Cleaning and Quality Control

This section tidies the raw data, filters out contaminants, and calculates relative protein intensities for each gland.

```{r qc-major-minor-flageliform-aggregate}
# Major Gland QC
MQ_data.MaSp <- MQ_data.major.intensity %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE) %>%
  mutate(Type = "Other",
         Type = replace(Type, grepl("MaSp[1-4]",geneName), "MaSp"),
         Type = replace(Type, grepl("CON_",geneName), "Contaminants"),
         Type = replace(Type, grepl("REV_",geneName), "FALSE"))

MQ_data.MaSp.QC <- MQ_data.MaSp %>%
  filter(intensity > 1000000, Type != "Contaminants", Type != "FALSE") %>%
  group_by(Gland, spider, Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.MaSp) %>%
  mutate(relativeIntensity = intensity / TotalIntensity)

MQ_data.MaSp.QC.summary <- MQ_data.MaSp.QC %>% 
  group_by(geneName, Part) %>%
  summarise(relativeIntensity = mean(relativeIntensity)) %>%
  filter(relativeIntensity > 0.0001) %>% 
  mutate(gland = "Major")

# Minor Gland QC
MQ_data.MiSp <- MQ_data.minor.intensity %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE) %>%
  mutate(Type = "Other",
         Type = replace(Type, grepl("MiSp",geneName), "MiSp"),
         Type = replace(Type, grepl("CON_",geneName), "Contaminants"),
         Type = replace(Type, grepl("REV_",geneName), "FALSE"))

MQ_data.MiSp.QC <- MQ_data.MiSp %>%
  filter(intensity > 10000, Type %in% c("Other", "MiSp"), Sample != "Mi_Silk_1") %>%
  group_by(Gland, spider, Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.MiSp) %>%
  mutate(relativeIntensity = intensity / TotalIntensity)

MQ_data.MiSp.QC.summary <- MQ_data.MiSp.QC %>% 
  group_by(geneName, Part) %>%
  summarise(relativeIntensity = mean(relativeIntensity)) %>%
  filter(relativeIntensity > 0.0001) %>% 
  mutate(gland = "Minor")

# Combine and save major and minor summaries
bind_rows(MQ_data.MiSp.QC.summary, MQ_data.MaSp.QC.summary) -> MQ_data.AmpGl.QC.summary
write_csv(MQ_data.AmpGl.QC.summary, "../../data/proteomics/MQ_data.AmpGl.QC.summary.csv")

# Flagelliform Gland QC
MQ_data.FlSp <- MQ_data.intensity.Fl %>% 
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("Gland","Part","spider"), remove = FALSE) %>%
  mutate(Type = "Other",
         Type = replace(Type, grepl("FlSp",geneName), "FlSp"),
         Type = replace(Type, grepl("FlSp-like",geneName), "FlSp-like"),
         Type = replace(Type, grepl("AcSp",geneName), "AcSp"),
         Type = replace(Type, grepl("CON_",geneName), "Contaminants"),
         Type = replace(Type, grepl("REV_",geneName), "FALSE"))

MQ_data.FlSp.QC <- MQ_data.FlSp %>%
  filter(intensity > 10000, Type %in% c("Other", "FlSp", "AcSp", "FlSp-like")) %>%
  group_by(Gland, spider, Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.FlSp) %>%
  mutate(relativeIntensity = intensity / TotalIntensity)

# Aggregate Gland QC
MQ_data.Ag <- MQ_data.intensity.Ag %>% 
  select(-2) %>%
  pivot_longer(names_to = "Sample", values_to = "intensity", -geneName) %>%
  separate(Sample, into = c("ID","spider","Side","Part"),sep = "_", remove = FALSE) %>%
  mutate(Type = "Other",
         Type = replace(Type, grepl("AgSp",geneName), "AgSp"),
         Type = replace(Type, grepl("Spidroin2",geneName), "AgSp-like"),
         Type = replace(Type, grepl("AcSp",geneName), "AcSp"),
         Type = replace(Type, grepl("CON_",geneName), "Contaminants"),
         Type = replace(Type, grepl("REV_",geneName), "FALSE"))

MQ_data.Ag.QC <- MQ_data.Ag %>%
  filter(intensity > 10000, Type %in% c("Other", "AgSp","AcSp", "AgSp-like")) %>%
  group_by(Side, spider, Part) %>%
  summarise(TotalIntensity = sum(intensity)) %>%
  inner_join(MQ_data.Ag) %>%
  mutate(relativeIntensity = intensity / TotalIntensity)
```

## 📈 Data Visualization and Comparison

This section contains code for plotting and comparing the normalized and relative intensities of proteins across different glands and tissue parts.

### Normalized Intensities

```{r normalized-intensities-analysis}
# Combine and normalize data
Proteomics.data <- bind_rows(MQ_data.FlSp.QC, MQ_data.MaSp.QC, MQ_data.MiSp.QC)
Proteomics.data.summary <- Proteomics.data %>%
  mutate(logInt = log10(intensity+1)) %>%
  filter(logInt > 3) %>%
  group_by(geneName, Gland, Part) %>%
  summarise(n = n(), intensity = sum(logInt)/n) %>%
  separate_longer_delim(geneName, delim = ";")

Proteomics.data.summary <- Proteomics.data.summary %>%
  group_by(Gland, Part) %>%
  summarise(mean = mean(intensity), sd = sd(intensity)) %>%
  ungroup() %>%
  inner_join(Proteomics.data.summary) %>%
  mutate(int_norm = (intensity - mean)/sd )
Proteomics.data.summary <- Proteomics.data.summary %>% left_join(silkGenes)

# Separate silk proteins for specific plots
Proteomics.data.silk.summary <- Proteomics.data.summary %>% filter(!is.na(geneNameProtein))
Proteomics.data.other.summary <- Proteomics.data.summary %>% filter(is.na(geneNameProtein))

# Gland-to-gland comparison plots
Proteomics.data.silk.gland.summary <- Proteomics.data.silk.summary %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneNameProtein, class, Gland, int_norm) %>%
  pivot_wider(names_from = "Gland", values_from = "int_norm", values_fill = -2)
Proteomics.data.other.gland.summary <- Proteomics.data.other.summary %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneName, Gland, class, int_norm) %>%
  pivot_wider(names_from = "Gland", values_from = "int_norm", values_fill = -2) %>%
  inner_join(signalPeptideInfo) %>%
  mutate(Term = replace(Term,is.na(Term),"NoSignalPeptide"))

ggplot(data = Proteomics.data.other.gland.summary,  aes(x = Mi, y = Ma))+ 
  geom_point(aes(size = Term))+
  geom_point(data = Proteomics.data.silk.gland.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.gland.summary, aes(label = geneNameProtein, color = class ), force = 15)+
  geom_abline(slope = 1, intercept = c(0,0))+
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) -> glands.plot.ma.mi

ggplot(data = Proteomics.data.other.gland.summary,  aes(x = Ma, y = Fl))+ 
  geom_point(aes(size = Term))+
  geom_point(data = Proteomics.data.silk.gland.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.gland.summary, aes(label = geneNameProtein, color = class ), force = 15)+
  geom_abline(slope = 1, intercept = c(0,0))+
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) -> glands.plot.ma.fl

ggplot(data = Proteomics.data.other.gland.summary,  aes(x = Mi, y = Fl))+ 
  geom_point(aes(size = Term))+
  geom_point(data = Proteomics.data.silk.gland.summary, aes(color = class))+
  geom_text_repel(data = Proteomics.data.silk.gland.summary, aes(label = geneNameProtein, color = class ), force = 15)+
  geom_abline(slope = 1, intercept = c(0,0))+
  scale_size_manual(values = c(NoSignalPeptide = 0.1, signalP = 1)) -> glands.plot.mi.fl

glands.plot.ma.mi
glands.plot.ma.fl
glands.plot.mi.fl
```

### Relative Intensities

```{r relative-intensities-analysis}
# Re-combine data for relative intensity analysis
Proteomics.data <- bind_rows(MQ_data.FlSp, MQ_data.MaSp, MQ_data.MiSp)
Proteomics.data.mean <- Proteomics.data %>%
  filter(intensity > 1000) %>%
  group_by(geneName, Gland, Part) %>%
  summarise(n = n(), intensity = mean(intensity))

# Calculate relative intensity
Proteomics.data.relative <- Proteomics.data.mean %>%
  group_by(Gland, Part) %>%
  summarise(Total = sum(intensity)) %>%
  inner_join(Proteomics.data.mean) %>%
  mutate(relativeIntensity = intensity/Total)

# Separate silk proteins from others
Proteomics.data.relative <- Proteomics.data.relative %>% left_join(silkGenes)
Proteomics.silk.relative <- Proteomics.data.relative %>% filter(!is.na(geneNameProtein))
Proteomics.other.relative <- Proteomics.data.relative %>% filter(is.na(geneNameProtein))
min_intensity <- min(Proteomics.data.relative$relativeIntensity, na.rm = TRUE)

# Pivot data for plotting
Proteomics.silk.gland.relative <- Proteomics.silk.relative %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneNameProtein, class, Gland, relativeIntensity) %>%
  pivot_wider(names_from = "Gland", values_from = "relativeIntensity", values_fill = (min_intensity))
Proteomics.other.gland.relative <- Proteomics.other.relative %>%
  filter(Part == "Gland") %>%
  ungroup() %>%
  select(geneName, Gland, relativeIntensity) %>%
  pivot_wider(names_from = "Gland", values_from = "relativeIntensity", values_fill = (min_intensity))

# Create the plot with log scales and percentage labels
Gland.plot <- ggplot(data =Proteomics.other.gland.relative,  aes(x = Mi, y = Ma))+ 
  geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.gland.relative, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.gland.relative, aes(label = geneNameProtein, color = class ))
  
library(scales)
Gland.plot +
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent ) +
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent) +
  theme_light() + 
  ggtitle("Gland Major vs Minor") + 
  geom_abline(intercept = 0, slope = 1) -> Gland.plot.ma.mi



# Create the plot with log scales and percentage labels
Gland.plot <- ggplot(data =Proteomics.other.gland.relative,  aes(x = Mi, y = Fl))+ 
  geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.gland.relative, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.gland.relative, aes(label = geneNameProtein, color = class ))
  
library(scales)
Gland.plot +
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent ) +
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent) +
  theme_light() + 
  ggtitle("Gland Flagelliform vs Minor") + 
  geom_abline(intercept = 0, slope = 1) -> Gland.plot.fl.mi


# Create the plot with log scales and percentage labels
Gland.plot <- ggplot(data =Proteomics.other.gland.relative,  aes(x = Ma, y = Fl))+ 
  geom_point( size = 0.1)+
  geom_point(data = Proteomics.silk.gland.relative, aes(color = class))+
  geom_text_repel(data = Proteomics.silk.gland.relative, aes(label = geneNameProtein, color = class ))
  
library(scales)
Gland.plot +
  scale_y_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent ) +
  scale_x_continuous(trans = log10_trans(),limits = c(0.000001, 1), labels = percent) +
  theme_light() + 
  ggtitle("Gland Major vs Flagelliform") + 
  geom_abline(intercept = 0, slope = 1) -> Gland.plot.fl.ma




# Major vs minor
Gland.plot.ma.mi

# Major vs Flageliform
Gland.plot.fl.ma

# Minor vs Flageliform
Gland.plot.fl.mi


```

## 📊 Summary and Export
This section summarizes the findings and exports the processed data for future use.

```{r summary-and-export}
# Export the final processed data
write_csv(Proteomics.data.relative, "../../data/proteomics/Proteomics.data.relative.csv")
```





